### 2023-11-26 notebooks/prokaryote_sampling.Rmd
Initial commit

---
title: "LUCA sampling"
# knit: (function(input_file, encoding) {
#   rmarkdown::render(input_file,
date: "`r format(Sys.time(), '%d-%b-%Y')`"
author: "Giacomo Mutti"
output:
    rmdformats::html_clean:
      number_sections: false
      code_folding: hide
      self_contained: true
      thumbnails: false
      lightbox: false
    highlight: kate
---

<style>
h1, h2, h3, h4, h5, h6, legend {
    font-family: Helvetica, sans-serif;
    font-weight: 700;
    color: #00a276;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
knitr::opts_chunk$set(crop = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(ggtree)
library(ape)
library(tidyverse)
library(patchwork)
library(kableExtra)
source("scripts/functions.R")
```

# Prokaryotes

In this notebook we will get a representative prokaryotic taxon sampling for a LUCA dataset to compare deep homologies inferred by blast or foldseek. In order to do this we can use **[GTDB](https://gtdb.ecogenomic.org/)**. It has selected reference genomes for both *Bacteria* and *Archaea*. There is a species tree available for both kingdoms:

* **[bacterial tree](https://data.gtdb.ecogenomic.org/releases/latest/bac120.tree)**
* **[archaeal tree](https://data.gtdb.ecogenomic.org/releases/latest/ar53.tree)**

Each tree with associated metadata:

* **[bacterial metadata](https://data.gtdb.ecogenomic.org/releases/latest/bac120_metadata.tsv.gz)**
* **[archaeal metadata](https://data.gtdb.ecogenomic.org/releases/latest/ar53_metadata.tsv.gz)**

As all genomes in uniprot have structural predictions we also need to check if the genomes are available. We can look in the **[uniprot genomes](https://ftp.uniprot.org/pub/databases/uniprot/current_release/knowledgebase/reference_proteomes/STATS)** table.

```{r params, class.source = 'fold-show'}
min_completeness <- 90
max_contamination <- 5
stepwise_species <- 1/10
target_species <- 20
```
```{r files}
# meta
uniprot_genomes <- "data/meta/uniprot_genomes.tsv"
uniprot_ref <- "data/meta/uniprot_ref.tsv"
ar_meta_file <- "data/meta/ar53_metadata.tsv.gz"
bac_meta_file <- "data/meta/bac120_metadata.tsv.gz"
taxids_in_AF <- "data/meta/taxids_in_alphafold.txt"
# LUCA eukaryotes
euka <- "data/input_tables/Hsapopi.csv"
# trees
bac120_file <- "data/sptrees/bac120.tree"
ar53_file <- "data/sptrees/ar53.tree"
```

```{r read_files, cache=TRUE}
taxids_AF <- as.numeric(readLines(taxids_in_AF))

up_ref <- read_delim(uniprot_ref, show_col_types = FALSE)
up_genomes_raw <- read_delim(uniprot_genomes, show_col_types = FALSE, col_types = c("cncccccc")) %>% 
  left_join(select(up_ref, Proteome_ID, SUPERREGNUM), by = "Proteome_ID")

up_genomes <- up_genomes_raw %>%
  mutate(assembly=gsub(" .*", "", `Assembly(Previous_Assembly)`)) %>%
  filter(Proteome_ID %in% up_ref$Proteome_ID, Tax_ID %in% taxids_AF) %>% 
  select(Proteome_ID, Species_Name, assembly)

bac <- read.tree(bac120_file)
ar <- read.tree(ar53_file)

gtdb_cols <- c("accession", "checkm_contamination", "checkm_completeness",
               "genome_size", "gtdb_taxonomy", "gtdb_representative",
               "ncbi_assembly_level", "ncbi_genbank_assembly_accession",
               "ncbi_organism_name", "protein_count")

bac_meta <- read_delim(bac_meta_file, show_col_types = FALSE) %>%
  select(all_of(gtdb_cols)) %>%
  filter(gtdb_representative,
         checkm_completeness >= min_completeness,
         checkm_contamination <= max_contamination,
         ncbi_genbank_assembly_accession %in% up_genomes$assembly)

asgards <- c("GCA_016933575.1", "GCA_008000775.1")

ar_meta <- read_delim(ar_meta_file, show_col_types = FALSE) %>%
  select(all_of(gtdb_cols)) %>%
  filter((gtdb_representative & checkm_completeness >= min_completeness & checkm_contamination <= max_contamination &
         ncbi_genbank_assembly_accession %in% up_genomes$assembly) | ncbi_genbank_assembly_accession %in% asgards)

```

There are **`r length(bac$tip.label)`** reference bacterias and **`r length(ar$tip.label)`** archaeas in the species trees.

Of these, there are **`r length(unique(bac_meta$ncbi_genbank_assembly_accession))`** Bacterias and **`r length(unique(ar_meta$ncbi_genbank_assembly_accession))`** Archaea that are in **Uniprot** and **AF** and respect the completeness (greater then **`r min_completeness`** and contamination (lower than **`r max_contamination`**) criteria. Notably, as they are crucial, we included the two Asgard genomes shared by GTDB and Uniprot: `r paste(asgards , collapse=",")`, even though they did not pass the checkM criterias (although not by much).

We therefore subset the species trees with these genomes. We can use **[Treemer](https://bmcbioinformatics.biomedcentral.com/articles/10.1186/s12859-018-2164-8)**: a tree pruning algorithm optimized to keep as much as information as possible. the `-r` parameter is what determines the speed of the process. We decided to use 1/250 of the leaves (if 1000 leaves, remove 4 tips before recomputing distances, which is the slow part). We will keep `r target_species` tips. With `-mc -lm` option we can force treemer to include all asgards genomes.

```{r intermediate_trees}
ar_intermediate <- drop.tip(ar, ar$tip.label[-match(ar_meta$accession, ar$tip.label)])
bac_intermediate <- drop.tip(bac, bac$tip.label[-match(bac_meta$accession, bac$tip.label)])

mc <- length(asgards)
asgard_tips <- ar_meta[ar_meta$ncbi_genbank_assembly_accession %in% asgards, ]$accession
write_csv(tibble(asgard_tips, rep("Asgard", mc)), paste0(ar53_file, ".filtered.lm"), col_names = FALSE)

write.tree(ar_intermediate, paste0(ar53_file, ".filtered"))
write.tree(bac_intermediate, paste0(bac120_file, ".filtered"))
```

```{r treemer, cache=TRUE}
py="python3"
ar_resolution <- round(length(ar_intermediate$tip.label)*stepwise_species)
args=paste("scripts/Treemmer_v0.3.py", paste0(ar53_file, ".filtered"),
           "-X",target_species,"-r",ar_resolution,"-lm",paste0(ar53_file, ".filtered.lm"),"-mc",mc,"-np -v 0 -lp 0")
system2(command=py, args=args)
bac_resolution <- round(length(bac_intermediate$tip.label)*stepwise_species)
args=paste("scripts/Treemmer_v0.3.py", paste0(bac120_file, ".filtered"),
           "-X",target_species,"-r",bac_resolution,"-np -v 0 -lp 0")
system2(command=py, args=args)
```

### Reduced trees

```{r plot trees, fig.align = "center", fig.width=14, fig.height=8, out.width="100%"}
ar_final_file <- paste0(ar53_file, ".filtered_trimmed_tree_X_", target_species)
ar_final <- read.tree(ar_final_file)

bac_final_file <- paste0(bac120_file, ".filtered_trimmed_tree_X_", target_species)
bac_final <- read.tree(bac_final_file)

ar_plot <- fortify(ar_intermediate) %>%
  left_join(ar_meta, by = c("label"="accession")) %>%
  mutate(is_selected=label %in% ar_final$tip.label, is_asgard=label %in% asgard_tips) %>% 
  ggtree(layout = "circular") +
  ggtitle("Archaea selected genomes") +
  geom_tippoint(aes(color=checkm_completeness, shape=is_asgard), . %>% filter(is_selected), size=4) +
  xlim(c(0, max(vcv(ar_intermediate)))) +
  scale_color_gradientn(colours = wespal, limits=c(min_completeness-5, 100))

bac_plot <- fortify(bac_intermediate) %>%
  left_join(bac_meta, by = c("label"="accession")) %>%
  mutate(is_selected=label %in% bac_final$tip.label) %>% 
  ggtree(layout = "circular") +
  ggtitle("Bacteria selected genomes") +
  geom_tippoint(aes(color=checkm_completeness), . %>% filter(is_selected), size=4) +
  xlim(c(0, max(vcv(bac_intermediate)))) +
  scale_color_gradientn(colours = wespal, limits=c(min_completeness-5, 100))

trees_plot <- ar_plot + bac_plot & theme(legend.position = "bottom")
trees_plot + plot_layout(guides = "collect")
```

```{r final_table}
ar_final_table <- filter(ar_meta, accession %in% ar_final$tip.label) %>% 
  left_join(up_genomes, by=c("ncbi_genbank_assembly_accession"="assembly")) %>% 
  mutate(phylum=gsub(";.*", "", gsub(".*;p__", "", gtdb_taxonomy)), kingdom="Archaea") %>%
  select(Proteome_ID, ncbi_genbank_assembly_accession, Species_Name, phylum,
         protein_count, genome_size, checkm_completeness, checkm_contamination) %>% 
  arrange(phylum)

bac_final_table <- filter(bac_meta, accession %in% bac_final$tip.label) %>% 
  left_join(up_genomes, by=c("ncbi_genbank_assembly_accession"="assembly")) %>% 
  mutate(phylum=gsub(";.*", "", gsub(".*;p__", "", gtdb_taxonomy)), kingdom="Bacteria") %>% 
  select(Proteome_ID, ncbi_genbank_assembly_accession, Species_Name, phylum,
         protein_count, genome_size, checkm_completeness, checkm_contamination) %>% 
  arrange(phylum)

rbind(ar_final_table, bac_final_table) %>% 
  kbl(booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(bootstrap_options = "striped",
              font_size = 9,
              latex_options = c("hold_position","repeat_header"),position = "left")
```

# Run LUCA analysis

We will add then these 40 proteomes to the 18 Eukaryotes of the *H. sapiens Opisthokonta* dataset and create the file *data/input_tables/LUCA.csv* that will be then used in the pipeline.

```{r create_input_table}
read_delim(euka, delim = "\t", col_names = colnames(up_genomes_raw), show_col_types = FALSE) %>% 
  rbind(up_genomes_raw[up_genomes_raw$Proteome_ID %in% ar_final_table$Proteome_ID, ]) %>% 
  rbind(up_genomes_raw[up_genomes_raw$Proteome_ID %in% bac_final_table$Proteome_ID, ]) %>% 
  write_delim("data/input_tables/LUCA.csv", delim = "\t", col_names = FALSE)
```

First you need to download the structures, do this in local:

```{bash download, eval=FALSE, class.source = "fold-show"}
snakemake --configfile data/configs/LUCA.yaml -p -j2 --until download_pdbs -s workflow/download_data.smk
snakemake --configfile data/configs/LUCA.yaml -p -j2 --until download_uniprot_ids -s workflow/download_data.smk
```

Then run on MN the remaining steps to decompress:

```{bash untar, eval=FALSE, class.source = "fold-show"}
snakemake --configfile data/configs/LUCA.yaml -p -j24 -s workflow/download_data.smk
```

Finally, run the pipeline until the homology call:

```{bash smk, eval=FALSE, class.source = "fold-show"}
snakemake --configfile data/configs/LUCA.yaml -p -j24 --until plot_evalues
```

### 2023-11-26 scripts/cif2fasta.py

@@ -1,0 +1,794 @@
+#!/usr/bin/env python
+
+"""
+Created on Mon Jun 15 21:49:32 2015
+
+@author: Harald Voehringer
+"""
+
+from __future__ import print_function
+import sys, os, glob, gzip, textwrap, itertools
+from optparse import OptionParser
+from collections import defaultdict
+from os.path import splitext
+from pdbx.reader.PdbxReader import PdbxReader
+from multiprocessing import Pool
+
+DEBUG_MODE = False
+MIN_SEQ_LEN = None
+SCOP_LIBRARY = False
+
+THREE2ONE = {
+        'CYS': 'C', 'ASP': 'D', 'SER': 'S', 'GLN': 'Q', 'LYS': 'K', 'ILE': 'I', 'PRO': 'P', 
+        'THR': 'T', 'PHE': 'F', 'ASN': 'N', 'GLY': 'G', 'HIS': 'H', 'LEU': 'L', 'ARG': 'R', 
+        'TRP': 'W', 'ALA': 'A', 'VAL': 'V', 'GLU': 'E', 'TYR': 'Y', 'MET': 'M', 'MSE': 'M',
+        'HYP': 'P', 'MLY': 'K', 'SEP': 'S', 'TPO': 'T', 'CSO': 'C', 'PTR': 'Y', 'KCX': 'K',
+        'CME': 'C', 'CSD': 'A', 'CAS': 'C', 'MLE': 'L', 'DAL': 'A', 'CGU': 'E', 'DLE': 'L',
+        'FME': 'M', 'DVA': 'V', 'OCS': 'C', 'DPR': 'P', 'MVA': 'V', 'TYS': 'Y', 'M3L': 'K',
+        'SMC': 'C', 'ALY': 'K', 'CSX': 'C', 'DCY': 'C', 'NLE': 'L', 'DGL': 'E', 'DSN': 'S',
+        'CSS': 'C', 'DLY': 'K', 'MLZ': 'K', 'DPN': 'F', 'DAR': 'R', 'PHI': 'F', 'IAS': 'D',
+        'DAS': 'D', 'HIC': 'H', 'MP8': 'P', 'DTH': 'T', 'DIL': 'I', 'MEN': 'N', 'DTY': 'Y',
[…]

### 2023-11-26 scripts/compare_mcl.R

@@ -1,0 +1,26 @@
+blines <- readLines("results/mcl/blast.I12.test")
+blines <- blines[grepl("\\t", blines)]
+flines <- readLines("results/mcl/foldseek.I12.test")
+flines <- flines[grepl("\\t", flines)]
+
+list_b <- lapply(blines[1:1000], function(x) strsplit(x, "\\t")[[1]])
+list_f <- lapply(flines[1:1000], function(x) strsplit(x, "\\t")[[1]])
+
+# combns <- gtools::combinations(length(list_b), length(list_f), )
+nrows <- length(list_b)
+ncols <- length(list_f)
+
+result_matrix <- matrix(0, nrow = nrows, ncol = ncols)
+
+for (i in 1:nrows){
+    set_blast <- list_b[i][[1]]
+  for (j in 1:ncols){
+    set_fs <- list_f[j][[1]]
+    norm_value <- pmax(length(set_blast), length(set_fs))
+    result_matrix[i,j] <- length(intersect(set_blast, set_fs))/norm_value
+  }
+}
+
+
+heatmap(result_matrix)
+

### 2023-11-26 scripts/compare_sampling.R

@@ -1,0 +1,378 @@
+library("optparse")
+
+option_list <- list(
+  make_option(c("-i", "--input"),
+              type = "character", default = NULL,
+              help = "input table file", dest = "input"
+  ),
+  make_option(c("-m", "--meta"),
+              type = "character", default = NULL,
+              help = "meta file", dest = "meta"
+  ),
+  make_option(c("-t", "--taxidmap"),
+              type = "character", default = NULL,
+              help = "taxidmap file", dest = "taxidmap"
+  ),
+  make_option(c("-b", "--blast"),
+              type = "character", default = NULL,
+              help = "blast file", dest = "blast"
+  ),
+  make_option(c("-f", "--fs"),
+              type = "character", default = NULL,
+              help = "foldseek file", dest = "fs"
+  ),
+  make_option(c("--fb"),
+              type = "character", default = NULL,
+              help = "foldseek brh", dest = "fs_brh"
+  ),
+  make_option(c("--bb"),
+              type = "character", default = NULL,
+              help = "blast brh", dest = "blast_brh"
+  ),
+  make_option(c("-o", "--output"),
+              type = "character", default = NULL,
+              help = "eda plot filename", metavar = "character", dest = "outfile"
+  ),
+  make_option(c("--o2"),
+              type = "character", default = NULL,
+              help = "saturation plot filename", metavar = "character", dest = "outfile2"
+  ),
+  make_option(c("--sf"),
+              type = "character", default = NULL,
+              help = "self foldseek", dest = "self_fs"
+  ),
+  make_option(c("--sb"),
+              type = "character", default = NULL,
+              help = "self blast", dest = "self_blast"
+  ),
+  make_option(c("-e", "--evalue"),
+              type = "numeric", default = 1e-3,
+              help = "e value for tree sequences", dest = "min_evalue"
+  ),
+  make_option(c("-s", "--max_seqs"),
+              type = "integer", default = 150,
[…]

### 2023-11-26 scripts/compare_sampling_old.R

@@ -1,0 +1,197 @@
+library("optparse")
+
+option_list <- list(
+  make_option(c("-b", "--blast"),
+              type = "character", default = NULL,
+              help = "blast file", dest = "blast"
+  ),
+  make_option(c("-f", "--fs"),
+              type = "character", default = NULL,
+              help = "foldseek file", dest = "fs"
+  ),
+  make_option(c("--fb"),
+              type = "character", default = NULL,
+              help = "foldseek brh", dest = "fs_brh"
+  ),
+  make_option(c("--bb"),
+              type = "character", default = NULL,
+              help = "blast brh", dest = "blast_brh"
+  ),
+  make_option(c("-o", "--output"),
+              type = "character", default = NULL,
+              help = "output filename", metavar = "character", dest = "outfile"
+  )
+)
+
+desc <- "Produces informative plot on foldseek and blast results comparison"
+
+opt_parser <- OptionParser(option_list = option_list, description = desc)
+opt <- parse_args(opt_parser)
+
+library(tidyverse)
+library(patchwork)
+source("scripts/functions.R")
+theme_set(theme_bw())
+
+# opt <- NULL
+# opt$fs <- "results/Ncra_opistho/homology/UP000001805_fs.tsv"
+# opt$blast <- "results/Ncra_opistho/homology/UP000001805_blast.tsv"
+# opt$outfile <- "results/Ncra_opistho/plots/plot_intersection.png"
+# opt$fs_brh <- "results/Ncra_opistho/homology/UP000001805_fs_brh.tsv"
+# opt$blast_brh <- "results/Ncra_opistho/homology/UP000001805_blast_brh.tsv"
+
+fs <- read_delim(opt$fs, col_names = fs_columns, show_col_types = FALSE)
+blast <- read_delim(opt$blast, col_names = blast_columns, show_col_types = FALSE)
+
+seeds <- union(unique(fs$query), unique(blast$query))
+
+evals <- c(Inf, 1e-2, 1e-3, 1e-5)
+min_covs <- c(0, 30, 60)
+
+fs_brhs <- read_delim(opt$fs_brh, col_names = c("query", "target"), show_col_types = FALSE) %>% 
+  group_by(query) %>% 
+  summarise(fs_brhs=list(target))
+
[…]

### 2023-11-26 scripts/compare_trees.R

@@ -1,0 +1,126 @@
+library("optparse")
+
+option_list <- list(
+  make_option(c("--seq"),
+              type = "character", default = NULL,
+              help = "sequence trees", dest = "seq"
+  ),
+  make_option(c("--struct"),
+              type = "character", default = NULL,
+              help = "structure trees", dest = "struct"
+  ),
+  make_option(c("--structml"),
+              type = "character", default = NULL,
+              help = "structure ml trees", dest = "structml"
+  ),
+  make_option(c("-r", "--ranger"),
+              type = "character", default = NULL,
+              help = "ranger results", dest = "ranger"
+  ),
+  make_option(c("-o", "--output"),
+              type = "character", default = NULL,
+              help = "output filename", metavar = "character", dest = "outfile"
+  )
+)
+
+desc <- "Produces informative plot on trees comparison"
+
+opt_parser <- OptionParser(option_list = option_list, description = desc)
+opt <- parse_args(opt_parser)
+
+library(tidyverse)
+library(ggtree)
+library(patchwork)
+library(ggpubr)
+
+# opt <- NULL
+# opt$ranger <- "test/tcs/prova"
+# struct_trees <- "results/Ncra_opistho/trees/UP000001805_fs_foldtrees.txt"
+# structml_trees <- "results/Ncra_opistho/trees/UP000001805_fs_mltrees.txt"
+# seq_trees <- "results/Ncra_opistho/trees/UP000001805_blast_trees.txt"
+# ranger_result <- "results/Ncra_opistho/reco/UP000001805_ranger.tsv"
+
+theme_set(theme_bw())
+
+df_struct <- read_delim(opt$struct, col_names = c("gene", "tree_struct"), show_col_types = FALSE)
+# mutate(.id=factor(row_number()))
+df_structml <- read_delim(opt$structml, col_names = c("gene", "tree_structml"), show_col_types = FALSE) 
+df_seq <- read_delim(opt$seq, col_names = c("gene", "tree_seq"), show_col_types = FALSE)
+# mutate(.id=factor(row_number()))
+
+df <- full_join(df_struct, df_seq) %>% 
+  full_join(df_structml) %>% 
[…]

### 2023-11-26 scripts/explore_lddt.R

@@ -1,0 +1,15 @@
+library(tidyverse)
+
+theme_set(theme_bw())
+
+seed <- "UP000001805"
+
+lddt_file <- "results/Ncra_opistho/db/db_lddt.tsv"
+lddt_proteomes <- read_delim(lddt_file)
+
+ggplot(lddt_proteomes, aes(fct_reorder(proteome, mean_lddt), mean_lddt)) + 
+  geom_violin(fill="#35B779") + 
+  geom_boxplot(width=0.1, alpha=0, color="black", outlier.alpha = 1) +
+  ylim(c(0,100)) +
+  theme(axis.text.x = element_text(angle=70, hjust=1))
+

### 2023-11-26 scripts/functions.R

@@ -1,0 +1,29 @@
+blast_columns <- c("query","target","pident","length","mismatch","gapopen",
+                   "qstart","qend","sstart","send","evalue","bitscore",
+                   "qcov","qcovhsp","qlen","slen","staxids")
+
+fs_columns <- c("query","target","pident","length","mismatch","gapopen",
+                "qstart","qend","sstart","send","evalue","bitscore",
+                "lddt","alntmscore","rmsd","prob","qcov","tcov")
+
+
+wespal <- wesanderson::wes_palette("Zissou1", 100, type = "continuous")
+
+
+add_marginal <- function(df, plot, x, y, fill, palette){
+  xdens <- axis_canvas(plot, axis = "x")+
+    geom_density(data = df, aes(x = .data[[x]], fill = .data[[fill]]),
+                 alpha = 0.7, size = 0.2) +
+    coord_cartesian(expand=0) + 
+    scale_fill_manual(values = palette)
+  ydens <- axis_canvas(plot, axis = "y", coord_flip = TRUE) +
+    geom_density(data = df, aes(x = .data[[y]], fill = .data[[fill]]),
+                 alpha = 0.7, size = 0.2)+
+    coord_flip(expand = 0) +
+    scale_fill_manual(values = palette)
+  
+  p1 <- insert_xaxis_grob(plot, xdens, grid::unit(.2, "null"), position = "top")
+  p2 <- insert_yaxis_grob(p1, ydens, grid::unit(.2, "null"), position = "right")
+  ggdraw(p2)
+}
+

### 2023-11-26 scripts/get_BRH.py

@@ -1,0 +1,71 @@
+#!/usr/bin/env python
+
+#Script used to obtain the BRH of a pair of reciprocal blast results
+
+import argparse
+
+#Load blast results into memory after filtering based on evalue and overlap thresholds
+def filter_blast(BlastFile, threshold, evalue_thr=1e-05, overlap_thr=0.2):
+	hits = {}
+	stats = {}
+	# if evalue_thr > 0.1:
+	# 	print("WARNING: High e-value threshold!")
+	for line in open(BlastFile):
+		line = line.strip()
+		dades = line.split("\t")
+		if "#" not in line:
+			# try:
+			# 	overlap = (float(dades[7]) - float(dades[6])) / float(len(seqs[dades[0]]))
+			# except:
+			# 	overlap = 0.0
+			# if overlap > overlap_thr and float(dades[10]) < evalue_thr:
+			if float(dades[10]) < float(evalue_thr):
+				if dades[0] not in hits:
+					hits[dades[0]] = []
+					# stats[dades[0]] = {}
+				if dades[1] not in hits[dades[0]] and len(hits[dades[0]]) < threshold:
+					hits[dades[0]].append(dades[1])
+					# stats[dades[0]][dades[1]] = {}
+					# stats[dades[0]][dades[1]]["overlap"] = overlap
+					# stats[dades[0]][dades[1]]["eval"] = float(dades[10])
+					# stats[dades[0]][dades[1]]["ident"] = float(dades[2])
+	return hits
+
+#Get best reciprocal hits
+def get_BRH(hits1,hits2):
+	BBH = set([])
+	taken = set([])
+	for c1 in hits1:
+		h1 = hits1[c1][0]
+		if h1 in hits2:
+			if hits2[h1][0] == c1:
+				# if c1 > h1:
+				pair = c1 +"\t"+ h1
+				# else:
+				# 	pair = h1 + "\t"+c1
+				BBH.add(pair)
+				taken.add(h1)
+				taken.add(c1)
+	return BBH,taken
+
+
+if __name__ == "__main__":
+	parser = argparse.ArgumentParser(description="Obtains pairs of BRH")
+	parser.add_argument('-h1', dest='h1',required=True,
+						help=('blast where seed species is species 1.'),
+						nargs='*')
+	parser.add_argument('-h2', dest='h2',required=True,
+						help=('blast where seed species is species 2.'),
+						nargs='*')
+	parser.add_argument("-e", dest="evalue",
[…]

### 2023-11-26 scripts/mask_structures.py

@@ -1,0 +1,40 @@
+#!/usr/bin/env python
+import argparse
+from Bio import SeqIO
+import pandas as pd
+import glob
+
+
+# Parse data
+parser = argparse.ArgumentParser(
+    description="Mask a fasta file based on structural confidence prediction. Residues with low lddt will be Xs"
+)
+
+parser.add_argument("-i", "--in", dest="infile", default="None", 
+                    help="input file",required=True)
+parser.add_argument("-o", "--out", dest="out", default="None", 
+                    help="output file",required=True)
+parser.add_argument("-s", "--struct", dest="struct", default="None", 
+                    help="structure directory",required=True)
+parser.add_argument('-m', '--min', dest='min',
+                    help='Minimum lddt', type=int, default=50)
+
+if __name__ == '__main__':
+    args = parser.parse_args()
+    min_lddt = args.min
+
+    out_seqs = []
+
+    for record in SeqIO.parse(args.infile, "fasta"):
+        gene_id = record.name
+        json_conf = glob.glob(args.struct+'/*/confidence/'+gene_id+'-confidence_v4.json.gz')[0]
+
+        confidence = pd.read_json(json_conf)
+        low_confidence_sites = list(confidence[confidence['confidenceScore']<min_lddt]['residueNumber'])
+
+        for site in low_confidence_sites:
+            record.seq = record.seq[:site-1] + 'X' + record.seq[site:]
+
+        out_seqs.append(record)
+
+    SeqIO.write(out_seqs, args.out, "fasta")

### 2023-11-26 software/foldtree/compute_scores.py

@@ -1,0 +1,100 @@
+import treescore
+import toytree
+import pandas as pd
+import subprocess
+import argparse
+import numpy as np
+
+from pathlib import Path
+# Parse data
+parser = argparse.ArgumentParser(
+    description="Run Ranger DTL to compute most parsimonious number of Duplication and Losses"
+)
+
+parser.add_argument('--trees', dest='trees',required=True,
+						help=('tree files for each method'),
+						nargs='*')
+# parser.add_argument("-i", "--in", dest="infile", default="None", 
+#                     help="input file",required=True)
+parser.add_argument("-o", "--out", dest="out", default="None", 
+                    help="output file", required=True)
+parser.add_argument("-t", "--tax", dest="taxidmap", default="None", 
+                    help="taxidmap file", required=True)
+parser.add_argument("-l", "--lineage", dest="lineage", default="None", 
+                    help="lineage file", required=True)
+parser.add_argument("-s", "--sptree", dest="sptree", default="None", 
+                    help="species tree",required=True)
+parser.add_argument("--tmp", dest="tmp", default="/tmp/Ranger_input.txt", 
+                    help="tmp file for Ranger")
+
+
+# def get_species_name(nodename):
+#     taxid = taxidmap[taxidmap['query']==nodename].iloc[0]['species']
+#     mnemo = meta[meta['taxid']==taxid].iloc[0]['mnemo']
+#     return mnemo
+
+T_score = 2000
+
+if __name__ == '__main__':
+    args = parser.parse_args()
+
+    uniprot_df = pd.read_csv(args.lineage, sep="\t", names=['query', 'mnemo','tax'])
+    taxidmap = pd.read_csv(args.taxidmap, names=['query', 'species'], sep="\t")
+
+    sptree = toytree.tree(args.sptree)
+    tmpfile = args.tmp
+
+    df = []
+
+    for inline in args.trees:
+        method = Path(inline).stem.split('_',1)[1]
+        print(method)
+        with open(inline) as tf:
+            for line in tf:
+                line = line.strip().split()
+                id = line[0]
+                tree = toytree.tree(line[1], format = 0)
+                tree = treescore.label_leaves(tree, taxidmap, uniprot_df)
+                overlap = treescore.getTaxOverlap(tree.treenode)
+
+                leaves = tree.get_tip_labels()
+                n_sps = len(set(leaves))
+                n_tips = len(leaves)
+
+                with open(tmpfile, 'w') as outfile:
+                    outfile.write(sptree.write(tree_format=9)+'\n')
[…]

### 2023-11-26 software/foldtree/foldseek2tree.py

@@ -1,0 +1,243 @@
+#foldssek cluster
+import subprocess ,shlex
+import numpy as np
+from scipy.spatial.distance import cdist
+import statsmodels
+import toytree
+import pandas as pd
+import re
+import os
+from scipy.stats import chi2
+import argparse
+
+def consensustree(treelist):
+	'''get a consensus tree from a list of tree files
+	
+	Parameters
+	----------
+	treelist : list
+		list of tree files
+
+	'''
+	#creat a multitree object from the list of tree files
+	treelist = [toytree.tree(i , format = 0 ) for i in treelist]
+	mt = toytree.mtree(treelist)
+	#get the consensus tree
+	ct = mt.get_consensus_tree( )
+	return ct
+
+#smooth distmat with MDS
+def MDS_smooth(distmat):
+	mds = MDS(n_components=int(distmat.shape[0]/2) )#, dissimilarity='precomputed'  )
+	distmat = mds.fit_transform(1-distmat )
+	distmat = cdist(distmat,distmat, 'minkowski', p=1.5 )
+	return distmat
+
+
+
+def Tajima_dist(kn_ratio, bfactor=1, iter = 100, delta=10**-5):
+	#approx of b*log(1-p/b) from https://doi.org/10.1093/oxfordjournals.molbev.a040031
+	#set all entries to smaller than bfactor
+	kn_ratio[kn_ratio > bfactor] = bfactor - delta
+	taj = np.add.reduce([ (kn_ratio**(np.ones(kn_ratio.shape)*i) )/ (bfactor**(i-1)*i) for i in range(1,iter) ] )
+	#set diagonal to 0
+	np.fill_diagonal(taj, 0)
+	return taj
+
+
+def runargs(args):
+	'''run a command line command
+	
+	Parameters
+	----------
+	args : str
+		command line command
+	'''
+	
+	args = shlex.split( args)
+	p = subprocess.run( args )
+	return p
+	
+def runFoldseekdb(folder , outfolder , foldseekpath = 'foldseek'):
+	'''run foldseek createdb
+	
+	parameters
+	----------
+	folder : str
+		path to folder with pdb files
+	outfolder : str 
+		path to output folder
+	
+
+	'''
+	args = foldseekpath + ' createdb '+  folder + ' '+ outfolder+'structblobDB '
+	p = runargs(args)
+	return outfolder+'structblobDB '
+
[…]

### 2023-11-26 software/foldtree/foldtree.py

@@ -1,0 +1,136 @@
+
+import foldseek2tree
+import corecut 
+import argparse
+import numpy as np
+import pandas as pd
+import os
+
+def structblob2tree(input_folder, outfolder, tmpfolder, corefolder,
+                    overwrite = False, correction = False,
+                    fastmepath = 'fastme', foldseekpath = 'foldseek', delta = 0.0001,
+                    kernel = 'fident', core = False, hitthresh = .8, minthresh = .6):
+    '''
+    run fold tree pipeline for a folder of pdb files
+    
+    Parameters
+    ----------
+    input_folder : str
+        path to folder with pdb files   
+    outfolder : str
+        path to output folder   
+    overwrite : bool
+        overwrite existing foldseek output  
+    fastmepath : str    
+        path to fastme executable
+    quicktreepath : str 
+        path to quicktree executable
+    foldseekpath : str  
+        path to foldseek executable 
+    delta : float   
+        small number to replace negative branch lengths with, default is .0001
+    correction : str    
+        correction method to use, either 'tajima' or 'none'
+    kernel : str    
+        kernel to use, either 'fident', 'lddt' or 'alntmscore'
+    
+    '''
+    #check if the foldseek output is already there
+    input_folder = os.path.join(input_folder, '')
+    if os.path.exists(outfolder + '_allvall.tsv') and overwrite == False:
+        print('found foldseek output, skipping foldseek')
+        alnres = outfolder + '_allvall.tsv'
+    else:
+        alnres = foldseek2tree.runFoldseek_allvall_EZsearch(input_folder, outfolder + '_allvall.tsv', tmpfolder, foldseekpath = foldseekpath)
+    
+    if core == True:
+        corefolder = os.path.join(corefolder, '')
+        corecut.extract_core(alnres, outfolder+'.core',  hitthresh = hitthresh, minthresh = minthresh, 
+                             corefolder = corefolder, structfolder = input_folder)
+        if os.path.exists(outfolder + '_core_allvall.tsv') and overwrite == False:
+            print('found foldseek core output, skipping foldseek')
+            alnres = outfolder + '_core_allvall.tsv'
+        else:
+            alnres = foldseek2tree.runFoldseek_allvall_EZsearch(corefolder, outfolder + '_core_allvall.tsv', tmpfolder, foldseekpath = foldseekpath)
+
[…]

### 2023-11-26 software/foldtree/treescore.py

@@ -1,0 +1,261 @@
+import toytree
+import toyplot
+
+#taxonomy overlap score
+
+def standard_treedraw( tre, sizes= None , colors= None ,fixed_order=None, fixed_position=None , 
+ts = None,  save_file = False  , tiplabels = None ,  layout='c', edge_type='p' ):
+	
+	'''
+	standard treedraw function
+	
+	Parameters
+	----------
+	tre : toytree.tree
+		tree to draw
+	sizes : list
+		list of node sizes
+	colors : list
+		list of node colors
+	fixed_order : list
+		list of node names in order to draw
+	fixed_position : list
+		list of node positions
+	ts : treestlye : str
+		style of tree to draw (default = 'c')
+	save_file : str
+		path to save file
+	tiplabels : list
+		list of tip labels to draw
+	layout : str
+		layout of tree to draw (default = 'c')
+	edge_type : str
+		type of edge to draw (default = 'p')
+	'''
+
+	
+	if tiplabels is None:
+		tiplabels = tre.get_tip_labels()
+	canvas, axes, mark = tre.draw(  
+		ts = ts,
+		node_sizes=sizes,
+		node_colors=colors,
+		tip_labels_align=True,
+		scalebar=True,
+		fixed_order=fixed_order, 
+		fixed_position=fixed_position,
+		tip_labels=tiplabels,
+		layout=layout,
+		edge_type=edge_type,
+		tip_labels_style={
+			"fill": "#262626",
+			"font-size": "9px"}
+	)
+	if save_file:
+		toyplot.svg.render(canvas, save_file)
+
+
+
+def exp_score( v , exp = 1.5):
+	return v**exp
+
+def frac_score( v ):
+	return v+1
+
+#taxonomy overlap score
+def getTaxOverlap(node , treelen  = None , scorefun = frac_score):
+	
+	"""
+    Calculate the taxonomy overlap score for the given node in a phylogenetic tree.
+    
+    The taxonomy overlap score is defined as the number of taxonomic labels shared by all the leaf nodes
+    descended from the given node, plus the sum of the scores of all its children. If a leaf node has no
+    taxonomic label, it is not counted towards the score. The function also calculates the size of the
+    largest loss in lineage length, defined as the difference between the length of the set of taxonomic
+    labels shared by all the leaf nodes and the length of the longest set of taxonomic labels among the
+    children of the node.
+    
[…]

### 2024-05-28 workflow/notebooks/eda.Rmd
added notebooks, FTPY and some code improvements

---
title: "Structural analysis"
# knit: (function(input_file, encoding) {
#   rmarkdown::render(input_file,
date: "`r format(Sys.time(), '%d-%b-%Y')`"
author: "Giacomo Mutti"
output:
    rmdformats::html_clean:
        number_sections: false
        code_folding: hide
        self_contained: true
        fig_caption: true
        # thumbnails: false
        # lightbox: false
        gallery: true
        toc_depth: 3
        highlight: kate
---

```{r setup, include=FALSE, dev="CairoPNG"}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, cache.lazy = FALSE,
                      crop = TRUE, dev="CairoPNG")
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
knitr::opts_knit$set(root.dir = "../../")
```

```{r libraries, message=FALSE, warning=FALSE, cache=FALSE}
library(tidyverse)
library(patchwork)
library(ggtree)
library(cowplot)
library(glue)
source("workflow/scripts/functions.R")
theme_set(theme_classic())

models <- names(palettes_model)

params <- yaml::read_yaml("config/params.yaml")
params_dataset <- yaml::read_yaml("config/Hsapopi_test.yaml")

outdir <- glue(params$outdir, params_dataset$dataset)
```

```{r read_stuff, cache=FALSE}
seeds <- readLines(glue(outdir, "/ids/UP000005640_common.ids"))

tax <- read_delim(params_dataset$taxons_file, show_col_types=F)
taxmap <- read_delim(glue(outdir, "/db/taxidmap"), 
                     delim = "\t", col_names = c("label", "Tax_ID"),
                     show_col_types = FALSE) %>% 
  left_join(tax, by = "Tax_ID") %>% 
  dplyr::select(label, Proteome_ID, mnemo)

taxidmap <- read_delim(glue(outdir,"/db/taxidmap"), 
                       col_names = c("target", "Tax_ID"), show_col_types=F)

sptree <- read.tree(params_dataset$species_tree_labels)
translate <- deframe(distinct(dplyr::select(taxmap, Proteome_ID, mnemo)))
# sptree$tip.label <- names(translate)[match(translate, sptree$tip.label)]

species_dist_df <- tibble()
for (sp in sptree$tip.label){
  if (sp=="Hsap") {next}
  dist <- castor::get_all_distances_to_tip(sptree, "Hsap")[MRCA(sptree, "Hsap", sp)]
  mrca <- c(sptree$tip.label, sptree$node.label)[MRCA(sptree, "Hsap", sp)]
  species_dist_df <- bind_rows(species_dist_df, tibble(mnemo=sp, value=dist, mrca=mrca))
}

# read input table
lvls_tax <- rev(c(unique(tax$Clade), "archaea", "bacteria"))

table_columns <- c('Proteome_ID', 'Tax_ID', 'count1', 'count2', 'count3', 
                   'genome', 'source', 'species', 'mnemo')

table <- read_delim(params_dataset$input_table_file, show_col_types=F, 
                    delim = "\t", col_names = table_columns) %>% 
  left_join(tax, by = "Tax_ID") %>% 
  mutate(Clade = ifelse(is.na(Clade), mnemo, Clade),
         Clade = factor(Clade, levels = lvls_tax))
```

# Homology

```{r read_self_blast}
blast_self <- read_delim(glue(outdir, "/homology/allvall/",
                              params_dataset$seed, "_", params_dataset$seed, "_blast.tsv"),
                         show_col_types=FALSE,
                         col_names = blast_columns) %>% 
  filter(query %in% seeds) %>%
  filter(query==target) %>% 
  mutate(blast_selfbit=bitscore) %>% 
  select(query, blast_selfbit)

fs_self <- read_delim(glue(outdir, "/homology/allvall/",
                              params_dataset$seed, "_", params_dataset$seed, "_fs.tsv"), 
                      show_col_types = FALSE,
                         col_names = blast_columns) %>% 
  filter(query %in% seeds) %>%
  filter(query==target) %>% 
  mutate(fs_selfbit=bitscore) %>% 
  select(query, fs_selfbit)

self <- full_join(blast_self, fs_self, by="query")
```

```{r read_fs}
fs <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_fs.tsv"), 
                 col_names = fs_columns, show_col_types = FALSE) %>% 
  filter(query %in% seeds) %>%
  mutate(evalue = ifelse(evalue==0, 1e-180, evalue)) %>%
  left_join(taxidmap, by = "target")

fs_brhs <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_fs_brh.tsv"),
                     col_names = c("query", "target"), show_col_types = FALSE) %>% 
  filter(query %in% seeds) %>%
  mutate(method="fs_brh") 
```

```{r read_blast}
blast <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_blast.tsv"), 
                    col_names = blast_columns, show_col_types = FALSE) %>% 
  filter(query %in% seeds) %>%
  mutate(evalue = ifelse(evalue==0, 1e-180, evalue))

blast_brhs <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_blast_brh.tsv"),
                     col_names = c("query", "target"), show_col_types = FALSE) %>% 
  filter(query %in% seeds) %>%
  mutate(method="blast_brh")
```

```{r merge_homo}
df <- full_join(blast, fs,
                by = c("query", "target"), suffix=c("_blast", "_fs")) %>%
    filter(query!=target) %>% 
  left_join(self, by="query") %>%
  mutate(pident_fs = 100*pident_fs, 
         qcov_fs = 100*qcov_fs,
         singleton = case_when(is.na(evalue_fs) ~ "only_blast",
                               is.na(evalue_blast) ~ "only_fs", 
                               .default = "common"),
         blast_norm = bitscore_blast/blast_selfbit, fs_norm = bitscore_fs/fs_selfbit)

brhs <- rbind(blast_brhs, fs_brhs) %>% 
    group_by(query, target, method) %>%
    mutate(n=1) %>%
    pivot_wider(names_from = method, values_fill = 0, values_from = n) %>%
    mutate(common = ifelse(fs_brh+blast_brh==2, 1, 0))

# grouped_brhs <- brhs %>%
#   group_by(query) %>%
#   summarise(fs_brh = sum(fs_brh), blast_brh=sum(blast_brh), common_brh=sum(common))
```
## Blast vs Foldseek

There are these number of hits, divided by singleton class:

```{r eda_homo}
df %>% count(singleton)
```

We can explore different properties of blast and foldseek search divided by singleton class.

<!-- ```{r blast_vs_fs, echo = FALSE, message=FALSE, fig.align='center', fig.cap='Some cool caption', out.width='\\linewidth', fig.pos='H'} -->
<!-- knitr::include_graphics(glue(outdir, "/plots/", params_dataset$seed, "_homology.pdf")) -->
<!-- ``` -->

<center>
![*Blast vs foldseek properties*](../../draft/figures/UP000005640_homology.png)
</center>

## BRHs

We can explore the abundance of BRH hits for every seed protein across the different species.

```{r brhs, fig.width=12}
brhs %>% 
  left_join(select(taxmap, -Proteome_ID), by=c("target"="label")) %>% 
  mutate(mnemo=factor(mnemo, levels=sptree$tip.label)) %>% 
  # pivot_longer(cols = c(blast_brh, fs_brh, common)) %>% 
  filter(mnemo!="Hsap") %>% 
  pivot_longer(cols = c(blast_brh, fs_brh, common)) %>% 
  filter(value!=0) %>% 
  ggplot(aes(mnemo, fill=name)) + 
  geom_bar(position = "dodge") + 
  scale_fill_manual(values = palette_brh) + 
  scale_y_continuous(expand = expansion(0,0))
# select(df, query, target, singleton) %>% 
#   left_join(brhs) %>% 
#   group_by(singleton) %>%
#   summarise(#prop_brh = sum(!is.na(common))/n()*100,
#             prop_blast_brh = sum(blast_brh, na.rm = T)/n()*100, 
#             prop_fs_brh = sum(fs_brh, na.rm = T)/n()*100, 
#             prop_common_brh = sum(common, na.rm = T)/n()*100) %>% 
#   pivot_longer(!singleton) %>% 
#   ggplot(aes(name, value, color=singleton)) + 
#   geom_point(size=3) + 
#   scale_color_manual(values=palette_singleton) + 
#   labs(x="Type of BRH", y="% of BRH")
```

## Singletons structure quality

```{r lddt_singletons, fig.width=10, fig.height=5}
lddt_df <- read_delim(paste0(params$structure_dir, "/", tax$Proteome_ID, "/", tax$Proteome_ID, "_mean_plddt.tsv"),
           col_names = c("target", "mean_lddt"), show_col_types = FALSE)

df %>% 
  select(target, singleton) %>% 
  mutate(target=gsub("AF-|-F1", "", target)) %>% 
  left_join(lddt_df, by="target") %>% 
  ggplot(aes(mean_lddt, fill=singleton)) + 
  geom_density(alpha=.5) +
  scale_y_continuous(expand = expansion(0.01,0)) +
  scale_x_continuous(expand = expansion(0.01,0)) +
  scale_fill_manual(values = palette_singleton) 
```

We can see the meand plddt for singleton classes and common are generally better imputed proteins than the two singletons. Expectedly fs suffers more than blast.

We can also see more in detail some blast and foldseek features and how they differ.

<center>
![*Foldseek and blast properties divided by singleton class*](../../draft/figures/UP000005640_singletons.png)
</center>

# Post filtering blast

```{r filter_homo}
aln_seeds <- paste0("AF-",readLines(glue(outdir, "/ids/",params_dataset$seed,"_aln.ids")), "-F1")
blast_filtered <- filter(blast, 
                         query %in% aln_seeds,
                         length/qlen*100>params$coverage, 
                         length/slen*100>params$coverage,
                         evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

fs_filtered <- filter(fs, 
                      query %in% aln_seeds,
                      qcov*100>params$coverage, 
                      tcov*100>params$coverage, 
                      evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()
```

We filter blast and foldseek by `r params$coverage` query and target coverage and by `r params$eval_both` min evalue for our `r length(seeds)`. Only `r length(aln_seeds)` will go in the phylogenetic pipeline as they have more than 4 common hits. 
This results in `r nrow(blast_filtered)` blast rows and `r nrow(fs_filtered)` fs rows.

## Originations

```{r originations}
blast_originations <- blast_filtered %>% 
  select(query, target) %>% 
  left_join(taxmap, by=c("target"="label")) %>% 
  group_by(query) %>% 
  filter(mnemo!="Hsap") %>% 
  left_join(species_dist_df, by="mnemo") %>% 
  mutate(mrca=factor(mrca, levels=unique(species_dist_df$mrca), ordered = TRUE)) %>% 
  summarise(mrca=max(mrca)) %>% 
  group_by(mrca) %>% 
  count() %>% 
  mutate(method="blast")

fs_originations <- fs_filtered %>% 
  select(query, target) %>% 
  left_join(taxmap, by=c("target"="label")) %>% 
  group_by(query) %>% 
  filter(mnemo!="Hsap") %>% 
  left_join(species_dist_df, by="mnemo") %>% 
  mutate(mrca=factor(mrca, levels=unique(species_dist_df$mrca), ordered = TRUE)) %>% 
  summarise(mrca=max(mrca)) %>% 
  group_by(mrca) %>% 
  count() %>% 
  mutate(method="fs")

originations <- rbind(blast_originations, fs_originations)

originations %>% 
  ggplot(aes(n, fct_rev(mrca), fill=method)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_manual(values = palettes_method) + 
  scale_x_continuous(expand = expansion(0,0)) + 
  labs(y="MRCA")
```

We can see where is the "taxonomic mrca" (the most distant species) for every seed. Interestingly, foldssek is more than blast just in the two last nodes.

## Duplications

```{r duplications}
duplications_mrca <- read_delim(glue(outdir, "/reco/", params_dataset$seed,"_mrca.tsv"),
                                show_col_types = FALSE) %>% 
  separate(id, c("gene", "target", "alphabet", "model")) %>% 
  select(-alphabet) %>% 
  group_by(target, model, mrca) %>% 
  summarise(n=sum(n)) %>% 
  filter(model %in% c("LG", "FTPY")) %>% 
  mutate(mrca=factor(mrca, levels=fortify(sptree) %>% arrange(y) %>% pull(label), ordered = TRUE))

duplications_mrca %>% 
  filter(target %in% c("blast", "fs")) %>%
  ggplot(aes(n, mrca, fill=target)) + 
  geom_bar(position = "dodge", stat="identity") +
  facet_grid(~model) + 
  scale_x_continuous(expand = expansion(0,0)) +
  scale_fill_manual(values=palettes_method)

# fortify(sptree) %>% 
#   inner_join(duplications_mrca, by=c("label"="mrca")) %>% 
#   ggtree() + 
#   geom_point(aes(size=n)) + 
#   facet_grid(target~model)
```

We can also explore where the duplications called by RangerDTL are mapped in the species tree. For this we are only interested in *LG* and *FTPY* to reduce complexity.

# Trees

```{r read_trees}
df_trees <- read_delim(glue(outdir, "/trees/", params_dataset$seed, "_unrooted_trees.txt"),
                       show_col_types = FALSE,
                       col_names = c("gene", "target", "alphabet", "model", "tree")) %>% 
  mutate(model=factor(model, levels=models))
```

## Reconciliation

<center>
![*all 6 model results*](../../draft/figures/UP000005640_discordance.png)
</center>

<center>
![*Astral-PRO results*](../../draft/figures/UP000005640_astral_pro.png)
</center>

<center>
![*Distance 2 seed by singleton class*](../../draft/figures/UP000005640_distance.png)
</center>

## RF distance + varr2t

<center>
![*Various Tree analysis*](../../draft/figures/UP000005640_trees.png)
</center>

## Model fit

How many times is each model better?

```{r prop_better, fig.width=10}
reco <- read_delim(glue(outdir, "/reco/", params_dataset$seed, "_DTL.tsv"), show_col_types = FALSE) %>%
  mutate(model=factor(model, levels=models))

scores <- read_delim(glue(outdir, "/reco/", params_dataset$seed, "_scores.tsv"), show_col_types = FALSE) %>%
  left_join(reco) %>%
  mutate(model=factor(model, levels=models))

reco_prop_plot <- scores %>%
  group_by(id, targets) %>%
  mutate(norm_reco_score=(dups+losses)/n_tips) %>%
  slice_min(norm_reco_score, with_ties = TRUE) %>%
  mutate(best_score=1/n()) %>% 
  ggplot(aes(targets, best_score, fill=model)) +
  geom_bar(stat = "identity") +
  labs(subtitle = "How many times each model has the lowest D+L/tips?") +
  scale_fill_manual(values = palettes_model)

tcs_prop_plot <- scores %>%
  group_by(id, targets) %>%
  slice_max(score, with_ties = TRUE) %>%
  mutate(best_score=1/n()) %>% 
  ggplot(aes(targets, best_score, fill=model)) +
  geom_bar(stat = "identity") +
  labs(subtitle = "How many times each model has the highest TCS score?") +
  scale_fill_manual(values = palettes_model)

reco_prop_plot + tcs_prop_plot +  plot_layout(guides = 'collect')
```

## Distance between target classes

We can see the average normalized topological distance between the different groups in LG and FTPY.

```{r targets_topodist, fig.width=10, fig.height=5}
df_trees_dist <- filter(df_trees, model %in% c("LG", "FTPY"), target=="union")
ts_dist <- read.tree(text = df_trees_dist$tree)
names(ts_dist) <- paste(df_trees_dist$gene, df_trees_dist$target, df_trees_dist$alphabet, df_trees_dist$model, sep = "_")

df_distance <- tibble()

for (idx in 1:length(ts_dist)) {
    seed <- paste0("AF-",df_trees_dist$gene[idx],"-F1")
    tmp_df <- select(filter(df, query==seed), target, singleton)

    tree_dist <- as.matrix(adephylo::distTips(ts_dist[[idx]], method = "nNodes")) %>% 
      as.data.frame() %>% 
      rownames_to_column() %>% 
      pivot_longer(!rowname) %>% 
      filter(rowname>name) %>%
      mutate(value=value/max(value)) %>% 
      left_join(tmp_df, by = c("rowname"="target")) %>% 
      left_join(tmp_df, by = c("name"="target")) %>% 
      mutate(a=pmin(singleton.x, singleton.y),
             b=pmax(singleton.x, singleton.y)) %>% 
      group_by(a, b) %>%
      filter(!is.na(a), !is.na(b)) %>%
      summarise(mean_dist=mean(value), n=n(), .groups = "keep") %>%
      mutate(gene=seed, model=df_trees_dist$model[idx])

  df_distance <- bind_rows(df_distance, tree_dist)
}

df_distance %>% 
  group_by(a,b,model) %>% 
  summarise(wmean=weighted.mean(mean_dist,n)) %>% 
  ggplot(aes(a,b, fill=wmean)) + 
  geom_tile() +
  geom_text(aes(label=round(wmean,2))) +
  facet_grid(~model) +
  scale_fill_gradientn(colors=wespal) +
  coord_cartesian(expand = F) +
  theme(axis.title = element_blank(),
        legend.position = "none")
```

# CATH/SCOPe

# Runtimes

<center>
![*Various runtimes statistics*](../../draft/figures/UP000005640_runtime.png)
</center>

# TODOs

* CATH/SCOPe
* Benchmark: fix runtimes!
* Fix FTPY branch support
* Foldtree smk vs Foldtree python: DONE
* duplication in species tree: DONE

### 2024-05-28 workflow/notebooks/explore_examples.Rmd
added notebooks, FTPY and some code improvements

---
title: "Structural vs Sequence Dashboard"
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    theme:
      bg: "#FFFFFF"
      fg: "#00a276"
      primary: "#00a276"
      base_font:
        google: Roboto
      code_font:
        google: JetBrains Mono
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
# Install thematic and un-comment for themed static plots (i.e., ggplot2)
# thematic::thematic_rmd()
```

## Column {.tabset data-width="650"}

```{r libraries, message=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(ggtree))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggmsa))
library(patchwork)
library(shiny)

source("../scripts/functions.R")
theme_set(theme_classic())
```

```{r input_files}
models <- c("QT", "FT", "LG", "GTR", "3Di")

ids_input <- "../../results/test/ids/UP000005640_examples.ids"
trees <- "../../results/test/trees/UP000005640_unrooted_trees.txt"
taxidmap <- "../../results/test/db/taxidmap"
meta <- "../../data/meta/Hsap_taxon.csv"

seeds <- readLines(ids_input)
seed_sp <- gsub("_examples.ids", "", basename(ids_input))
seed_dir <- gsub("trees", "seeds", dirname(trees))

taxmap <- read_delim(taxidmap, 
                     delim = "\t", col_names = c("label", "Tax_ID")) %>% 
  left_join(read_delim(meta)) %>% 
  dplyr::select(label, Proteome_ID, mnemo)

fls <- list.files(paste0("~/mn5/projects/projects/structural_phylome/results/test/seeds/", seed_sp, "/", seeds),
                  full.names = TRUE)
ids <- fls[grepl("(blast|fs).*ids$", fls)]
x <- sapply(ids, readLines, simplify = FALSE)
names(x) <- gsub("\\.ids", "", basename(ids))
alns <- fls[grepl("commo.*clean$", fls)]
```

```{r}
selectInput("seed", label = "seed", choices = seeds)
```

```{r}
venn_input <- reactive({
  x[which(grepl(input$seed, names(x)))]
})

aln_3di_input <- reactive({
  Biostrings::readAAMultipleAlignment(filepath = alns[grepl(paste0(input$seed,".*3Di.alg"), alns)])
})

aln_aa_input <- reactive({
  Biostrings::readAAMultipleAlignment(filepath = alns[grepl(paste0(input$seed,".*aa.alg"), alns)])
})

# alns <- fls[grepl("commo.*clean$", fls)]
#   
# # 3Di aln
# aln_3di <- Biostrings::readAAMultipleAlignment(alns[grepl("3Di.alg", alns)])
# aln_3di_mat <- as.matrix(aln_3di@unmasked)
# pos_df_3di <- tibble(pos=1:ncol(aln_3di_mat), 
#                      cons=nrow(aln_3di_mat) - colSums(aln_3di_mat=="-" | aln_3di_mat=="X"))
# 
# aln_aa <- Biostrings::readAAMultipleAlignment(alns[grepl("aa.alg", alns)])
# aln_aa_mat <- as.matrix(aln_aa@unmasked)
# pos_df_aa <- tibble(pos=1:ncol(aln_aa_mat), 
#                      cons=nrow(aln_aa_mat) - colSums(aln_aa_mat=="-" | aln_aa_mat=="X"))
# 
# trees_files <- fls[grepl("nwk$", fls)]
# 
# trees <- read.tree(text = sapply(trees_files, readLines))
# names(trees) <- gsub(paste0(seed,"_|(3Di|aa)_|.nwk$"), "", basename(trees_files))

# conf_df <- tibble(label=union(x$blast, x$fs)) %>% 
#   left_join(taxmap) %>% 
#   mutate(confidence_file=paste0("results/data/structures/",
#                                 Proteome_ID,
#                                 "/confidence/",
#                                 label,
#                                 "-confidence_v4.json.gz")) %>% 
#   rowwise() %>% 
#   mutate(conf=mean(jsonlite::fromJSON(gzfile(confidence_file))$confidenceScore))
```

### Venn

```{r}
renderPlot(
  venn_input() %>% 
    ggvenn::ggvenn(auto_scale = TRUE)
)
```

### Alns 3di

```{r aln 3di}

renderPlot(
  aln_3di_input() %>% 
  ggmsa(seq_name = FALSE, font = NULL, show.legend = TRUE, border = NA) +
    labs(title = paste(input$seed, "- 3Di common hits cleaned aln")) +
    # geom_line(mapping = aes(x=pos, y=cons), data = pos_df_3di, inherit.aes = F) +
    coord_cartesian() + 
    theme_void() + 
    theme(axis.text.x = element_text())
)

```

### Aln AA

```{r aln_aa}

renderPlot(
  aln_aa_input() %>% 
  ggmsa(seq_name = FALSE, font = NULL, show.legend = TRUE, border = NA) +
    labs(title = paste(input$seed, "- 3Di common hits cleaned aln")) +
    # geom_line(mapping = aes(x=pos, y=cons), data = pos_df_3di, inherit.aes = F) +
    coord_cartesian() + 
    theme_void() + 
    theme(axis.text.x = element_text())
)

```

### 2024-05-28 software/foldtree/foldseek2tree.py

@@ -189,7 +189,7 @@ def postprocess(t, outree, delta=0 ):
 	print(tre)
 
 	for n in tre.treenode.traverse():
-		if n.dist< 0:
+		if n.dist < delta:
 			n.dist = delta
 	tre.write( outree, tree_format = 0 )
 

### 2024-05-28 software/foldtree/foldseekres2distmat_simple.py

@@ -1,6 +1,30 @@
-import foldseek2tree
 import numpy as np
 import pandas as pd
+
+def distmat_to_txt(identifiers, distmat, outfile):
+	'''
+	write out a distance matrix in fastme format
+
+	Parameters
+	----------
+	identifiers : list
+		list of identifiers for your proteins
+	distmat : np.array  
+		distance matrix
+	outfile : str   
+		path to output file
+
+	'''
+
+	#write out distmat in phylip compatible format
+	outstr = str(len(identifiers)) + '\n'
+	for i,pdb in enumerate(identifiers):
+		outstr += pdb + ' ' + ' '.join( [ "{:.4f}".format(d) for d in list( distmat[i,: ] )  ]  ) + '\n'
+	with open(outfile , 'w') as handle:
+		handle.write(outstr)
+		handle.close()
+	return outfile
+
 
 res = pd.read_table(snakemake.input[0], header = None)
 # print(res.head())
@@ -33,5 +57,5 @@ for i,k in enumerate(matrices):
     matrices[k] = 1-matrices[k]
     # print(matrices[k], np.amax(matrices[k]), np.amin(matrices[k]) )
     # np.save( infolder + k + '_distmat.npy' , matrices[k])
-    distmat_txt = foldseek2tree.distmat_to_txt(ids, matrices[k], snakemake.output[i])
+    distmat_txt = distmat_to_txt(ids, matrices[k], snakemake.output[i])
 

### 2024-05-28 workflow/scripts/analyze_apro.R

@@ -26,6 +26,7 @@ nodes_df <- trees_df %>%
   left_join(nodes_clades) %>% 
   mutate(freq=f1+f2+f3)
 
+
 pp1_plot <- nodes_df %>% 
     ggplot(aes(q1, label, fill=model, color=model, group=model)) +
     geom_linerange(aes(x=x, xmin=xmin, xmax=xmax, color=model),
@@ -38,10 +39,10 @@ pp1_plot <- nodes_df %>%
                position = position_dodge(width=.7)) +
   scale_fill_manual(values=palettes_model) +
   scale_color_manual(values = palettes_model_d) +
-  scale_shape_manual(values = c(21,22,23)) +
+  scale_shape_manual(values = c(21,22,23,24)) +
   guides(color = guide_legend(override.aes = list(pch = 21, size=3)),
          shape = guide_legend(override.aes = list(size=3))) +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom", legend.box="vertical")
 
 sptree_plot <- ggtree(labeled_sptree) + 
   geom_tiplab() + 

### 2024-05-28 workflow/scripts/analyze_bl.R

@@ -1,0 +1,78 @@
+suppressPackageStartupMessages(library(ggtree))
+suppressPackageStartupMessages(library(tidyverse))
+source("workflow/scripts/functions.R")
+
+theme_set(theme_classic())
+
+# all trees
+df_trees <- read_delim(snakemake@input[["trees"]], show_col_types = FALSE,
+                       col_names = c("gene", "targets", "alphabet", "model", "tree")) %>% 
+  mutate(model=factor(model, levels=models)) %>% 
+  filter(targets=="union")
+
+# Read all trees
+ts <- read.tree(text = df_trees$tree)
+names(ts) <- paste(df_trees$gene, df_trees$targets, df_trees$alphabet, df_trees$model, sep = "_")
+
+fs <- read_delim(snakemake@input[["fs"]], 
+                 col_names = fs_columns, show_col_types = FALSE) %>% 
+  filter(query %in% paste0("AF-", df_trees$gene, "-F1")) 
+  # filter(evalue<snakemake@params[["eval_both"]], 
+  #        qcov*100>snakemake@params[["coverage"]], 
+  #        tcov*100>snakemake@params[["coverage"]])
+
+blast <- read_delim(snakemake@input[["blast"]], 
+                    col_names = blast_columns, show_col_types = FALSE) %>% 
+  filter(query %in% paste0("AF-", df_trees$gene, "-F1")) #%>% 
+  # filter(evalue<snakemake@params[["eval_both"]], 
+  #        length/qlen*100>snakemake@params[["coverage"]], 
+  #        length/slen*100>snakemake@params[["coverage"]])
+
+
+df <- full_join(blast, fs,
+                by = c("query", "target"), suffix=c("_blast", "_fs")) %>%
+  mutate(pident_fs = 100*pident_fs, 
+         qcov_fs = 100*qcov_fs,
+         singleton = case_when(is.na(evalue_fs) ~ "only_blast",
+                               is.na(evalue_blast) ~ "only_fs", 
+                               .default = "common"))
+
+df_red <- select(df, query, target, singleton) 
+
+res <- NULL
+for (idx in 1:nrow(df_trees)){
+  t <- ts[[idx]]
+  seed <-  t$tip.label[grepl(df_trees$gene[idx], t$tip.label)]
+  # cophe <- cophenetic(t)
+  # total <- sum(cophe)
+  # bl <- cophe[seed,]
[…]

### 2024-05-28 workflow/scripts/compare_sampling.R

@@ -48,25 +48,25 @@ evals <- c(Inf, 1e-2, 1e-3, 1e-5)
 min_covs <- c(0, 30, 80)
 
 
-fs_brhs <- read_delim(snakemake@input[["fs_brh"]], col_names = c("query", "target"), show_col_types = FALSE) %>% 
-  mutate(method="blast_brh")
-  # group_by(query) %>%
-  # summarise(fs_brhs=list(target))
+# fs_brhs <- read_delim(snakemake@input[["fs_brh"]], col_names = c("query", "target"), show_col_types = FALSE) %>% 
+#   mutate(method="blast_brh")
+#   # group_by(query) %>%
+#   # summarise(fs_brhs=list(target))
 
-blast_brhs <- read_delim(snakemake@input[["blast_brh"]], col_names = c("query", "target"), show_col_types = FALSE) %>% 
-  mutate(method="fs_brh")
-  # group_by(query) %>%
-  # summarise(blast_brhs=list(target))
+# blast_brhs <- read_delim(snakemake@input[["blast_brh"]], col_names = c("query", "target"), show_col_types = FALSE) %>% 
+#   mutate(method="fs_brh")
+#   # group_by(query) %>%
+#   # summarise(blast_brhs=list(target))
 
-brhs <- rbind(blast_brhs, fs_brhs) %>% 
-  group_by(query, target, method) %>% 
-  mutate(n=1) %>% 
-  pivot_wider(names_from = method, values_fill = 0, values_from = n) %>% 
-  mutate(common = ifelse(fs_brh+blast_brh==2, 1, 0))
+# brhs <- rbind(blast_brhs, fs_brhs) %>% 
+#   group_by(query, target, method) %>% 
+#   mutate(n=1) %>% 
+#   pivot_wider(names_from = method, values_fill = 0, values_from = n) %>% 
+#   mutate(common = ifelse(fs_brh+blast_brh==2, 1, 0))
 
-grouped_brhs <- brhs %>%
-  group_by(query) %>% 
-  summarise(fs_brh = sum(fs_brh), blast_brh=sum(blast_brh), common_brh=sum(common))
+# grouped_brhs <- brhs %>%
+#   group_by(query) %>% 
+#   summarise(fs_brh = sum(fs_brh), blast_brh=sum(blast_brh), common_brh=sum(common))
 
 df <- full_join(select(blast, query, target, pident, evalue, bitscore, qcov),
[…]

### 2024-05-28 workflow/scripts/compare_sampling_trees.R

@@ -35,7 +35,7 @@ fs <- read_delim(snakemake@input[["fs"]],
          qcov*100>snakemake@params[["coverage"]], 
          tcov*100>snakemake@params[["coverage"]]) %>% 
   group_by(query) %>% 
-  mutate(is_top = row_number()<=snakemake@params[["max_seqs"]]) %>% 
+  filter(row_number()<=snakemake@params[["max_seqs"]]) %>% 
   ungroup()
 
 blast <- read_delim(snakemake@input[["blast"]], 
@@ -46,7 +46,7 @@ blast <- read_delim(snakemake@input[["blast"]],
          length/qlen*100>snakemake@params[["coverage"]], 
          length/slen*100>snakemake@params[["coverage"]]) %>% 
   group_by(query) %>% 
-  mutate(is_top = row_number()<=snakemake@params[["max_seqs"]]) %>% 
+  filter(row_number()<=snakemake@params[["max_seqs"]]) %>% 
   ungroup()
 # $11<{params.eval_both} && $4/$15*100>{params.coverage} && $4/$16*100
 
@@ -58,19 +58,6 @@ df <- full_join(blast, fs,
          singleton = case_when(is.na(evalue_fs) ~ "only_blast",
                                is.na(evalue_blast) ~ "only_fs", 
                                .default = "common"))
-
-plot_top <- df %>% 
-  select(query, singleton, contains("is_top")) %>% 
-  group_by(singleton, is_top_blast, is_top_fs) %>% 
-  count() %>% 
-  filter(singleton=="common") %>%  
-  ggplot(aes(is_top_blast, is_top_fs, fill=n)) + 
-  geom_tile() + 
-  geom_text(aes(label=n), color="white") +
-  coord_cartesian(expand = F) + 
-  scale_fill_viridis_c(option = "E") +
-  labs("Common hits in top 150") +
-  theme(legend.position = "none")
 
 # first of all let's explore common properties:
 # bitscore, evalue, pident, prop mismatch and prop gaps
@@ -95,7 +82,7 @@ plot_common <- df %>%
   scale_x_continuous(expand = c(0,0)) +
   scale_linetype_manual(values = c(2,1)) + 
   labs(subtitle = "Common properties distribution") + 
-  theme(legend.position = "none")
+  theme(legend.position = "bottom")
 
 
 plot_fs <- df %>% 
@@ -108,34 +95,34 @@ plot_fs <- df %>%
[…]

### 2024-05-28 workflow/scripts/compare_trees.R

@@ -6,9 +6,6 @@ source("workflow/scripts/functions.R")
 
 theme_set(theme_classic())
 
-# order of models
-models <- c("QT", "FT", "LG", "GTR", "3Di")
-
 # all trees
 df_trees <- read_delim(snakemake@input[["trees"]], show_col_types = FALSE,
                        col_names = c("gene", "target", "alphabet", "model", "tree")) %>% 
@@ -73,17 +70,17 @@ plot_3di <- filter(df_ml, model!="LG") %>%
   # geom_density() + 
   ggdist::stat_slab(aes(fill = after_stat(level)), .width = c(.5, .66, .95, 1)) +
   scale_fill_brewer() +
-  geom_text(data = . %>%
-              mutate(x=quantile(diff_BIC, .1)) %>% 
-              group_by(targets, better_3Di, x) %>% 
-              summarise(n=n()) %>% 
-              pivot_wider(names_from = better_3Di, values_from = n) %>% 
-              mutate(prop_better=`TRUE`/(`TRUE`+`FALSE`)), 
-            mapping = aes(x=x, y=0.5, 
-                          label=paste0(round(prop_better*100, 2),"%"))) +
+  # geom_text(data = . %>%
+  #             mutate(x=quantile(diff_BIC, .1)) %>% 
+  #             group_by(targets, better_3Di, x) %>% 
+  #             summarise(n=n()) %>% 
+  #             pivot_wider(names_from = better_3Di, values_from = n) %>% 
+  #             mutate(prop_better=`TRUE`/(`TRUE`+`FALSE`)), 
+  #           mapping = aes(x=x, y=0.5, 
+  #                         label=paste0(round(prop_better*100, 2),"%"))) +
   geom_vline(xintercept = 0) + 
   coord_cartesian(expand = 0) + 
-  facet_wrap(~targets, nrow = 1) +
+  facet_grid(~targets, scales="free_x") +
   labs(x="BIC 3Di - BIC GTR", y="Density", subtitle = "GTR vs 3Di BIC") + 
   theme(legend.position = "bottom")
 
@@ -167,7 +164,7 @@ plot_tcs <- inner_join(TCS_df, TCS_df, by = c("id", "targets"),
 
 
 plot_varr2t <- scores %>% 
-  filter(model %in% c("3Di", "LG", "GTR")) %>% 
+  # filter(model %in% c("3Di", "LG", "GTR")) %>% 
[…]

### 2024-05-28 workflow/scripts/functions.R

@@ -1,3 +1,7 @@
+interesting_types <- c("Topological domain", "Domain", 
+                       "Region", "Helix", "Zinc finger",
+                       "Beta strand", "Turn", "Coiled coil", "Transmembrane")
+
 blast_columns <- c("query","target","pident","length","mismatch","gapopen",
                    "qstart","qend","sstart","send","evalue","bitscore",
                    "qcov","qcovhsp","qlen","slen","staxids")
@@ -6,16 +10,21 @@ fs_columns <- c("query","target","pident","length","mismatch","gapopen",
                 "qstart","qend","sstart","send","evalue","bitscore",
                 "lddt","alntmscore","rmsd","prob","qcov","tcov")
 
-palettes_model <- c("#BF360C", "#75C323", "#40A79D", "#FFA000", "#D4E157")
-names(palettes_model) <- c("LG", "3Di", "GTR", "QT", "FT")
+models <- c("LG", "3Di", "GTR", "QT", "FT", "FTPY")
+
+palettes_model <- c("#BF360C", "#75C323", "#40A79D", "#FFA000", "#D4E157","#646B00")
+names(palettes_model) <- models
 palettes_model_d <- colorspace::darken(palettes_model, amount = 0.6)
-names(palettes_model_d) <- c("LG", "3Di", "GTR", "QT", "FT")
+names(palettes_model_d) <- models
 
-palettes_method <- c("#F23030", "#F29F05", "#267365")
-names(palettes_method) <- c("common", "blast", "fs")
+palettes_method <- c("#F23030", "#F29F05", "#267365", "#D6FAB7")
+names(palettes_method) <- c("common", "blast", "fs", "union")
 
-palette_singleton <- c("#F23030", "#F29F05", "#267365")
-names(palette_singleton) <- c("common", "only_blast", "only_fs")
+palette_singleton <- c("#F23030", "#F29F05", "#267365", "#D6FAB7")
+names(palette_singleton) <- c("common", "only_blast", "only_fs", "union")
+
+palette_brh <- c("#F23030","#F29F05", "#267365")
+names(palette_brh) <- c("common","blast_brh", "fs_brh")
 
 palettes_alphabet <- c("#327AD9", "#BF3475", "#F29F05")
 names(palettes_alphabet) <- c("3Di", "aa", "fident")
[…]

### 2024-05-28 workflow/scripts/plot_example.R

@@ -1,0 +1,208 @@
+suppressPackageStartupMessages(library(ggtree))
+suppressPackageStartupMessages(library(tidyverse))
+suppressPackageStartupMessages(library(ggmsa))
+library(patchwork)
+
+source("workflow/scripts/functions.R")
+theme_set(theme_classic())
+
+seeds <- readLines(snakemake@input[["ids"]])
+
+seed_sp <- gsub("_examples.ids", "", basename(snakemake@input[["ids"]]))
+seed_dir <- gsub("trees", "seeds", dirname(snakemake@input[["trees"]]))
+
+taxmap <- read_delim(snakemake@input[["taxidmap"]], 
+                     delim = "\t", col_names = c("label", "Tax_ID"),
+                     show_col_types = FALSE) %>% 
+  left_join(read_delim(snakemake@input[["meta"]],
+                       show_col_types = FALSE), by = "Tax_ID") %>% 
+  dplyr::select(label, Proteome_ID, mnemo)
+
+
+sptree <- read.tree(snakemake@input[["sptree"]])
+translate <- deframe(distinct(dplyr::select(taxmap, Proteome_ID, mnemo)))
+sptree$tip.label <- names(translate)[match(translate, sptree$tip.label)]
+
+species_dist_df <- as.matrix(adephylo::distTips(sptree, method = "nNodes"))[snakemake@params[["seed"]], ] %>% 
+  enframe(name = "Proteome_ID", value = "dist")
+
+gff <- rtracklayer::readGFF(snakemake@input[["gff"]]) %>%
+  as_tibble() %>% 
+  filter(seqid %in% c(seeds), 
+         type %in% c(interesting_types, "Chain"))
+
+reco <- read_delim(snakemake@input[["reco"]], show_col_types = FALSE) %>% 
+  mutate(model=factor(model, levels=models))
+
+plot_list = list()
+
+for (seed in seeds) {
+  print(seed)
+  fls <- list.files(paste0("results/test/seeds/", seed_sp, "/", seed), full.names = TRUE)
+  ids <- fls[grepl("(blast|fs).*ids$", fls)]
+
+  x <- sapply(ids, readLines, simplify = FALSE)
+  names(x) <- gsub(".*_|\\.ids", "", basename(ids))
+  
+  singleton_df <- list("common" = intersect(x$fs, x$blast), 
+                 "only_fs" = setdiff(x$fs, x$blast), 
[…]

### 2024-05-28 workflow/scripts/temp_BRH.R

@@ -1,0 +1,10 @@
+suppressPackageStartupMessages(library(ggtree))
+suppressPackageStartupMessages(library(tidyverse))
+source("workflow/scripts/functions.R")
+
+theme_set(theme_classic())
+
+# order of models
+models <- c("QT", "FT", "LG", "GTR", "3Di")
+
+

### 2024-06-04 workflow/notebooks/CATH.Rmd
Added CATH benchmark

---
title: "CATH analysis"
# knit: (function(input_file, encoding) {
#   rmarkdown::render(input_file,
date: "`r format(Sys.time(), '%d-%b-%Y')`"
author: "Giacomo Mutti"
output:
    rmdformats::html_clean:
        number_sections: false
        code_folding: hide
        self_contained: true
        fig_caption: true
        # thumbnails: false
        # lightbox: false
        gallery: true
        toc_depth: 3
        highlight: kate
---

```{r setup, include=FALSE, dev="CairoPNG"}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, cache.lazy = FALSE,
                      crop = TRUE, dev = "CairoPNG")
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
knitr::opts_knit$set(root.dir = "../../")
```

```{r libraries, message=FALSE, warning=FALSE, cache=FALSE}
library(tidyverse)
library(ggraph)
library(tidygraph)
library(patchwork)
library(ggtree)
library(cowplot)
library(glue)
source("workflow/scripts/functions.R")
theme_set(theme_classic())

models <- names(palettes_model)

params <- yaml::read_yaml("config/params.yaml")
params_dataset <- yaml::read_yaml("config/Hsapopi_test.yaml")

outdir <- glue(params$outdir, params_dataset$dataset)
```

```{r read_dfs}
CATH_df <- read_delim(list.files(params$cath_dir, full.names = TRUE),
                      delim = "\t", id = "proteome", show_col_types = FALSE) %>%
  # filter(!is.na(Gene3D)) %>% 
  mutate(proteome = gsub("_cath.tsv", "", basename(proteome)),
         n_3d_domains = str_count(Gene3D, ";"),
         n_Pfam_domains = str_count(Pfam, ";")) %>%
  # filter(n_domains==1) %>% 
  mutate(target=paste0("AF-", Entry, "-F1")) %>% 
    select(-proteome, Entry)

nms <- read_delim("test/cath/cath-names.txt", 
                  comment = "#", delim = "    ",
                  col_names = c("Gene3D", "abbr", "name"), 
                  show_col_types = FALSE) %>% 
  select(-abbr) %>% 
  mutate(name=gsub("^:", "", name))

nms_pfam <- read_delim("test/cath/Pfam-A.clans.tsv", 
           col_names = c("Pfam", "clan", "boh", "abbr", "name"), 
           show_col_types = FALSE) %>% 
  select(-name, -clan, -boh)

long_CATH_df <- CATH_df %>% 
  # filter(target %in% unique(blast$query)) %>% 
  mutate(Gene3D=strsplit(Gene3D, ";")) %>% 
  unnest(Gene3D) %>% 
  separate(col = Gene3D, into = c("C", "A", "T", "H"), sep = "\\.", remove = FALSE) %>% 
  mutate(CA = paste(C,A,sep = "."), CAT=paste(C,A,T,sep = ".")) %>% 
  left_join(nms, by=c("C"="Gene3D")) %>%  
  left_join(nms, by=c("CA"="Gene3D"), suffix = c("_C", "_CA")) %>% 
  left_join(nms, by=c("CAT"="Gene3D"), suffix = c("_CA", "_CAT")) %>% 
  left_join(nms, by=c("Gene3D"), suffix = c("_CAT", "_CATH"))

long_CATH_nms <- long_CATH_df %>%
    select(-c(Entry, Pfam, n_Pfam_domains, target, n_3d_domains)) %>% 
    distinct()
```

```{r read_blast}
seeds <- paste0("AF-",readLines(glue(outdir, "/ids/",params_dataset$seed,"_aln.ids")), "-F1")

fs <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_fs.tsv"), 
                 col_names = fs_columns, show_col_types = FALSE) %>% 
  # filter(query %in% seeds) %>%
  mutate(evalue = ifelse(evalue==0, 1e-180, evalue)) %>% 
  filter(qcov*100>params$coverage,
         tcov*100>params$coverage,
         evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

blast <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_blast.tsv"), 
                    col_names = blast_columns, show_col_types = FALSE) %>% 
  # filter(query %in% seeds) %>%
  mutate(evalue = ifelse(evalue==0, 1e-180, evalue)) %>% 
  filter(length/qlen*100>params$coverage, 
         length/slen*100>params$coverage,
         evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

df <- full_join(mutate(select(fs, query, target), method="fs"),
                mutate(select(blast, query, target), method="blast"),
          by = c("query", "target"), suffix=c("_fs", "_blast")) %>% 
    filter(query!=target) %>% 
    mutate(singleton = case_when(is.na(method_fs) ~ "only_blast",
                                 is.na(method_blast) ~ "only_fs", 
                                 TRUE ~ "common")) %>% 
  select(-contains("method"))
```

# CATH exploration

First of all we can see more or less how many domains (3d and Pfam) there are in all species?

By analyzing only monodomains we simplify everything and we still consider the majority of the data.

```{r monodomain, fig.width=6, fig.height=5}
CATH_df %>% 
  pivot_longer(c(n_3d_domains, n_Pfam_domains), values_to = "N_domains") %>% 
  mutate(name=str_split_i(name, "_", 2)) %>% 
  ggplot(aes(N_domains, fill=name)) + 
  geom_histogram(binwidth = 1, position = "dodge", color="black") +
  scale_x_continuous(breaks = seq(1,100,1)) +
  scale_y_continuous(expand = expansion(0,0)) + 
  scale_fill_manual(values=palette_domains)
```

From now on we will analyze the results based on the top 150 filtered hits for each query (not only the test seeds).
We can check how many distinct CATHs there are in targets in monodomain queries. Here it shows that foldseek recovers more weird stuff (analogous etc).

```{r monodomain_distinct, fig.width=12, fig.height=8}
plot_CATH <- inner_join(df, CATH_df, by = "target") %>%
  filter(n_3d_domains==1) %>%
  group_by(query, singleton) %>% 
  summarise(n_distinct_CATH = n_distinct(Gene3D)) %>% 
  ggplot(aes(n_distinct_CATH, fill=singleton)) + 
  geom_histogram(binwidth = 1, position = "dodge") +
  # facet_grid(n_domains~., scales = "free_y") +
  scale_x_continuous(breaks = seq(1,100,1)) + 
  scale_y_continuous(expand = expansion(0,0)) + 
  scale_fill_manual(values=palette_singleton)

plot_Pfam <- inner_join(df, CATH_df, by = "target") %>%
  filter(n_Pfam_domains==1) %>% 
  group_by(query, singleton) %>% 
  summarise(n_distinct_Pfam = n_distinct(Pfam)) %>% 
  ggplot(aes(n_distinct_Pfam, fill=singleton)) + 
  geom_histogram(binwidth = 1, position = "dodge") +
  # facet_grid(n_domains~., scales = "free_y") +
  scale_x_continuous(breaks = seq(1,100,1)) + 
  scale_y_continuous(expand = expansion(0,0)) + 
  scale_fill_manual(values=palette_singleton)

(plot_CATH / plot_Pfam) + plot_layout(guides = 'collect')
```

We can also see when the Pfam and CATH disagree in monodomain proteins, this could be explained by the fact the structure is more conserved, therefore less FALSEs where sequence is more prone to this.

```{r combn, fig.width=4, fig.height=8}
df %>% 
  select(-contains("method")) %>%
  left_join(CATH_df, by=c("query"="target")) %>% 
  left_join(CATH_df, by=c("target"), suffix = c("_query", "_target")) %>% 
  filter(query!=target, n_Pfam_domains_query==1, n_3d_domains_query==1,
          n_Pfam_domains_target==1, n_3d_domains_target==1) %>% 
  mutate(is_same_3d=Gene3D_query==Gene3D_target, is_same_Pfam=Pfam_query==Pfam_target) %>% 
  group_by(singleton) %>%
  count(is_same_3d, is_same_Pfam) %>% 
  filter(!is.na(is_same_3d), !is.na(is_same_Pfam)) %>%
  ggplot(aes(is_same_3d,is_same_Pfam,fill=n)) +
  geom_tile(color="black") +
  geom_text(aes(label=n), color="white") +
  facet_grid(singleton~.) + 
  scale_fill_gradientn(colours = wespal) + 
  coord_cartesian(expand = 0) +
  theme(legend.position = "none")
```

If we want to include multidomain proteins we can see the % of overlap between Pfam and 3D domains.

```{r multi, fig.width=10, fig.height=8}
df_multi <- df %>% 
    # head(n = 100000) %>%
  left_join(CATH_df, by=c("query"="target")) %>% 
  filter(!is.na(Pfam), !is.na(Gene3D)) %>%
  left_join(CATH_df, by="target") %>% 
  rowwise() %>% 
  mutate(Gene3D.x=strsplit(Gene3D.x, ";"),
         Pfam.x=strsplit(Pfam.x, ";"),
         Gene3D.y=strsplit(Gene3D.y, ";"),
         Pfam.y=strsplit(Pfam.y, ";")) %>% 
  mutate(n_3d_common=length(intersect(Gene3D.x,Gene3D.y)),
         n_Pfam_common=length(intersect(Pfam.x,Pfam.y))) %>% 
  # filter(n_Pfam_common>0, n_3d_common>0) %>% 
  select(-c(Pfam.x,Pfam.y,Gene3D.x,Gene3D.y)) %>% 
  mutate(overlap_3d=n_3d_common/pmax(n_3d_domains.x,n_3d_domains.y),
         overlap_Pfam=n_Pfam_common/pmax(n_Pfam_domains.x,n_Pfam_domains.y)) %>% 
  select(query, target, singleton, contains("overlap"), contains("common"))

df_multi %>% 
    pivot_longer(contains("overlap")) %>% 
    ggplot( aes(value, fill=singleton)) +
    geom_histogram(binwidth = 0.1, position = "dodge") +
    facet_grid(name~.) +
    scale_x_continuous(breaks = seq(0,1,0.1)) +
    scale_y_continuous(expand = expansion(0,0)) + 
    scale_fill_manual(values=palette_singleton)
```

<!-- This is experimental. -->

<!-- ```{r graphs, fig.height=20, fig.width=15, eval=FALSE} -->
<!-- diff <- df %>%  -->
<!--   select(-contains("method")) %>% -->
<!--   left_join(CATH_df, by=c("query"="target")) %>%  -->
<!--   left_join(CATH_df, by=c("target"), suffix = c("_query", "_target")) %>%  -->
<!--       filter(query!=target, n_Pfam_domains_query==1, n_3d_domains_query==1, -->
<!--           n_Pfam_domains_target==1, n_3d_domains_target==1) %>%  -->
<!--   filter(Pfam_query != Pfam_target | Gene3D_query != Gene3D_target) %>%  -->
<!--     select(singleton, Pfam_query, Pfam_target, Gene3D_query, Gene3D_target) -->

<!-- graph_struct <- filter(diff, Gene3D_query!=Gene3D_target) %>%  -->
<!--     group_by(singleton, Gene3D_query, Gene3D_target) %>%  -->
<!--     count() %>%  -->
<!--     filter(n>30) %>%  -->
<!--     mutate(Gene3D_query=gsub(";","",Gene3D_query), -->
<!--            Gene3D_target=gsub(";","",Gene3D_target)) %>%  -->
<!--     select(Gene3D_query, Gene3D_target, n, singleton) %>%  -->
<!--     as_tbl_graph() -->

<!-- graph_pfam <- filter(diff, Pfam_query!=Pfam_target) %>%  -->
<!--     group_by(singleton, Pfam_query, Pfam_target) %>%  -->
<!--     count() %>%  -->
<!--     filter(n>30) %>%  -->
<!--     mutate(Pfam_query=gsub(";","",Pfam_query), -->
<!--            Pfam_target=gsub(";","",Pfam_target)) %>%  -->
<!--     select(Pfam_query, Pfam_target, n, singleton) %>%  -->
<!--     left_join(nms_pfam,by=c("Pfam_query"="Pfam")) %>%  -->
<!--     left_join(nms_pfam,by=c("Pfam_target"="Pfam")) %>% -->
<!--     ungroup() %>%  -->
<!--     select(abbr.x, abbr.y, n, singleton) %>%  -->
<!--     as_tbl_graph() -->

<!-- graph_struct_plot <- ggraph(graph_struct, "graphopt") +  -->
<!--     geom_edge_fan(aes(width=n, color=singleton), show.legend = FALSE) +  -->
<!--     geom_node_point() +  -->
<!--     geom_node_text(aes(label = name),  colour = 'black', size=4, -->
<!--                   show.legend = FALSE, family = "serif", repel = T) + -->
<!--     # facet_edges(~singleton) +  -->
<!--     scale_edge_color_manual(values = palette_singleton) -->

<!-- graph_pfam_plot <- ggraph(graph_pfam, "stress") +  -->
<!--     geom_edge_fan(aes(width=n, color=singleton), show.legend = FALSE) +  -->
<!--     geom_node_point() +  -->
<!--     geom_node_text(aes(label = name),  colour = 'black', size=4, -->
<!--                   show.legend = FALSE, family = "serif", repel = T) + -->
<!--     # facet_edges(~singleton) +  -->
<!--     scale_edge_color_manual(values = palette_singleton) -->

<!-- graph_pfam_plot / graph_struct_plot -->
<!-- ``` -->

# TODO

### 2024-06-04 workflow/scripts/explore_CATH.R

@@ -1,0 +1,66 @@
+library(tidyverse)
+source("workflow/scripts/functions.R")
+theme_set(theme_classic())
+
+CATH_df <- read_delim(list.files("results/data/cath/", full.names = TRUE), delim = "\t", id = "proteome") %>% 
+  filter(!is.na(Gene3D)) %>% 
+  mutate(proteome=gsub("_cath.tsv", "", basename(proteome)), 
+         n_domains = str_count(Gene3D, ";")) %>% 
+  # filter(n_domains==1) %>% 
+  mutate(#Gene3D = gsub(";", "", Gene3D),
+         target=paste0("AF-", Entry, "-F1"))
+
+nms <- read_delim("test/cath/cath-names.txt", comment = "#", delim = "    ",
+                  col_names = c("Gene3D", "abbr", "name")) %>% 
+  select(-abbr) %>% 
+  mutate(name=gsub("^:", "", name))
+
+ggplot(CATH_df, aes(n_domains)) + 
+  geom_histogram(binwidth = 1, position = "dodge") +
+  scale_x_continuous(breaks = seq(1,100,1))
+
+# you can join filtered blast and foldseek information (the ones that reach the trees)
+# Then see the distinct number of CATH, CAT, CA and C
+
+blast <- read_delim(readLines("test/cath/testblast.files"), col_names = blast_columns)
+fs <- read_delim(readLines("test/cath/testfs.files"), col_names = fs_columns)
+
+df <- full_join(mutate(select(fs, query, target), method="fs"), mutate(select(blast, query, target), method="blast"),
+          by = c("query", "target"), suffix=c("_fs", "_blast")) %>% 
+  mutate(singleton = case_when(is.na(method_fs) ~ "only_blast",
+                        is.na(method_blast) ~ "only_fs", 
+                        .default = "common"))
+
+inner_join(df, select(CATH_df, -proteome, -Entry)) %>%
+  group_by(query, singleton, n_domains) %>% 
+  summarise(n_distinct_CATH = n_distinct(Gene3D)) %>% 
+  filter(n_domains==1) %>% 
+  ggplot(aes(n_distinct_CATH, fill=singleton)) + 
+  geom_histogram(binwidth = 1, position = "dodge") +
+  # facet_grid(n_domains~., scales = "free_y") +
[…]

### 2024-06-04 workflow/scripts/functions.R

@@ -33,6 +33,10 @@ wespal <- wesanderson::wes_palette("Zissou1", 100, type = "continuous")
 palettes_step <- c( "#F0E442", "#009E73", "#CC79A7")
 names(palettes_step) <- c("foldtree", "iqtree", "quicktree")
 
+palette_domains <- c("#31b7bcff", "#95c11fff")
+names(palette_domains) <- c("Pfam", "3d")
+
+
 add_marginal <- function(df, plot, x, y, fill, palette){
   xdens <- axis_canvas(plot, axis = "x")+
     geom_density(data = df, aes(x = .data[[x]], fill = .data[[fill]]),
@@ -83,3 +87,44 @@ get_idxs_df <- function(seqs, aln, trimmed) {
   }
   return(df_idxs)
 }
+
+
+plot_dom <- function(tree, seed, gff, prot, singleton_df, title="") {
+
+  tree <- ape::root(tree, paste0("AF-", seed, "-F1"), resolve.root=TRUE)
+  tree$edge.length <- (tree$edge.length/max(diag(ape::vcv(tree))))
+
+  lbls <- fortify(tree, ladderize = FALSE) %>% 
+    filter(isTip) %>% 
+    arrange(y) %>% 
+    pull(label)
+
+  a <- gff %>%
+    filter(seqid %in% gsub("AF-|-F1", "", tree$tip.label), type=="Domain") %>% 
+    mutate(Note=gsub(";..*", "", gsub("Note=", "", rest)),
+           seqid=factor(seqid, levels=gsub("AF-|-F1", "", lbls), ordered = TRUE)) %>% 
+    ggplot(aes(x=start, xend=end, y=seqid, yend=seqid)) + 
+    geom_segment(aes(x=1, xend=width,
+                     y=seqid, yend=seqid), 
+                 data = prot %>% mutate(seqid=factor(seqid, levels=gsub("AF-|-F1", "", lbls), ordered = TRUE))) +
+    geom_segment(aes(x=start, xend=end,
+                     y=seqid, yend=seqid, color=Note),  
+                 linewidth = 2) + 
+    geom_text(aes(label = Note, x=(start+end)/2, y=seqid), size=2) +
+    theme_void() +
+    labs(title = title) +
+    theme(legend.position = "none")
+  
+  b <- fortify(tree, ladderize = FALSE) %>% 
+    left_join(singleton_df, by="label") %>% 
[…]

### 2024-06-04 workflow/scripts/plot_example.R

@@ -23,18 +23,22 @@ sptree <- read.tree(snakemake@input[["sptree"]])
 translate <- deframe(distinct(dplyr::select(taxmap, Proteome_ID, mnemo)))
 sptree$tip.label <- names(translate)[match(translate, sptree$tip.label)]
 
-species_dist_df <- as.matrix(adephylo::distTips(sptree, method = "nNodes"))[snakemake@params[["seed"]], ] %>% 
+species_dist_df <- as.matrix(adephylo::distTips(sptree, method = "nNodes"))[seed_sp, ] %>% 
   enframe(name = "Proteome_ID", value = "dist")
 
-gff <- rtracklayer::readGFF(snakemake@input[["gff"]]) %>%
-  as_tibble() %>% 
-  filter(seqid %in% c(seeds), 
-         type %in% c(interesting_types, "Chain"))
+gff <- read_delim(c(snakemake@input[["gff"]]), 
+                  delim = "\t", comment = "#", 
+                  col_names = c("seqid", "source", "type",
+                                "start", "end", "score",
+                                "strand", "phase", "rest"),
+                  show_col_types = FALSE) %>% 
+  filter(type %in% c(interesting_types, "Chain"))
 
 reco <- read_delim(snakemake@input[["reco"]], show_col_types = FALSE) %>% 
   mutate(model=factor(model, levels=models))
 
 plot_list = list()
+plot_domains = list()
 
 for (seed in seeds) {
   print(seed)
@@ -81,16 +85,17 @@ for (seed in seeds) {
 
   seqs <- Biostrings::readAAStringSet(fls[grepl("union.*_aa.seqs", fls)])
   length <- nchar(as.character(seqs[[paste0("AF-", seed, "-F1")]]))
-  prot <- tibble(seqid=seed, start=1, end=length)
+  prot <- as.data.frame(seqs@ranges) %>% 
+    mutate(seqid=gsub("AF-|-F1", "", names))
   
   plot_seed <- gff %>%
     filter(seqid==seed, type!="Chain") %>% 
     rowwise() %>% 
-    mutate(Note=ifelse(length(Note)==0, "", Note)) %>% 
+    mutate(Note=gsub(";..*", "", gsub("Note=", "", rest))) %>% 
     # filter(!type %in% c("Helix", "Beta strand", "Turn")) %>% 
     ggplot() + 
-    geom_segment(aes(x=start, xend=end,
[…]

### 2024-06-26 workflow/notebooks/draft.Rmd
fixed apro plot

---
title: "Structural analysis"
# knit: (function(input_file, encoding) {
#   rmarkdown::render(input_file,
date: "`r format(Sys.time(), '%d-%b-%Y')`"
author: "Giacomo Mutti"
output:
    rmdformats::html_clean:
        number_sections: false
        code_folding: hide
        self_contained: true
        fig_caption: true
        # thumbnails: false
        # lightbox: false
        gallery: true
        toc_depth: 3
        highlight: kate
---

```{r setup, include=FALSE, dev="CairoPNG"}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, cache.lazy = FALSE,
                      crop = TRUE, dev="CairoPNG")
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
knitr::opts_knit$set(root.dir = "../../")
```

```{r libraries, message=FALSE, warning=FALSE, cache=FALSE}
library(tidyverse)
library(patchwork)
library(ggtree)
library(cowplot)
library(glue)
source("workflow/scripts/functions.R")
theme_set(theme_classic())

models <- names(palettes_model)

params <- yaml::read_yaml("config/params.yaml")
params_dataset <- yaml::read_yaml("config/Hsapopi_test.yaml")

outdir <- glue(params$outdir, params_dataset$dataset)
```

```{r read_stuff, cache=FALSE}
seeds <- readLines(glue(outdir, "/ids/UP000005640_common.ids"))

tax <- read_delim(params_dataset$taxons_file, show_col_types=F)
taxmap <- read_delim(glue(outdir, "/db/taxidmap"), 
                     delim = "\t", col_names = c("label", "Tax_ID"),
                     show_col_types = FALSE) %>% 
  left_join(tax, by = "Tax_ID") %>% 
  dplyr::select(label, Proteome_ID, mnemo)

taxidmap <- read_delim(glue(outdir,"/db/taxidmap"), 
                       col_names = c("target", "Tax_ID"), show_col_types=F)

sptree <- read.tree(params_dataset$species_tree_labels)
translate <- deframe(distinct(dplyr::select(taxmap, Proteome_ID, mnemo)))
# sptree$tip.label <- names(translate)[match(translate, sptree$tip.label)]

species_dist_df <- tibble()
for (sp in sptree$tip.label){
  if (sp=="Hsap") {next}
  dist <- castor::get_all_distances_to_tip(sptree, "Hsap")[MRCA(sptree, "Hsap", sp)]
  mrca <- c(sptree$tip.label, sptree$node.label)[MRCA(sptree, "Hsap", sp)]
  species_dist_df <- bind_rows(species_dist_df, tibble(mnemo=sp, value=dist, mrca=mrca))
}

# read input table
lvls_tax <- rev(c(unique(tax$Clade), "archaea", "bacteria"))

table_columns <- c('Proteome_ID', 'Tax_ID', 'count1', 'count2', 'count3', 
                   'genome', 'source', 'species', 'mnemo')

table <- read_delim(params_dataset$input_table_file, show_col_types=F, 
                    delim = "\t", col_names = table_columns) %>% 
  left_join(tax, by = "Tax_ID") %>% 
  mutate(Clade = ifelse(is.na(Clade), mnemo, Clade),
         Clade = factor(Clade, levels = lvls_tax))

lddt_df <- read_delim(paste0(params$structure_dir, "/", tax$Proteome_ID, "/", tax$Proteome_ID, "_mean_plddt.tsv"),
           col_names = c("target", "mean_lddt"), show_col_types = FALSE)

aln_seeds <- readLines(glue(outdir,"/ids/", params_dataset$seed, "_aln.ids"))
```

# Homology

```{r read_self_blast}
blast_self <- read_delim(glue(outdir, "/homology/allvall/",
                              params_dataset$seed, "_", params_dataset$seed, "_blast.tsv"),
                         show_col_types=FALSE,
                         col_names = blast_columns) %>% 
  # filter(query %in% seeds) %>%
  filter(query==target) %>% 
  mutate(blast_selfbit=bitscore) %>% 
  select(query, blast_selfbit)

fs_self <- read_delim(glue(outdir, "/homology/allvall/",
                              params_dataset$seed, "_", params_dataset$seed, "_fs.tsv"), 
                      show_col_types = FALSE,
                         col_names = blast_columns) %>% 
  # filter(query %in% seeds) %>%
  filter(query==target) %>% 
  mutate(fs_selfbit=bitscore) %>% 
  select(query, fs_selfbit)

self <- full_join(blast_self, fs_self, by="query")
```

```{r read_fs}
fs <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_fs.tsv"), 
                 col_names = fs_columns, show_col_types = FALSE) %>% 
  # filter(query %in% seeds) %>%
  mutate(evalue = ifelse(evalue==0, 1e-180, evalue)) %>%
  left_join(taxidmap, by = "target")

fs_brhs <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_fs_brh.tsv"),
                     col_names = c("query", "target"), show_col_types = FALSE) %>% 
  # filter(query %in% seeds) %>%
  mutate(method="fs_brh") 
```

```{r read_blast}
blast <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_blast.tsv"), 
                    col_names = blast_columns, show_col_types = FALSE) %>% 
  # filter(query %in% seeds) %>%
  mutate(evalue = ifelse(evalue==0, 1e-180, evalue))

blast_brhs <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_blast_brh.tsv"),
                     col_names = c("query", "target"), show_col_types = FALSE) %>% 
  # filter(query %in% seeds) %>%
  mutate(method="blast_brh")
```

```{r merge_homo}
df <- full_join(blast, fs,
                by = c("query", "target"), suffix=c("_Blast", "_Foldseek")) %>%
    filter(query!=target) %>% 
  left_join(self, by="query") %>%
  mutate(pident_Foldseek = 100*pident_Foldseek, 
         qcov_Foldseek = 100*qcov_Foldseek,
         singleton = case_when(is.na(evalue_Foldseek) ~ "only Blast",
                               is.na(evalue_Blast) ~ "only Foldseek", 
                               .default = "Common"))

brhs <- rbind(blast_brhs, fs_brhs) %>% 
    group_by(query, target, method) %>%
    mutate(n=1) %>%
    pivot_wider(names_from = method, values_fill = 0, values_from = n) %>%
    mutate(common = ifelse(fs_brh+blast_brh==2, 1, 0))

# grouped_brhs <- brhs %>%
#   group_by(query) %>%
#   summarise(fs_brh = sum(fs_brh), blast_brh=sum(blast_brh), common_brh=sum(common))
blast_filtered <- filter(blast, 
                         query %in% paste0("AF-", aln_seeds, "-F1"),
                         length/qlen*100>params$coverage, 
                         length/slen*100>params$coverage,
                         evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

fs_filtered <- filter(fs, 
                      query %in% paste0("AF-", aln_seeds, "-F1"),
                      qcov*100>params$coverage, 
                      tcov*100>params$coverage, 
                      evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

df_filtered <- full_join(blast_filtered, fs_filtered,
                by = c("query", "target"), suffix=c("_Blast", "_Foldseek")) %>%
    filter(query!=target) %>% 
  left_join(self, by="query") %>%
  mutate(pident_Foldseek = 100*pident_Foldseek, 
         qcov_Foldseek = 100*qcov_Foldseek,
         singleton = case_when(is.na(evalue_Foldseek) ~ "only Blast",
                               is.na(evalue_Blast) ~ "only Foldseek", 
                               .default = "Common"))
```

```{r cath}
CATH_df <- read_delim(list.files(params$cath_dir, full.names = TRUE),
                      delim = "\t", id = "proteome", show_col_types = FALSE) %>%
  # filter(!is.na(Gene3D)) %>% 
  mutate(proteome = gsub("_cath.tsv", "", basename(proteome)),
         n_3d_domains = str_count(Gene3D, ";"),
         n_Pfam_domains = str_count(Pfam, ";")) %>%
  # filter(n_domains==1) %>% 
  mutate(target=paste0("AF-", Entry, "-F1")) %>% 
    select(-proteome, Entry)

nms <- read_delim("test/cath/cath-names.txt", 
                  comment = "#", delim = "    ",
                  col_names = c("Gene3D", "abbr", "name"), 
                  show_col_types = FALSE) %>% 
  select(-abbr) %>% 
  mutate(name=gsub("^:", "", name))

nms_pfam <- read_delim("test/cath/Pfam-A.clans.tsv", 
           col_names = c("Pfam", "clan", "boh", "abbr", "name"), 
           show_col_types = FALSE) %>% 
  select(-name, -clan, -boh)

long_CATH_df <- CATH_df %>% 
  # filter(target %in% unique(blast$query)) %>% 
  mutate(Gene3D=strsplit(Gene3D, ";")) %>% 
  unnest(Gene3D) %>% 
  separate(col = Gene3D, into = c("C", "A", "T", "H"), sep = "\\.", remove = FALSE) %>% 
  mutate(CA = paste(C,A,sep = "."), CAT=paste(C,A,T,sep = ".")) %>% 
  left_join(nms, by=c("C"="Gene3D")) %>%  
  left_join(nms, by=c("CA"="Gene3D"), suffix = c("_C", "_CA")) %>% 
  left_join(nms, by=c("CAT"="Gene3D"), suffix = c("_CA", "_CAT")) %>% 
  left_join(nms, by=c("Gene3D"), suffix = c("_CAT", "_CATH"))

long_CATH_nms <- long_CATH_df %>%
    select(-c(Entry, Pfam, n_Pfam_domains, target, n_3d_domains)) %>% 
    distinct()
```

## Fig2

```{r fig2}
evals <- c(Inf, 1e-2, 1e-3, 1e-5)
min_covs <- c(0, 30, 80)
eval_p <- ggplot(df, aes(-log10(evalue_Blast), -log10(evalue_Foldseek))) + 
  stat_bin_2d(breaks=seq(0,20,0.4)) + 
  scale_fill_gradientn(colors=wespal) +
  geom_hline(yintercept = -log10(evals), linetype=2) + 
  geom_vline(xintercept = -log10(evals), linetype=2) + 
  coord_fixed(expand = 0) +
  labs(x="-log10(Evalue Blast)",y="-log10(Evalue Foldseek)") +
  scale_x_continuous(limits = c(0, 20)) +
  scale_y_continuous(limits = c(0, 20)) +
  theme(legend.position = "none", panel.grid = element_blank()) 
eval_p <- add_marginal(mutate(df, 
                              evalue_Blast=-log10(evalue_Blast), 
                              evalue_Foldseek=-log10(evalue_Foldseek)), 
                       eval_p, "evalue_Blast", "evalue_Foldseek", "singleton", palette_singleton_p)

pident_p <- ggplot(df, aes(pident_Blast, pident_Foldseek)) + 
  stat_bin_2d(breaks=seq(0,100,2)) + 
  scale_fill_gradientn(colors=wespal) +
  # scale_fill_viridis_c(option = "C") +
  coord_fixed(expand = 0) +
  scale_x_continuous(limits = c(0, 100)) +
  scale_y_continuous(limits = c(0, 100)) +
  theme(legend.position = "none", panel.grid = element_blank())
pident_p <- add_marginal(df, pident_p, "pident_Blast", "pident_Foldseek", "singleton", palette_singleton_p)

qcov_p <- ggplot(df, aes(qcov_Blast, qcov_Foldseek)) + 
  stat_bin_2d(breaks=seq(0,100,2)) + 
  geom_hline(yintercept = min_covs, linetype=2) + 
  geom_vline(xintercept = min_covs, linetype=2) + 
  scale_x_continuous(limits = c(0, 100)) +
  scale_y_continuous(limits = c(0, 100)) +
  scale_fill_gradientn(colors=wespal) +
  coord_fixed(expand = 0) +
  theme(legend.position = "none", panel.grid = element_blank()) 
qcov_p <- add_marginal(df, qcov_p, "qcov_Blast", "qcov_Foldseek", "singleton", palette_singleton_p)

lddt_p <- df %>% 
  select(lddt, singleton) %>% 
  ggplot(aes(color=singleton, lddt)) + 
  geom_density() + 
  # scale_y_continuous(expand = c(0.01,0)) +
  # scale_x_continuous(limits = c(params$low_confidence, 100),
  #                    expand = c(0,0)) +
  scale_color_manual(values = palette_singleton_p)+ 
  theme(legend.position = "none", 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

alntmscore_p <- df %>% 
  select(alntmscore, singleton) %>% 
  ggplot(aes(color=singleton, alntmscore)) + 
  geom_density() + 
  # coord_cartesian(expand = 0) +
  scale_color_manual(values = palette_singleton_p)+ 
  theme(legend.position = "none", 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

plddt_p <- df %>% 
  select(target, singleton) %>% 
  mutate(target=gsub("AF-|-F1", "", target)) %>% 
  left_join(lddt_df, by="target") %>% 
  ggplot(aes(color=singleton, mean_lddt)) + 
  geom_density(bw=1) + 
  # scale_y_continuous(expand = c(0.01,0)) +
  scale_x_continuous(limits = c(params$low_confidence, 100)) +
  scale_color_manual(values = palette_singleton_p)+ 
  theme(legend.position = "none", 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

df_multi <- df_filtered %>% 
    # head(n = 100000) %>%
  left_join(CATH_df, by=c("query"="target")) %>% 
  filter(!is.na(Pfam), !is.na(Gene3D)) %>% 
  left_join(CATH_df, by="target") %>% 
  rowwise() %>% 
  mutate(Gene3D.x=strsplit(Gene3D.x, ";"),
         Pfam.x=strsplit(Pfam.x, ";"),
         Gene3D.y=strsplit(Gene3D.y, ";"),
         Pfam.y=strsplit(Pfam.y, ";")) %>% 
  mutate(n_3d_common=length(intersect(Gene3D.x,Gene3D.y)),
         n_Pfam_common=length(intersect(Pfam.x,Pfam.y))) %>% 
  # filter(n_Pfam_common>0, n_3d_common>0) %>% 
  select(-c(Pfam.x,Pfam.y,Gene3D.x,Gene3D.y)) %>% 
  mutate(Jaccard_3d=n_3d_common/pmax(n_3d_domains.x,n_3d_domains.y),
         Jaccard_Pfam=n_Pfam_common/pmax(n_Pfam_domains.x,n_Pfam_domains.y)) %>% 
  select(query, target, singleton, contains("Jaccard"), contains("common"))

CATH_plot <- df_multi %>% 
    pivot_longer(contains("Jaccard"), values_to = "Jaccard") %>% 
    mutate(name=gsub("Jaccard_", "", name),
           name=gsub("3d", "CATH", name)) %>% 
    ggplot(aes(Jaccard, fill=singleton)) +
    geom_histogram(binwidth = 0.1, position = "dodge") +
    facet_grid(name~.) +
    scale_x_continuous(breaks = seq(0,1,0.1)) +
    scale_y_continuous(expand = expansion(0,0)) + 
    scale_fill_manual(values=palette_singleton_p) + 
  theme(legend.position = "bottom")

# prop_p <- group_by(df_filtered) %>% 
#   count(query, singleton) %>% 
#   group_by(query) %>% 
#   mutate(nn=sum(n), prop=n/nn) %>% 
#   ggplot(aes(prop, singleton, color=singleton)) + 
#   geom_boxplot() + 
#   labs(x="Proportion in tree", y="") +
#   scale_color_manual(values = palette_singleton_p) + 
#   theme(legend.position = "none")

fig2_first_row <- (pident_p | eval_p | qcov_p)
fig2_second_row <- ((lddt_p / alntmscore_p / plddt_p) | CATH_plot) + plot_layout(widths = c(1,2))
fig2 <-  (fig2_first_row / fig2_second_row)

ggsave("draft/figures/fig2.pdf", fig2, width = 9, height = 6)
```

## Fig3

```{r read_trees}
df_trees <- read_delim(glue(outdir,"/trees/", params_dataset$seed, "_unrooted_trees.txt"),
                       show_col_types = FALSE,
                       col_names = c("gene", "target", "alphabet", "model", "tree")) %>% 
  mutate(model=factor(model, levels=models)) %>% 
  filter(model!="FTPY")

# ML trees with selected model and LL
df_ml <- read_delim(glue(outdir,"/trees/", params_dataset$seed, "_mltrees.txt"), 
                    delim = "\t", 
                    show_col_types=FALSE) %>% 
  # mutate(fn=gsub("_mltrees.txt", "", basename(fn))) %>% 
  # separate(fn, c("gene", "target", "alphabet", "model"), sep = "_") %>% 
  mutate(model=factor(gsub("3DI", "3Di", 
                           gsub("GTR20", "GTR",
                                gsub("\\+.*", "", Model))), levels=models), 
         rate=sub("\\+", "", gsub("GTR20|3DI|LG", "", Model)))

# Read all trees
ts <- read.tree(text = df_trees$tree)
names(ts) <- paste(df_trees$gene, df_trees$target, df_trees$alphabet, df_trees$model, sep = "_")

# Compare trees with RF distance
rf_df <- filter(df_trees, target=="common") %>% 
  select(-alphabet) %>% 
  inner_join(x=., y = ., by = c("gene", "target")) %>%
  filter(!is.na(tree.x), !is.na(tree.y)) %>%
  rowwise() %>%
  mutate(RF = phangorn::RF.dist(normalize = TRUE, 
                                read.tree(text = tree.x), read.tree(text=tree.y)),
         n_tips = length(read.tree(text = tree.x)$tip.label)) %>%
  select(-tree.x, -tree.y)

rf_text <- group_by(rf_df, model.x, model.y) %>% 
  filter(as.integer(model.x)<=as.integer(model.y)) %>% 
  summarise(value=median(RF, na.rm=T))

# get BS distribution
df_bs <- fortify(ts) %>%
  filter(!isTip) %>%
  mutate(support=as.numeric(label)) %>%
  separate(.id, c("gene", "target", "alphabet", "model")) %>%
  filter(model!="QT") %>% 
  filter(!is.na(support)) %>% 
  select(gene, target, model, support)

# Reconciliation ranger+TCS scores
reco <- read_delim(glue(outdir,"/reco/", params_dataset$seed, "_DTL.tsv"), 
                   show_col_types = FALSE)

scores <- read_delim(glue(outdir,"/reco/", params_dataset$seed, "_scores.tsv"), 
                     show_col_types = FALSE) %>% 
  left_join(reco) %>% 
  mutate(model=factor(model, levels=models),
         scores, n_events = (dups+losses)/n_tips)
```

```{r fig3}
my_comparisons <- list( c("LG", "3Di"), c("3Di", "GTR"), c("LG", "GTR") )
plot_bs <- df_bs %>% 
  # group_by(gene, target, model) %>% 
  # summarise(mn_bs=mean(support)) %>% 
  ggplot(aes(model, support, fill=target)) +
  # geom_violin(position = position_dodge(width = 1)) +
  geom_boxplot() +
  # ggdist::stat_pointinterval(position = position_dodge(width = 1), color="black") +
  scale_fill_manual(values = palettes_method) +
  theme_classic() +
  labs(y="Boostrap", x="") +
  theme(legend.position = "none") 
  # ggpubr::stat_compare_means(aes(group = model),
  #                            comparisons = my_comparisons,
  #                            label = "p.format")

rf_plot <- rf_df %>% 
  ggplot(aes(RF)) +
  geom_tile(aes(x=0.5, y=0.5, fill=value), data = rf_text) +
  geom_text(aes(x=0.5, y=.5, label=round(value, 2)), color="white", data = rf_text) +
  geom_rug(linewidth=.1, data= . %>% filter(as.integer(model.x)>as.integer(model.y))) +
  geom_density(aes(y = after_stat(scaled)), data= . %>% filter(as.integer(model.x)>as.integer(model.y))) + 
  scale_fill_distiller(palette="YlGnBu", direction = -1, limits = c(0,1), 
                       name="Median RF") + 
  scale_x_continuous(breaks = c(0,.25,.5, .75,1), labels = function(x) sprintf("%.2g", x)) +
  coord_fixed(xlim = c(0,1), expand=0) +
  # facet_grid(method~method2, switch = "both") +
  facet_grid(model.x~model.y) +
  # ggh4x::facet_nested(targets+model~targets2+model2, nest_line = element_line(linetype = 1)) +
  labs(x="RF Distance") +
  theme_classic() +
  theme(legend.position = "bottom", 
        panel.background = element_rect(fill='transparent'),
        strip.placement = "outside",
        strip.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip.text = element_text(size=6),
        axis.text.x = element_text(size=6),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())

comps <- NULL
idx <- 1
for (el in df_trees$target %>% unique()){
  for (el2 in df_trees$target %>% unique()){
    if (el>el2){
    comps[[idx]] <- c(el, el2)
    idx <- idx+1
    }
  }
}

ttest <- scores %>%
  filter(model!="FTPY", id %in% id_tmp) %>%
  group_by(model) %>% 
  rstatix::t_test(n_events~targets) %>% 
  rstatix::add_xy_position(x="model", dodge = 0.8)

bxp <- scores %>% 
  filter(model!="FTPY", id %in% id_tmp) %>%
  ggpubr::ggboxplot(x = "model", y = "n_events", fill = "targets",
                    palette = palettes_method)

bxp +
ggpubr::stat_pvalue_manual(ttest, label = "p")

ggplot(scores,aes(n_tips, dups+losses, color=targets)) + 
  geom_point(size=.1) +
  geom_smooth(method = "lm", alpha=.2) +
  facet_wrap(~model, nrow = 1,scales = "free") +
  # coord_cartesian(xlim = c(0,50), ylim = c(0,250)) +
  scale_color_manual(values = palettes_method)

plot_DL <- scores %>% 
  filter(model!="FTPY") %>% 
  ggplot(aes(model, n_events, fill=targets)) +
  geom_boxplot() +
  labs(x="", y="(D+L)/# Tips") +
  # ggthemes::scale_fill_solarized() +
  scale_fill_manual(values = palettes_method) +
  theme(legend.position = "none")

var_r2t <- scores %>% 
  filter(model!="FTPY", targets=="common") %>% 
  ggplot(aes(color=model, variance_r2t)) + 
  geom_density(size=1, alpha=.5) +
  # geom_histogram(position = "dodge") + 
  coord_cartesian(expand = 0) +
  scale_x_continuous(limits = c(0,.2)) + 
  # facet_grid(~targets) + 
  labs(x="Variance Root-to-tip") +
  scale_color_manual(values = palettes_model) +
  theme(legend.position = "bottom")

fig3 <- rf_plot | ((plot_bs | plot_DL) / var_r2t)
ggsave("draft/figures/fig3.pdf", fig3, width = 11, height = 6)
```

```{r FigS1}
df_red <- select(df_filtered, query, target, singleton)

res <- NULL
for (idx in 1:nrow(df_trees)){
  t <- ts[[idx]]
  seed <-  t$tip.label[grepl(df_trees$gene[idx], t$tip.label)]
  nn <- as.matrix(adephylo::distTips(t, method = "nNodes"))[seed, ]
  max_dist <- max(nn)

  row <- enframe(nn, name = "target", value="nn_dist") %>% 
    mutate(query=seed, 
           max_dist=max_dist,
           targets=df_trees$target[idx],
           alphabet=df_trees$alphabet[idx],
           model=df_trees$model[idx]) %>% 
    left_join(df_red, by=c("query", "target"))
  res <- bind_rows(res, row)
}

# for what this should be grouped?
plot_dist <- res %>% 
  filter(query!=target) %>% 
  group_by(query, singleton, model) %>%
  summarise(median_dist=median(nn_dist/max_dist), n=n()) %>% 
  ggplot(aes(median_dist, model,  fill=singleton)) + 
  ggdist::stat_halfeye(alpha=.8, point_colour = NA, .width = 0, height=.6, 
                       position = position_nudge(y=.15)) +
  geom_boxplot(
    width=.2,
    position = position_dodge(),
    # removing outliers
    outlier.color = NA,
    alpha = 1,
    # position= position_nudge(y=-.5)
  ) + 
  scale_x_continuous(expand = c(0,0.01), limits = c(0,1)) +
  theme(panel.border = element_rect(fill = NA)) + 
  scale_fill_manual(values = palette_singleton_p) + 
  theme(legend.position = "bottom")

labeled_sptree <- read.tree("data/sptrees/homo_internal.spTree.nw")
nodes_clades <- fortify(labeled_sptree) %>% 
  # filter(!isTip) %>% 
  mutate(ordered = rank(y)) %>% 
  mutate(label_int = fct_reorder(label, y)) %>%
  select(label_int, ordered)

clades <- read_delim(params_dataset$taxons_file, show_col_types = FALSE)

apro_files <- list.files(glue(outdir, "/reco"), 
                         pattern = "*support.nwk", full.names = TRUE)
apro_trees <- read.tree(text = sapply(apro_files, readLines))
names(apro_trees) <- sapply(str_split(basename(apro_files), "_"), 
                            function(x) paste0(x[2], "_", x[4]))

trees_df <- fortify(apro_trees) 
nodes_df <- trees_df %>% 
  group_by(.id) %>% 
  mutate(ordered = rank(y)) %>% 
  left_join(nodes_clades) %>% 
  separate(.id, c("targets", "model")) %>% 
  filter(!isTip, label!="") %>% 
  mutate(label = gsub("\\[|\\]|\\'", "", str_replace_all(label, "[a-z]{1,2}[0-9]=", ""))) %>% 
  separate(label, c("pp1", "pp2", "pp3", "f1", "f2", "f3", "q1", "q2", "q3"), 
           ";", convert = TRUE) %>% 
  mutate(freq=f1+f2+f3)

pp1_plot <- nodes_df %>%
    filter(model!="FTPY") %>% 
    ggplot(aes(q1, label_int, fill=model, color=model, group=model)) +
    geom_linerange(aes(x=x, xmin=xmin, xmax=xmax, color=model),
                   data = . %>% 
                     group_by(model, label_int) %>% 
                     summarise(x=mean(q1), xmin=min(q1), xmax=max(q1)),
                   position = position_dodge(width=.7), color="grey70") +
    geom_point(aes(shape=targets),
               position = position_dodge(width=.7)) +
  scale_fill_manual(values=palettes_model) +
  scale_color_manual(values = palettes_model_d) +
  scale_shape_manual(values = c(21,22,23,24)) +
  labs(y="Node") +
  guides(color = guide_legend(override.aes = list(pch = 21, size=3)),
         shape = guide_legend(override.aes = list(size=3))) +
  theme(legend.position = "bottom", legend.box="vertical") + 
  theme_bw()

sptree_plot <- ggtree(labeled_sptree) + 
  geom_tiplab() + 
  geom_nodelab(color="firebrick", 
               nudge_x = -.1, hjust = 1, 
               nudge_y = .2, size=2)+
  scale_x_continuous(expand = c(.2,.2))

fig_s1 <- (plot_dist + sptree_plot + pp1_plot) + plot_layout(widths = c(1.4, 1, 1.6))

ggsave("draft/figures/figS1.pdf", fig_s1, width = 14, height = 6)
```

### 2024-06-26 workflow/notebooks/eda.Rmd
fixed apro plot

@@
 
+First of all we can explore the model selection for the 3 different Maximum likelihood trees (GTR, 3Di and LG).
+We can see how the BIC scores between GTR and 3Di change and that generally 3DI seems to fit better (when BIC 3DI-BIC GTR < 0 it means that the 3di model is better).
+ModelFinder generally agrees for +R and +G across models.
+This is different for +F as it is clearly different when comparing LG and either of the structure models.
+For what concerns +I (Inv) this pattern is less clear.
+This could be explained as sequences and structure (although recoded) do not evolve in the same way and the empirical frequencies of the recoded status have completely different interpretations.
+
 ## Model fit

### 2024-06-26 workflow/scripts/analyze_apro.R

@@ -7,9 +7,10 @@ source("workflow/scripts/functions.R")
 
 labeled_sptree <- read.tree(snakemake@input[["sptree"]])
 nodes_clades <- fortify(labeled_sptree) %>% 
-  filter(!isTip) %>% 
-  mutate(label = fct_reorder(label, y)) %>% 
-  select(node, label)
+  # filter(!isTip) %>% 
+  mutate(ordered = rank(y)) %>% 
+  mutate(label_int = fct_reorder(label, y)) %>%
+  select(label_int, ordered)
 
 clades <- read_delim(snakemake@input[["groups"]], show_col_types = FALSE)
 
@@ -18,8 +19,10 @@ apro_trees <- read.tree(text = sapply(apro_files, readLines))
 names(apro_trees) <- sapply(str_split(basename(apro_files), "_"), function(x) paste0(x[2], "_", x[4]))
 
 trees_df <- fortify(apro_trees) 
-nodes_df <- trees_df %>% 
-  separate(.id, c("targets", "model")) %>% 
+nodes_df <- trees_df %>%   group_by(.id) %>% 
+  mutate(ordered = rank(y)) %>% 
+  left_join(nodes_clades) %>% 
+  separate(.id, c("targets", "model")) %>%  
   filter(!isTip, label!="") %>% 
   mutate(label = gsub("\\[|\\]|\\'", "", str_replace_all(label, "[a-z]{1,2}[0-9]=", ""))) %>% 
   separate(label, c("pp1", "pp2", "pp3", "f1", "f2", "f3", "q1", "q2", "q3"), ";", convert = TRUE) %>% 

### 2024-06-26 workflow/scripts/functions.R

@@ -16,14 +16,18 @@ palettes_model <- c("#BF360C", "#75C323", "#40A79D", "#FFA000", "#D4E157","#646B
 names(palettes_model) <- models
 palettes_model_d <- colorspace::darken(palettes_model, amount = 0.6)
 names(palettes_model_d) <- models
+# "#FF0000" "#00A08A" "#F2AD00" "#F98400" "#5BBCD6"
 
-palettes_method <- c("#F23030", "#F29F05", "#267365", "#D6FAB7")
+palettes_method <- c("#00A08A", "#d63a3a", "#ffaa0e", "#6f63bb")
 names(palettes_method) <- c("common", "blast", "fs", "union")
 
-palette_singleton <- c("#F23030", "#F29F05", "#267365", "#D6FAB7")
+palette_singleton <- c("#00A08A", "#d63a3a", "#ffaa0e", "#6f63bb")
 names(palette_singleton) <- c("common", "only_blast", "only_fs", "union")
 
-palette_brh <- c("#F23030","#F29F05", "#267365")
+palette_singleton_p <- c("#00A08A", "#d63a3a", "#ffaa0e", "#6f63bb")
+names(palette_singleton_p) <- c("Common", "only Blast", "only Foldseek", "Union")
+
+palette_brh <- c("#00A08A","#d63a3a", "#ffaa0e")
 names(palette_brh) <- c("common","blast_brh", "fs_brh")
 
 palettes_alphabet <- c("#327AD9", "#BF3475", "#F29F05")

### 2024-06-26 software/foldtree/compute_scores.py
simplified input formats + added conda option

@@ -1,72 +1,0 @@
-import treescore
-import toytree
-import pandas as pd
-import subprocess
-# import argparse
-import numpy as np
-
-from pathlib import Path
-
-
-T_score = 2000
-
-if __name__ == '__main__':
-    # args = parser.parse_args()
-    uniprot_df = pd.read_csv(snakemake.input[1], sep="\t", names=['query', 'mnemo','tax'])
-    taxidmap = pd.read_csv(snakemake.input[2], names=['query', 'species'], sep="\t")
-    # sptree = toytree.tree(snakemake.input[3])
-    # tmpfile = snakemake.output[0]
-
-    df = []
-    with open(snakemake.input[0]) as treefile:
-        for line in treefile:
-            line = line.strip().split()
-            gene = line[0]
-            targets = line[1]
-            alphabet = line[2]
-            model = line[3]
-            tree = toytree.tree(line[4], format = 0)
-            tree = treescore.label_leaves(tree, taxidmap, uniprot_df)
-            overlap = treescore.getTaxOverlap(tree.treenode)
-
-            leaves = tree.get_tip_labels()
-            n_sps = len(set(leaves))
-            n_tips = len(leaves)
-
-            # TCS score
-            taxscore = tree.treenode.score
-            # Branch lengths, normalized by total length
-            lengths = np.array([node.dist for node in tree.treenode.traverse()])
-            total_length = np.sum(lengths)
-            lengths /= total_length
-
-            # TCS score of root??
-            treescore.getTaxOverlap_root(tree.treenode)
-            root_score = treescore.sum_rootscore(tree.treenode)
-            # Tip to root distance
-            distances = np.array([ node.get_distance(tree.treenode) for node in tree.treenode.get_leaves() ])
-            # distances_norm = distances / np.mean(distances)
-
-            row = [gene, targets, alphabet, model, n_sps, n_tips, 
-                    taxscore, total_length, np.mean(lengths), np.mean(distances), 
-                    np.var(distances), root_score]
-            df.append(row)
-
-    outdf = pd.DataFrame(df, columns=['id', 'targets', 'alphabet', 'model', 'n_sps', 'n_tips',
-                                      'score', 'tree_length', 'mean_normalized_length', 'mean_r2t',
-                                      'variance_r2t', 'root_score'])
-
-    outdf.to_csv(snakemake.output[0], index=False, sep="\t")
-
-
[…]

### 2024-06-26 software/foldtree/corecut.py

@@ -1,101 +1,0 @@
-import pandas as pd
-import Bio.PDB
-import numpy as np
-import os
-import tqdm
-
-def extract_core(resdf, outfile,  hitthresh = .8, minthresh = .6, corefolder = 'core_structs/', structfolder = 'structs/', cterfolder = 'cter_structs/' , nterfolder = 'nter_structs/'):
-	"""
-
-	Extract a core of structures from a results file
-
-	Parameters
-		resdf: path to results file
-		outfile: path to output file
-		hitthresh: proportion of structures that need to map to a residue for it to be included in the core
-		minthresh: if no residues meet the hitthresh, the minimum proportion of structures that need to map to a residue for it to be included in the core
-		corefolder: name of folder to output core structures to
-		structfolder: name of folder to find structures in
-		cterfolder: name of folder to find cter structures in
-		nterfolder: name of folder to find nter structures in
-
-
-	"""
-	
-	#read all results
-	# folder =''.join([ sub + '/' for sub in resdf.split('/')[:-1] ])
-	# print(folder)
-	resdf = pd.read_table(resdf, header = None)
-	resdf.columns = 'query,target,fident,alnlen,mismatch,gapopen,qstart,qend,tstart,tend,evalue,bits,lddt,lddtfull,alntmscore'.split(',')
-
-	print('extracting core')
-	print('hitthresh: ' + str(hitthresh))
-	print('minthresh: ' + str(minthresh))
-
-	queries = resdf['query'].unique()
-	nqueries = len(queries)
-	print(resdf.head(), nqueries)
-	#map hits to each struc
-	hits = {}
-	#proportion of structures that need to map to a residue fo
-	with tqdm.tqdm(total=nqueries) as pbar:
-		for i,q in enumerate(queries):
-			sub = resdf[resdf['query'] == q]
-			hitvec = np.zeros((1, max(sub['qend'])) )
-			for idx,r in sub.iterrows():
-				hitvec[0,r['qstart']:r['qend']] = hitvec[0,r['qstart']:r['qend']]+1
-			hitvec /= nqueries
-			core = np.where(hitvec>hitthresh)[1]
-			try:
-				hits[q]= {'min': np.amin(core), 'max': np.amax(core)}
-			except:
-				#be more lenient...
-				subthresh = np.amax(hitvec)
[…]

### 2024-06-26 software/foldtree/foldseek2tree.py

@@ -1,243 +1,0 @@
-#foldssek cluster
-import subprocess ,shlex
-import numpy as np
-from scipy.spatial.distance import cdist
-import statsmodels
-import toytree
-import pandas as pd
-import re
-import os
-from scipy.stats import chi2
-import argparse
-
-def consensustree(treelist):
-	'''get a consensus tree from a list of tree files
-	
-	Parameters
-	----------
-	treelist : list
-		list of tree files
-
-	'''
-	#creat a multitree object from the list of tree files
-	treelist = [toytree.tree(i , format = 0 ) for i in treelist]
-	mt = toytree.mtree(treelist)
-	#get the consensus tree
-	ct = mt.get_consensus_tree( )
-	return ct
-
-#smooth distmat with MDS
-def MDS_smooth(distmat):
-	mds = MDS(n_components=int(distmat.shape[0]/2) )#, dissimilarity='precomputed'  )
-	distmat = mds.fit_transform(1-distmat )
-	distmat = cdist(distmat,distmat, 'minkowski', p=1.5 )
-	return distmat
-
-
-
-def Tajima_dist(kn_ratio, bfactor=1, iter = 100, delta=10**-5):
-	#approx of b*log(1-p/b) from https://doi.org/10.1093/oxfordjournals.molbev.a040031
-	#set all entries to smaller than bfactor
-	kn_ratio[kn_ratio > bfactor] = bfactor - delta
-	taj = np.add.reduce([ (kn_ratio**(np.ones(kn_ratio.shape)*i) )/ (bfactor**(i-1)*i) for i in range(1,iter) ] )
-	#set diagonal to 0
-	np.fill_diagonal(taj, 0)
-	return taj
-
-
-def runargs(args):
-	'''run a command line command
-	
-	Parameters
-	----------
-	args : str
-		command line command
-	'''
-	
-	args = shlex.split( args)
-	p = subprocess.run( args )
-	return p
-	
-def runFoldseekdb(folder , outfolder , foldseekpath = 'foldseek'):
-	'''run foldseek createdb
-	
-	parameters
-	----------
-	folder : str
-		path to folder with pdb files
-	outfolder : str 
-		path to output folder
-	
-
-	'''
-	args = foldseekpath + ' createdb '+  folder + ' '+ outfolder+'structblobDB '
-	p = runargs(args)
-	return outfolder+'structblobDB '
-
[…]

### 2024-06-26 software/foldtree/foldseekres2distmat_simple.py

@@ -1,61 +1,0 @@
-import numpy as np
-import pandas as pd
-
-def distmat_to_txt(identifiers, distmat, outfile):
-	'''
-	write out a distance matrix in fastme format
-
-	Parameters
-	----------
-	identifiers : list
-		list of identifiers for your proteins
-	distmat : np.array  
-		distance matrix
-	outfile : str   
-		path to output file
-
-	'''
-
-	#write out distmat in phylip compatible format
-	outstr = str(len(identifiers)) + '\n'
-	for i,pdb in enumerate(identifiers):
-		outstr += pdb + ' ' + ' '.join( [ "{:.4f}".format(d) for d in list( distmat[i,: ] )  ]  ) + '\n'
-	with open(outfile , 'w') as handle:
-		handle.write(outstr)
-		handle.close()
-	return outfile
-
-
-res = pd.read_table(snakemake.input[0], header = None)
-# print(res.head())
-#get the folder of the input file
-infolder = snakemake.input[0].split('/')[:-1]
-infolder = ''.join( [i + '/' for i in infolder])+'/'
-res[0] = res[0].map(lambda x :x.replace('.cif', ''))
-res[1] = res[1].map(lambda x :x.replace('.cif', ''))
-res.columns = 'query,target,fident,lddt,alntmscore'.split(',')
-ids = list( set(list(res['query'].unique()) + list(res['target'].unique())))
-pos = { protid : i for i,protid in enumerate(ids)}
-kernels = ['fident']#, 'alntmscore', 'lddt']
-
-#set kernel columns to float
-for k in kernels:
-    res[k] = res[k].astype(float)
-#change nan to 0
-res = res.fillna(0)
-matrices = { k:np.zeros((len(pos), len(pos))) for k in kernels }
-# print(res)
-
-#calc kernel for tm, aln score, lddt
-for idx,row in res.iterrows():
-    for k in matrices:
-        matrices[k][pos[row['query']] , pos[row['target']]] += row[k]
-        matrices[k][pos[row['target']] , pos[row['query']]] += row[k]
-
-for i,k in enumerate(matrices):
-    matrices[k] /= 2
-    matrices[k] = 1-matrices[k]
-    # print(matrices[k], np.amax(matrices[k]), np.amin(matrices[k]) )
-    # np.save( infolder + k + '_distmat.npy' , matrices[k])
-    distmat_txt = distmat_to_txt(ids, matrices[k], snakemake.output[i])
-

### 2024-06-26 software/foldtree/foldtree.py

@@ -1,138 +1,0 @@
-
-import foldseek2tree
-import corecut 
-import argparse
-import numpy as np
-import pandas as pd
-import os
-
-def structblob2tree(input_folder, outfolder, tmpfolder, corefolder, outtree, 
-                    overwrite = False, correction = False,
-                    fastmepath = 'fastme', foldseekpath = 'foldseek', delta = 0.0001,
-                    kernel = 'fident', core = False, hitthresh = .8, minthresh = .6):
-    '''
-    run fold tree pipeline for a folder of pdb files
-    
-    Parameters
-    ----------
-    input_folder : str
-        path to folder with pdb files   
-    outfolder : str
-        path to output folder   
-    overwrite : bool
-        overwrite existing foldseek output  
-    fastmepath : str    
-        path to fastme executable
-    quicktreepath : str 
-        path to quicktree executable
-    foldseekpath : str  
-        path to foldseek executable 
-    delta : float   
-        small number to replace negative branch lengths with, default is .0001
-    correction : str    
-        correction method to use, either 'tajima' or 'none'
-    kernel : str    
-        kernel to use, either 'fident', 'lddt' or 'alntmscore'
-    
-    '''
-    #check if the foldseek output is already there
-    input_folder = os.path.join(input_folder, '')
-    if os.path.exists(outfolder + '_allvall.tsv') and overwrite == False:
-        print('found foldseek output, skipping foldseek')
-        alnres = outfolder + '_allvall.tsv'
-    else:
-        alnres = foldseek2tree.runFoldseek_allvall_EZsearch(input_folder, outfolder + '_allvall.tsv', tmpfolder, foldseekpath = foldseekpath)
-    
-    if core == True:
-        corefolder = os.path.join(corefolder, '')
-        corecut.extract_core(alnres, outfolder+'.core',  hitthresh = hitthresh, minthresh = minthresh, 
-                             corefolder = corefolder, structfolder = input_folder)
-        if os.path.exists(outfolder + '_core_allvall.tsv') and overwrite == False:
-            print('found foldseek core output, skipping foldseek')
-            alnres = outfolder + '_core_allvall.tsv'
-        else:
-            alnres = foldseek2tree.runFoldseek_allvall_EZsearch(corefolder, outfolder + '_core_allvall.tsv', tmpfolder, foldseekpath = foldseekpath)
-
[…]

### 2024-06-26 software/foldtree/treescore.py

@@ -1,261 +1,0 @@
-import toytree
-import toyplot
-
-#taxonomy overlap score
-
-def standard_treedraw( tre, sizes= None , colors= None ,fixed_order=None, fixed_position=None , 
-ts = None,  save_file = False  , tiplabels = None ,  layout='c', edge_type='p' ):
-	
-	'''
-	standard treedraw function
-	
-	Parameters
-	----------
-	tre : toytree.tree
-		tree to draw
-	sizes : list
-		list of node sizes
-	colors : list
-		list of node colors
-	fixed_order : list
-		list of node names in order to draw
-	fixed_position : list
-		list of node positions
-	ts : treestlye : str
-		style of tree to draw (default = 'c')
-	save_file : str
-		path to save file
-	tiplabels : list
-		list of tip labels to draw
-	layout : str
-		layout of tree to draw (default = 'c')
-	edge_type : str
-		type of edge to draw (default = 'p')
-	'''
-
-	
-	if tiplabels is None:
-		tiplabels = tre.get_tip_labels()
-	canvas, axes, mark = tre.draw(  
-		ts = ts,
-		node_sizes=sizes,
-		node_colors=colors,
-		tip_labels_align=True,
-		scalebar=True,
-		fixed_order=fixed_order, 
-		fixed_position=fixed_position,
-		tip_labels=tiplabels,
-		layout=layout,
-		edge_type=edge_type,
-		tip_labels_style={
-			"fill": "#262626",
-			"font-size": "9px"}
-	)
-	if save_file:
-		toyplot.svg.render(canvas, save_file)
-
-
-
-def exp_score( v , exp = 1.5):
-	return v**exp
-
-def frac_score( v ):
-	return v+1
-
-#taxonomy overlap score
-def getTaxOverlap(node , treelen  = None , scorefun = frac_score):
-	
-	"""
-    Calculate the taxonomy overlap score for the given node in a phylogenetic tree.
-    
-    The taxonomy overlap score is defined as the number of taxonomic labels shared by all the leaf nodes
-    descended from the given node, plus the sum of the scores of all its children. If a leaf node has no
-    taxonomic label, it is not counted towards the score. The function also calculates the size of the
-    largest loss in lineage length, defined as the difference between the length of the set of taxonomic
-    labels shared by all the leaf nodes and the length of the longest set of taxonomic labels among the
-    children of the node.
-    
[…]

### 2024-06-26 workflow/envs/sp_python.post-deploy.sh

@@ -1,0 +1,9 @@
+# install pdbx
+git clone https://github.com/soedinglab/pdbx $CONDA_PREFIX/lib/pdbx
+cd $CONDA_PREFIX/lib/pdbx
+mkdir build
+cd build
+cmake -DUserInstallOption=ON ../
+make install
+
+cp -r build/lib/pdbx/ $CONDA_PREFIX/lib/python*/

### 2024-06-26 workflow/envs/sp_tree.post-deploy.sh

@@ -1,0 +1,12 @@
+# install astral-pro
+git clone https://github.com/chaoszhang/ASTER $CONDA_PREFIX/lib/ASTER
+cd $CONDA_PREFIX/lib/ASTER
+make astral-pro
+cp bin/astral-pro $CONDA_PREFIX/bin/
+cd -
+
+# install Ranger-DTL
+# wget https://compbio.mit.edu/ranger-dtl/ranger-dtl-linux.tar.gz -O $CONDA_PREFIX/lib/RangerDTL.tar.gz
+# tar xvzf $CONDA_PREFIX/lib/RangerDTL.tar.gz -C $CONDA_PREFIX/bin/ #--strip-components=2 ./bin/foldmason
+
+# 404 error. let's wait and see if they fix it, for now use one already installed.

### 2024-06-26 workflow/envs/sp_utils.post-deploy.sh

@@ -1,0 +1,6 @@
+# install foldmason
+# find other binaries here: https://mmseqs.com/foldmason/
+mkdir -p $CONDA_PREFIX/lib/
+wget https://mmseqs.com/foldmason/foldmason-linux-avx2.tar.gz -O $CONDA_PREFIX/lib/foldmason.tar.gz
+
+tar xvzf $CONDA_PREFIX/lib/foldmason.tar.gz -C $CONDA_PREFIX/bin/ --strip-components=2 ./bin/foldmason

### 2024-06-26 workflow/scripts/Treemmer_v0.3.py

@@ -1,673 +1,0 @@
-#Treemmer_v0.3
-
-#Copyright 2019 Fabrizio Menardo
-
-
-#   This program is free software: you can redistribute it and/or modify
-#    it under the terms of the GNU General Public License as published by
-#    the Free Software Foundation, either version 3 of the License, or
-#    (at your option) any later version.
-#
-#    This program is distributed in the hope that it will be useful,
-#    but WITHOUT ANY WARRANTY; without even the implied warranty of
-#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-#    GNU General Public License for more details.
-
-#    You should have received a copy of the GNU General Public License
-#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
-
-# Dependencies:
-#  ete3 
-#  joblib
-#  numpy
-#  matplotlib	
-
-# Tremmer_v0.3 is compatible with joblib 0.13.1 and 0.13.2
-
-
-
-
-# If you use Treemmer for your research, please cite:
-# Treemmer: a tool to reduce large phylogenetic datasets with minimal loss of diversity. Menardo et. al (2018),BMC Bioinformatics 19:164. https://doi.org/10.1186/s12859-018-2164-8
-
-from collections import defaultdict
-from joblib import Parallel, delayed
-from ete3 import Tree
-import sys
-import random
-import operator
-import argparse
-import csv
-
-
-
-############################################################			define arg type float 0 < X > 1		###############################################################
-
-def restricted_float(x):
-    x = float(x)
-    if x < 0.0 or x > 1.0:
-        raise argparse.ArgumentTypeError("%r not in range [0.0, 1.0]"%(x,))
-    return (x)
-
-
-##########################################			FIND LEAVES NEIGHBORS OF A LEAF (2 NODE OF DISTANCE MAX) and calc DISTANCE			#######################	
-
-def find_N(t,leaf):
-	dlist ={}
-	parent= leaf.up
-	dist_parent=leaf.dist
-	flag=0
-
-	if arguments.verbose==3:
-		print ("leaf findN at iteration:	" + str(counter))
-		print (leaf)
-		print ("parent findN at iteration:	" + str(counter))
-		print (parent)
-		print (parent.get_children())
-
-	sister_flag=0
[…]

### 2024-06-26 workflow/scripts/analyze_apro.R

@@ -12,7 +12,7 @@ nodes_clades <- fortify(labeled_sptree) %>%
   mutate(label_int = fct_reorder(label, y)) %>%
   select(label_int, ordered)
 
-clades <- read_delim(snakemake@input[["groups"]], show_col_types = FALSE)
+clades <- read_delim(snakemake@input[["table"]], show_col_types = FALSE)
 
 apro_files <- snakemake@input[["trees"]]
 apro_trees <- read.tree(text = sapply(apro_files, readLines))

### 2024-06-26 workflow/scripts/analyze_runtimes.R

@@ -7,7 +7,7 @@ df <- read_delim(snakemake@input[["time"]],
                  show_col_types = FALSE) %>% 
   mutate(basename=gsub(".txt", "", basename),
          step=basename(dirname),
-         max_rss_gb = measurements::conv_unit(max_rss, from = "MB", to = "GB")) %>% 
+         mutate(max_rss_gb = max_rss/1024)) %>% 
   separate(basename, c("seed", "gene", "method", "alphabet", "model"))
 
 

### 2024-06-26 workflow/scripts/compare_sampling.R

@@ -7,18 +7,18 @@ theme_set(theme_bw())
 taxidmap <- read_delim(snakemake@input[["taxidmap"]], 
                        col_names = c("target", "Tax_ID"), show_col_types=F)
 
-tax <- read_delim(snakemake@input[["groups"]], show_col_types=F)
-# read input table
-lvls_tax <- rev(c(unique(tax$Clade), "archaea", "bacteria"))
+# tax <- read_delim(snakemake@input[["groups"]], show_col_types=F)
+# # read input table
+# lvls_tax <- rev(c(unique(tax$Clade), "archaea", "bacteria"))
 
-table_columns <- c('Proteome_ID', 'Tax_ID', 'count1', 'count2', 'count3', 
-                   'genome', 'source', 'species', 'mnemo')
+# table_columns <- c('Proteome_ID', 'Tax_ID', 'count1', 'count2', 'count3', 
+#                    'genome', 'source', 'species', 'mnemo')
 
 table <- read_delim(snakemake@input[["table"]], show_col_types=F, 
-                    delim = "\t", col_names = table_columns) %>% 
-  left_join(tax, by = "Tax_ID") %>% 
-  mutate(Clade = ifelse(is.na(Clade), mnemo, Clade),
-         Clade = factor(Clade, levels = lvls_tax))
+                    delim = "\t") 
+  # left_join(tax, by = "Tax_ID") %>% 
+  # mutate(Clade = ifelse(is.na(Clade), mnemo, Clade),
+  #        Clade = factor(Clade, levels = lvls_tax))
 
 blast_self <- read_delim(snakemake@input[["self_blast"]], col_names = blast_columns) %>% 
   filter(query==target) %>% 
@@ -272,94 +272,3 @@ fs_p <- fs %>%
 saturation_plot <- (blast_p | fs_p) + plot_layout(guides = "collect")
 ggsave(snakemake@output[["saturation"]], saturation_plot, width = 7, height = 5, dpi = 300)
 
-# blast_single_df <- blast %>% 
-#   filter(evalue<opt$min_evalue) %>% 
-#   group_by(query, staxids) %>% 
-#   count() %>% 
-#   ungroup(staxids) %>% 
-#   mutate(is_single=ifelse(any(n>1), FALSE, TRUE), nspecies=length(unique(staxids))) %>% 
-#   group_by(query, is_single, nspecies) %>% 
-#   summarise()
-# 
[…]

### 2024-06-26 workflow/scripts/compare_sampling_old.R

@@ -1,197 +1,0 @@
-library("optparse")
-
-option_list <- list(
-  make_option(c("-b", "--blast"),
-              type = "character", default = NULL,
-              help = "blast file", dest = "blast"
-  ),
-  make_option(c("-f", "--fs"),
-              type = "character", default = NULL,
-              help = "foldseek file", dest = "fs"
-  ),
-  make_option(c("--fb"),
-              type = "character", default = NULL,
-              help = "foldseek brh", dest = "fs_brh"
-  ),
-  make_option(c("--bb"),
-              type = "character", default = NULL,
-              help = "blast brh", dest = "blast_brh"
-  ),
-  make_option(c("-o", "--output"),
-              type = "character", default = NULL,
-              help = "output filename", metavar = "character", dest = "outfile"
-  )
-)
-
-desc <- "Produces informative plot on foldseek and blast results comparison"
-
-opt_parser <- OptionParser(option_list = option_list, description = desc)
-opt <- parse_args(opt_parser)
-
-library(tidyverse)
-library(patchwork)
-source("scripts/functions.R")
-theme_set(theme_bw())
-
-# opt <- NULL
-# opt$fs <- "results/Ncra_opistho/homology/UP000001805_fs.tsv"
-# opt$blast <- "results/Ncra_opistho/homology/UP000001805_blast.tsv"
-# opt$outfile <- "results/Ncra_opistho/plots/plot_intersection.png"
-# opt$fs_brh <- "results/Ncra_opistho/homology/UP000001805_fs_brh.tsv"
-# opt$blast_brh <- "results/Ncra_opistho/homology/UP000001805_blast_brh.tsv"
-
-fs <- read_delim(opt$fs, col_names = fs_columns, show_col_types = FALSE)
-blast <- read_delim(opt$blast, col_names = blast_columns, show_col_types = FALSE)
-
-seeds <- union(unique(fs$query), unique(blast$query))
-
-evals <- c(Inf, 1e-2, 1e-3, 1e-5)
-min_covs <- c(0, 30, 60)
-
-fs_brhs <- read_delim(opt$fs_brh, col_names = c("query", "target"), show_col_types = FALSE) %>% 
-  group_by(query) %>% 
-  summarise(fs_brhs=list(target))
-
[…]

### 2024-06-26 workflow/scripts/compare_sampling_trees.R

@@ -8,18 +8,18 @@ taxidmap <- read_delim(snakemake@input[["taxidmap"]],
                        col_names = c("target", "Tax_ID"), show_col_types=F)
 
 
-tax <- read_delim(snakemake@input[["groups"]], show_col_types=F)
-# read input table
-lvls_tax <- rev(c(unique(tax$Clade), "archaea", "bacteria"))
-
-table_columns <- c('Proteome_ID', 'Tax_ID', 'count1', 'count2', 'count3', 
-                   'genome', 'source', 'species', 'mnemo')
+# tax <- read_delim(snakemake@input[["groups"]], show_col_types=F)
+# # read input table
+# lvls_tax <- rev(c(unique(tax$Clade), "archaea", "bacteria"))
+# 
+# table_columns <- c('Proteome_ID', 'Tax_ID', 'count1', 'count2', 'count3', 
+#                    'genome', 'source', 'species', 'mnemo')
 
 table <- read_delim(snakemake@input[["table"]], show_col_types=F, 
-                    delim = "\t", col_names = table_columns) %>% 
-  left_join(tax, by = "Tax_ID") %>% 
-  mutate(Clade = ifelse(is.na(Clade), mnemo, Clade),
-         Clade = factor(Clade, levels = lvls_tax))
+                    delim = "\t") 
+  # left_join(tax, by = "Tax_ID") %>% 
+  # mutate(Clade = ifelse(is.na(Clade), mnemo, Clade),
+  #        Clade = factor(Clade, levels = lvls_tax))
 
 
 # seeds <- readLines("results/test/ids/UP000005640_common.ids")

### 2024-06-26 workflow/scripts/compare_trees.R

@@ -1,7 +1,7 @@
 suppressPackageStartupMessages(library(tidyverse))
 suppressPackageStartupMessages(library(ggtree))
 suppressPackageStartupMessages(library(patchwork))
-suppressPackageStartupMessages(library(ggpubr))
+# suppressPackageStartupMessages(library(ggpubr))
 source("workflow/scripts/functions.R")
 
 theme_set(theme_classic())

### 2024-06-26 workflow/scripts/foldseek2distmat.py

@@ -1,94 +1,0 @@
-# import foldseek2tree
-import numpy as np
-import pandas as pd
-import argparse
-
-# Parse data
-parser = argparse.ArgumentParser(
-    description="Get a distance matrix from foldseek result"
-)
-
-parser.add_argument(
-    "-i",
-    "--input",
-    dest="foldseek",
-    action="store",
-    default="None",
-    help="Input foldseek result",
-    required=True
-)
-parser.add_argument(
-    "-o",
-    "--output",
-    dest="out",
-    nargs='+',
-    action="store",
-    default="None",
-    help="output files",
-    required=True
-)
-
-
-def distmat_to_txt( identifiers , distmat, outfile):
-	'''
-	write out a distance matrix in fastme format
-
-	Parameters
-	----------
-	identifiers : list
-		list of identifiers for your proteins
-	distmat : np.array  
-		distance matrix
-	outfile : str   
-		path to output file
-
-	'''
-
-	#write out distmat in phylip compatible format
-	outstr = str(len(identifiers)) + '\n'
-	for i,pdb in enumerate(identifiers):
-		outstr += pdb + ' ' + ' '.join( [ "{:.4f}".format(d) for d in list( distmat[i,: ] )  ]  ) + '\n'
-	with open(outfile , 'w') as handle:
-		handle.write(outstr)
-		handle.close()
-	return outfile
-
-args = parser.parse_args()
-
-if __name__=="__main__":
-    res = pd.read_table(args.foldseek, header = None)
-    # print(res.head())
-
-    #get the folder of the input file
-    infolder = args.foldseek.split('/')[:-1]
-    infolder = ''.join( [i + '/' for i in infolder])+'/'
-    # print(infolder)
-    res[0] = res[0].map(lambda x :x.replace('-model_v4.cif.gz', ''))
-    res[1] = res[1].map(lambda x :x.replace('-model_v4.cif.gz', ''))
-
-    res.columns = 'query,target,fident,lddt,alntmscore'.split(',')
-    # res.columns = 'query,target,fident,alnlen,mismatch,gapopen,qstart,qend,tstart,tend,evalue,bits,qcov,tcov,lddt,alntmscore,rmsd,prob'.split(',')
-    ids = list( set(list(res['query'].unique()) + list(res['target'].unique())))
-    pos = { protid : i for i,protid in enumerate(ids)}
-    kernels = ['fident', 'alntmscore', 'lddt']
-
[…]

### 2024-06-26 workflow/scripts/foldtree/compute_scores.py

@@ -1,0 +1,72 @@
+import treescore
+import toytree
+import pandas as pd
+import subprocess
+# import argparse
+import numpy as np
+
+from pathlib import Path
+
+
+T_score = 2000
+
+if __name__ == '__main__':
+    # args = parser.parse_args()
+    uniprot_df = pd.read_csv(snakemake.input[1], sep="\t", names=['query', 'mnemo','tax'])
+    taxidmap = pd.read_csv(snakemake.input[2], names=['query', 'species'], sep="\t")
+    # sptree = toytree.tree(snakemake.input[3])
+    # tmpfile = snakemake.output[0]
+
+    df = []
+    with open(snakemake.input[0]) as treefile:
+        for line in treefile:
+            line = line.strip().split()
+            gene = line[0]
+            targets = line[1]
+            alphabet = line[2]
+            model = line[3]
+            tree = toytree.tree(line[4], format = 0)
+            tree = treescore.label_leaves(tree, taxidmap, uniprot_df)
+            overlap = treescore.getTaxOverlap(tree.treenode)
+
+            leaves = tree.get_tip_labels()
+            n_sps = len(set(leaves))
+            n_tips = len(leaves)
+
+            # TCS score
+            taxscore = tree.treenode.score
+            # Branch lengths, normalized by total length
+            lengths = np.array([node.dist for node in tree.treenode.traverse()])
+            total_length = np.sum(lengths)
+            lengths /= total_length
+
+            # TCS score of root??
+            treescore.getTaxOverlap_root(tree.treenode)
+            root_score = treescore.sum_rootscore(tree.treenode)
+            # Tip to root distance
+            distances = np.array([ node.get_distance(tree.treenode) for node in tree.treenode.get_leaves() ])
+            # distances_norm = distances / np.mean(distances)
+
+            row = [gene, targets, alphabet, model, n_sps, n_tips, 
+                    taxscore, total_length, np.mean(lengths), np.mean(distances), 
+                    np.var(distances), root_score]
+            df.append(row)
+
+    outdf = pd.DataFrame(df, columns=['id', 'targets', 'alphabet', 'model', 'n_sps', 'n_tips',
+                                      'score', 'tree_length', 'mean_normalized_length', 'mean_r2t',
+                                      'variance_r2t', 'root_score'])
+
+    outdf.to_csv(snakemake.output[0], index=False, sep="\t")
+
+
[…]

### 2024-06-26 workflow/scripts/foldtree/corecut.py

@@ -1,0 +1,101 @@
+import pandas as pd
+import Bio.PDB
+import numpy as np
+import os
+import tqdm
+
+def extract_core(resdf, outfile,  hitthresh = .8, minthresh = .6, corefolder = 'core_structs/', structfolder = 'structs/', cterfolder = 'cter_structs/' , nterfolder = 'nter_structs/'):
+	"""
+
+	Extract a core of structures from a results file
+
+	Parameters
+		resdf: path to results file
+		outfile: path to output file
+		hitthresh: proportion of structures that need to map to a residue for it to be included in the core
+		minthresh: if no residues meet the hitthresh, the minimum proportion of structures that need to map to a residue for it to be included in the core
+		corefolder: name of folder to output core structures to
+		structfolder: name of folder to find structures in
+		cterfolder: name of folder to find cter structures in
+		nterfolder: name of folder to find nter structures in
+
+
+	"""
+	
+	#read all results
+	# folder =''.join([ sub + '/' for sub in resdf.split('/')[:-1] ])
+	# print(folder)
+	resdf = pd.read_table(resdf, header = None)
+	resdf.columns = 'query,target,fident,alnlen,mismatch,gapopen,qstart,qend,tstart,tend,evalue,bits,lddt,lddtfull,alntmscore'.split(',')
+
+	print('extracting core')
+	print('hitthresh: ' + str(hitthresh))
+	print('minthresh: ' + str(minthresh))
+
+	queries = resdf['query'].unique()
+	nqueries = len(queries)
+	print(resdf.head(), nqueries)
+	#map hits to each struc
+	hits = {}
+	#proportion of structures that need to map to a residue fo
+	with tqdm.tqdm(total=nqueries) as pbar:
+		for i,q in enumerate(queries):
+			sub = resdf[resdf['query'] == q]
+			hitvec = np.zeros((1, max(sub['qend'])) )
+			for idx,r in sub.iterrows():
+				hitvec[0,r['qstart']:r['qend']] = hitvec[0,r['qstart']:r['qend']]+1
+			hitvec /= nqueries
+			core = np.where(hitvec>hitthresh)[1]
+			try:
+				hits[q]= {'min': np.amin(core), 'max': np.amax(core)}
+			except:
+				#be more lenient...
+				subthresh = np.amax(hitvec)
[…]

### 2024-06-26 workflow/scripts/foldtree/foldseek2tree.py

@@ -1,0 +1,243 @@
+#foldssek cluster
+import subprocess ,shlex
+import numpy as np
+from scipy.spatial.distance import cdist
+import statsmodels
+import toytree
+import pandas as pd
+import re
+import os
+from scipy.stats import chi2
+import argparse
+
+def consensustree(treelist):
+	'''get a consensus tree from a list of tree files
+	
+	Parameters
+	----------
+	treelist : list
+		list of tree files
+
+	'''
+	#creat a multitree object from the list of tree files
+	treelist = [toytree.tree(i , format = 0 ) for i in treelist]
+	mt = toytree.mtree(treelist)
+	#get the consensus tree
+	ct = mt.get_consensus_tree( )
+	return ct
+
+#smooth distmat with MDS
+def MDS_smooth(distmat):
+	mds = MDS(n_components=int(distmat.shape[0]/2) )#, dissimilarity='precomputed'  )
+	distmat = mds.fit_transform(1-distmat )
+	distmat = cdist(distmat,distmat, 'minkowski', p=1.5 )
+	return distmat
+
+
+
+def Tajima_dist(kn_ratio, bfactor=1, iter = 100, delta=10**-5):
+	#approx of b*log(1-p/b) from https://doi.org/10.1093/oxfordjournals.molbev.a040031
+	#set all entries to smaller than bfactor
+	kn_ratio[kn_ratio > bfactor] = bfactor - delta
+	taj = np.add.reduce([ (kn_ratio**(np.ones(kn_ratio.shape)*i) )/ (bfactor**(i-1)*i) for i in range(1,iter) ] )
+	#set diagonal to 0
+	np.fill_diagonal(taj, 0)
+	return taj
+
+
+def runargs(args):
+	'''run a command line command
+	
+	Parameters
+	----------
+	args : str
+		command line command
+	'''
+	
+	args = shlex.split( args)
+	p = subprocess.run( args )
+	return p
+	
+def runFoldseekdb(folder , outfolder , foldseekpath = 'foldseek'):
+	'''run foldseek createdb
+	
+	parameters
+	----------
+	folder : str
+		path to folder with pdb files
+	outfolder : str 
+		path to output folder
+	
+
+	'''
+	args = foldseekpath + ' createdb '+  folder + ' '+ outfolder+'structblobDB '
+	p = runargs(args)
+	return outfolder+'structblobDB '
+
[…]

### 2024-06-26 workflow/scripts/foldtree/foldseekres2distmat_simple.py

@@ -1,0 +1,61 @@
+import numpy as np
+import pandas as pd
+
+def distmat_to_txt(identifiers, distmat, outfile):
+	'''
+	write out a distance matrix in fastme format
+
+	Parameters
+	----------
+	identifiers : list
+		list of identifiers for your proteins
+	distmat : np.array  
+		distance matrix
+	outfile : str   
+		path to output file
+
+	'''
+
+	#write out distmat in phylip compatible format
+	outstr = str(len(identifiers)) + '\n'
+	for i,pdb in enumerate(identifiers):
+		outstr += pdb + ' ' + ' '.join( [ "{:.4f}".format(d) for d in list( distmat[i,: ] )  ]  ) + '\n'
+	with open(outfile , 'w') as handle:
+		handle.write(outstr)
+		handle.close()
+	return outfile
+
+
+res = pd.read_table(snakemake.input[0], header = None)
+# print(res.head())
+#get the folder of the input file
+infolder = snakemake.input[0].split('/')[:-1]
+infolder = ''.join( [i + '/' for i in infolder])+'/'
+res[0] = res[0].map(lambda x :x.replace('.cif', ''))
+res[1] = res[1].map(lambda x :x.replace('.cif', ''))
+res.columns = 'query,target,fident,lddt,alntmscore'.split(',')
+ids = list( set(list(res['query'].unique()) + list(res['target'].unique())))
+pos = { protid : i for i,protid in enumerate(ids)}
+kernels = ['fident']#, 'alntmscore', 'lddt']
+
+#set kernel columns to float
+for k in kernels:
+    res[k] = res[k].astype(float)
+#change nan to 0
+res = res.fillna(0)
+matrices = { k:np.zeros((len(pos), len(pos))) for k in kernels }
+# print(res)
+
+#calc kernel for tm, aln score, lddt
+for idx,row in res.iterrows():
+    for k in matrices:
+        matrices[k][pos[row['query']] , pos[row['target']]] += row[k]
+        matrices[k][pos[row['target']] , pos[row['query']]] += row[k]
+
+for i,k in enumerate(matrices):
+    matrices[k] /= 2
+    matrices[k] = 1-matrices[k]
+    # print(matrices[k], np.amax(matrices[k]), np.amin(matrices[k]) )
+    # np.save( infolder + k + '_distmat.npy' , matrices[k])
+    distmat_txt = distmat_to_txt(ids, matrices[k], snakemake.output[i])
+

### 2024-06-26 workflow/scripts/foldtree/foldtree.py

@@ -1,0 +1,138 @@
+
+import foldseek2tree
+import corecut 
+import argparse
+import numpy as np
+import pandas as pd
+import os
+
+def structblob2tree(input_folder, outfolder, tmpfolder, corefolder, outtree, 
+                    overwrite = False, correction = False,
+                    fastmepath = 'fastme', foldseekpath = 'foldseek', delta = 0.0001,
+                    kernel = 'fident', core = False, hitthresh = .8, minthresh = .6):
+    '''
+    run fold tree pipeline for a folder of pdb files
+    
+    Parameters
+    ----------
+    input_folder : str
+        path to folder with pdb files   
+    outfolder : str
+        path to output folder   
+    overwrite : bool
+        overwrite existing foldseek output  
+    fastmepath : str    
+        path to fastme executable
+    quicktreepath : str 
+        path to quicktree executable
+    foldseekpath : str  
+        path to foldseek executable 
+    delta : float   
+        small number to replace negative branch lengths with, default is .0001
+    correction : str    
+        correction method to use, either 'tajima' or 'none'
+    kernel : str    
+        kernel to use, either 'fident', 'lddt' or 'alntmscore'
+    
+    '''
+    #check if the foldseek output is already there
+    input_folder = os.path.join(input_folder, '')
+    if os.path.exists(outfolder + '_allvall.tsv') and overwrite == False:
+        print('found foldseek output, skipping foldseek')
+        alnres = outfolder + '_allvall.tsv'
+    else:
+        alnres = foldseek2tree.runFoldseek_allvall_EZsearch(input_folder, outfolder + '_allvall.tsv', tmpfolder, foldseekpath = foldseekpath)
+    
+    if core == True:
+        corefolder = os.path.join(corefolder, '')
+        corecut.extract_core(alnres, outfolder+'.core',  hitthresh = hitthresh, minthresh = minthresh, 
+                             corefolder = corefolder, structfolder = input_folder)
+        if os.path.exists(outfolder + '_core_allvall.tsv') and overwrite == False:
+            print('found foldseek core output, skipping foldseek')
+            alnres = outfolder + '_core_allvall.tsv'
+        else:
+            alnres = foldseek2tree.runFoldseek_allvall_EZsearch(corefolder, outfolder + '_core_allvall.tsv', tmpfolder, foldseekpath = foldseekpath)
+
[…]

### 2024-06-26 workflow/scripts/foldtree/treescore.py

@@ -1,0 +1,261 @@
+import toytree
+# import toyplot
+
+#taxonomy overlap score
+
+def standard_treedraw( tre, sizes= None , colors= None ,fixed_order=None, fixed_position=None , 
+ts = None,  save_file = False  , tiplabels = None ,  layout='c', edge_type='p' ):
+	
+	'''
+	standard treedraw function
+	
+	Parameters
+	----------
+	tre : toytree.tree
+		tree to draw
+	sizes : list
+		list of node sizes
+	colors : list
+		list of node colors
+	fixed_order : list
+		list of node names in order to draw
+	fixed_position : list
+		list of node positions
+	ts : treestlye : str
+		style of tree to draw (default = 'c')
+	save_file : str
+		path to save file
+	tiplabels : list
+		list of tip labels to draw
+	layout : str
+		layout of tree to draw (default = 'c')
+	edge_type : str
+		type of edge to draw (default = 'p')
+	'''
+
+	
+	if tiplabels is None:
+		tiplabels = tre.get_tip_labels()
+	canvas, axes, mark = tre.draw(  
+		ts = ts,
+		node_sizes=sizes,
+		node_colors=colors,
+		tip_labels_align=True,
+		scalebar=True,
+		fixed_order=fixed_order, 
+		fixed_position=fixed_position,
+		tip_labels=tiplabels,
+		layout=layout,
+		edge_type=edge_type,
+		tip_labels_style={
+			"fill": "#262626",
+			"font-size": "9px"}
+	)
+	if save_file:
+		toyplot.svg.render(canvas, save_file)
+
+
+
+def exp_score( v , exp = 1.5):
+	return v**exp
+
+def frac_score( v ):
+	return v+1
+
+#taxonomy overlap score
+def getTaxOverlap(node , treelen  = None , scorefun = frac_score):
+	
+	"""
+    Calculate the taxonomy overlap score for the given node in a phylogenetic tree.
+    
+    The taxonomy overlap score is defined as the number of taxonomic labels shared by all the leaf nodes
+    descended from the given node, plus the sum of the scores of all its children. If a leaf node has no
+    taxonomic label, it is not counted towards the score. The function also calculates the size of the
+    largest loss in lineage length, defined as the difference between the length of the set of taxonomic
+    labels shared by all the leaf nodes and the length of the longest set of taxonomic labels among the
+    children of the node.
+    
[…]

### 2024-06-26 workflow/scripts/get_yamls.sh

@@ -1,0 +1,9 @@
+
+#!/usr/bin/bash
+
+declare -a envs=("sp_utils" "sp_R" "sp_homology" "sp_tree" "sp_python")
+
+for env in "${envs[@]}"; do
+    echo $env
+    conda env export -n $env --from-history | grep -E -v "^name:|^prefix:" > workflow/envs/${env}.yaml
+done

### 2024-06-26 workflow/scripts/mask_structures.py

@@ -1,33 +1,17 @@
 #!/usr/bin/env python
-import argparse
 from Bio import SeqIO
 import pandas as pd
 import glob
 
 
-# Parse data
-parser = argparse.ArgumentParser(
-    description="Mask a fasta file based on structural confidence prediction. Residues with low lddt will be Xs"
-)
-
-parser.add_argument("-i", "--in", dest="infile", default="None", 
-                    help="input file",required=True)
-parser.add_argument("-o", "--out", dest="out", default="None", 
-                    help="output file",required=True)
-parser.add_argument("-s", "--struct", dest="struct", default="None", 
-                    help="structure directory",required=True)
-parser.add_argument('-m', '--min', dest='min',
-                    help='Minimum lddt', type=int, default=50)
-
 if __name__ == '__main__':
-    args = parser.parse_args()
-    min_lddt = args.min
+    min_lddt = snakemake.params[1]
 
     out_seqs = []
 
-    for record in SeqIO.parse(args.infile, "fasta"):
+    for record in SeqIO.parse(snakemake.input[0], "fasta"):
         gene_id = record.name
-        json_conf = glob.glob(args.struct+'/*/confidence/'+gene_id+'-confidence_v4.json.gz')[0]
+        json_conf = glob.glob(snakemake.params[0]+'/*/confidence/'+gene_id+'-confidence_v4.json.gz')[0]
 
         confidence = pd.read_json(json_conf)
         low_confidence_sites = list(confidence[confidence['confidenceScore']<min_lddt]['residueNumber'])
@@ -37,4 +21,4 @@ if __name__ == '__main__':
 
         out_seqs.append(record)
 
-    SeqIO.write(out_seqs, args.out, "fasta")
+    SeqIO.write(out_seqs, snakemake.output[0], "fasta")

### 2024-06-26 workflow/scripts/overlap.R

@@ -1,50 +1,0 @@
-source("scripts/functions.R")
-theme_set(theme_bw())
-opt$blast <- "results/Hsap_opistho//homology/UP000005640_blast.tsv"
-opt$fs <- "results/Hsap_opistho/homology/UP000005640_fs.tsv"
-
-fs <- read_delim(opt$fs, col_names = fs_columns, show_col_types = FALSE) %>% 
-  mutate(evalue = ifelse(evalue==0, 1e-180, evalue)) %>%
-  filter(query!=target) %>% 
-  filter(evalue<1e-3)
-  # left_join(taxidmap)
-blast <- read_delim(opt$blast, col_names = blast_columns, show_col_types = FALSE) %>% 
-  mutate(evalue = ifelse(evalue==0, 1e-180, evalue)) %>%
-  filter(query!=target) %>% 
-  filter(evalue<1e-3)
-
-df_overlap <- tibble()
-
-min_hits <- c(seq(50, 900, 100), 1000)
-
-for (hits in min_hits){
-      a <- group_by(blast, query) %>% 
-        arrange(desc(bitscore)) %>% 
-        slice_head(n = hits)
-      b <- group_by(fs, query) %>% 
-        arrange(desc(bitscore)) %>% 
-        slice_head(n = hits)
-
-      to_add <- full_join(select(a, query, target, pident, evalue, bitscore, qcov),
-                          select(b, query, target, pident, evalue, bitscore, qcov),
-                          by = c("query", "target"), suffix=c("_blast", "_fs")) %>% 
-        mutate(singleton = case_when(is.na(evalue_fs) ~ "only_blast",
-                                     is.na(evalue_blast) ~ "only_fs", .default = "common")) %>%
-        group_by(query, singleton) %>% 
-        count() %>%
-        ungroup(singleton) %>% 
-        mutate(n_tot = n/sum(n), n_hits = hits)
-
-      df_overlap <- bind_rows(df_overlap, to_add)
-}
-
-
-df_overlap %>%
-  group_by(n_hits, singleton) %>%
-  summarize(mn=median(n_tot)) %>%
-  ggplot(aes(n_hits, mn, color=singleton)) + 
-  geom_point() + 
-  # geom_smooth(method = lm, formula = y ~ splines::bs(x, 3)) +
-  geom_vline(xintercept = 150) +
-  geom_line(aes(group=singleton))
-  # scale_y_continuous(limits = c(0,1))

### 2024-06-26 workflow/scripts/temp_BRH.R

@@ -1,10 +1,0 @@
-suppressPackageStartupMessages(library(ggtree))
-suppressPackageStartupMessages(library(tidyverse))
-source("workflow/scripts/functions.R")
-
-theme_set(theme_classic())
-
-# order of models
-models <- c("QT", "FT", "LG", "GTR", "3Di")
-
-

### 2024-07-12 workflow/notebooks/draft.Rmd
fixed apro

@@
   scale_y_continuous(limits = c(0, 100)) +
+  labs(x="Identity Blast",y="Identity Foldseek") +
   theme(legend.position = "none", panel.grid = element_blank())
@@
   scale_y_continuous(limits = c(0, 100)) +
+  labs(x="Query cov. Blast",y="Query cov. Foldseek") +
   scale_fill_gradientn(colors=wespal) +
@@
   #                    expand = c(0,0)) +
-  scale_color_manual(values = palette_singleton_p)+ 
+  scale_color_manual(values = palette_singleton_p) + 
+  labs(x="LDDT") + 
   theme(legend.position = "none", 
@@
   # coord_cartesian(expand = 0) +
+  labs(x="TM-Score") + 
   scale_color_manual(values = palette_singleton_p)+ 
@@
   scale_x_continuous(limits = c(params$low_confidence, 100)) +
-  scale_color_manual(values = palette_singleton_p)+ 
+  scale_color_manual(values = palette_singleton_p) +   
+  labs(x="Mean target pLDDT") + 
   theme(legend.position = "none", 
@@
     mutate(name=gsub("Jaccard_", "", name),
-           name=gsub("3d", "CATH", name)) %>% 
-    ggplot(aes(Jaccard, fill=singleton)) +
-    geom_histogram(binwidth = 0.1, position = "dodge") +
+           name=gsub("3d", "CATH", name),
+           Jaccard=cut_width(Jaccard, .1, boundary = 0)) %>% 
+  filter(!is.na(Jaccard)) %>% 
+  group_by(name, singleton, Jaccard) %>% 
+  summarise(n=n()) %>% 
+  mutate(freq=n/sum(n)) %>%
+  ggplot(aes(Jaccard, freq, fill=singleton)) + 
+  geom_bar(stat = "identity",position = "dodge") + 
   facet_grid(name~.) +
-    scale_x_continuous(breaks = seq(0,1,0.1)) +
-    scale_y_continuous(expand = expansion(0,0)) + 
+  labs(y="% in group") +
+  scale_y_continuous(expand = expansion(0.01,0)) + 
   scale_fill_manual(values=palette_singleton_p) +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom",
+        panel.background = element_rect(fill=NA, color="black"),
+        axis.line = element_blank(),
+        axis.text.x = element_text(size=8))
 
@@
 fig2_second_row <- ((lddt_p / alntmscore_p / plddt_p) | CATH_plot) + plot_layout(widths = c(1,2))
-fig2 <-  (fig2_first_row / fig2_second_row)
+fig2 <-  (fig2_first_row / fig2_second_row) +
+  plot_annotation(tag_levels = 'A') & 
+  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
 
-ggsave("draft/figures/fig2.pdf", fig2, width = 9, height = 6)
+ggsave("draft/figures/fig2.pdf", fig2, width = 10.5, height = 8)
 ```
@@
 plot_bs <- df_bs %>% 
+  mutate(targets=target) %>% 
   # group_by(gene, target, model) %>% 
   # summarise(mn_bs=mean(support)) %>% 
-  ggplot(aes(model, support, fill=target)) +
+  ggplot(aes(model, support, fill=targets)) +
   # geom_violin(position = position_dodge(width = 1)) +
-  geom_boxplot() +
+  geom_boxplot(outlier.size = .1) +
   # ggdist::stat_pointinterval(position = position_dodge(width = 1), color="black") +
@@
   theme_classic() +
-  labs(y="Boostrap", x="") +
-  theme(legend.position = "none") 
+  labs(y="Boostrap", x="") 
+  # theme(legend.position = "bottom") 
   # ggpubr::stat_compare_means(aes(group = model),
@@
   scale_x_continuous(breaks = c(0,.25,.5, .75,1), labels = function(x) sprintf("%.2g", x)) +
-  coord_fixed(xlim = c(0,1), expand=0) +
+  coord_cartesian(xlim = c(0,1), expand=0) +
   # facet_grid(method~method2, switch = "both") +
@@
         axis.ticks.y = element_blank(),
-        axis.title.y = element_blank())
+        axis.title.y = element_blank(),
+        legend.margin=margin(0,0,0,0),
+        legend.box.margin=margin(-20,0,0,0))
 
@@
   ggplot(aes(model, n_events, fill=targets)) +
-  geom_boxplot() +
+  geom_boxplot(outlier.size = .1) +
   labs(x="", y="(D+L)/# Tips") +
   # ggthemes::scale_fill_solarized() +
-  scale_fill_manual(values = palettes_method) +
-  theme(legend.position = "none")
+  scale_fill_manual(values = palettes_method)
+  # theme(legend.position = "bottom")
 
@@
   # geom_histogram(position = "dodge") + 
-  coord_cartesian(expand = 0) +
+  # coord_cartesian(expand = 0) +
   scale_x_continuous(limits = c(0,.2)) + 
@@
   scale_color_manual(values = palettes_model) +
-  theme(legend.position = "bottom")
+  theme(legend.position = 'bottom',
+        legend.margin=margin(0,0,0,0),
+        legend.box.margin=margin(-10,0,0,0))
 
-fig3 <- rf_plot | ((plot_bs | plot_DL) / var_r2t)
+secondcolumn <- (plot_DL + plot_bs + 
+    plot_layout(guides = "collect", widths = c(5,3)) &
+    theme(legend.position = 'bottom',
+          legend.margin=margin(0,0,0,0),
+          legend.box.margin=margin(-20,0,0,0))) / var_r2t
+
+fig3 <- (rf_plot | secondcolumn) +
+  plot_annotation(tag_levels = 'A') & 
+  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
 ggsave("draft/figures/fig3.pdf", fig3, width = 11, height = 6)
@@
 
+nodes_df %>% 
+  ggplot(aes(label_int, f1+f2+f3, fill=model)) + 
+  geom_bar(stat="identity", position = "dodge") +
+  facet_grid(targets~.) +
+  scale_fill_manual(values=palettes_model)
+
 sptree_plot <- ggtree(labeled_sptree) + 
@@
   geom_nodelab(color="firebrick", 
-               nudge_x = -.1, hjust = 1, 
-               nudge_y = .2, size=2)+
+               nudge_x = -.15, hjust = 1, 
+               nudge_y = .3, size=2)+
   scale_x_continuous(expand = c(.2,.2))
 
-fig_s1 <- (plot_dist + sptree_plot + pp1_plot) + plot_layout(widths = c(1.4, 1, 1.6))
+plot_ntips <- scores %>% 
+  filter(model!="FTPY") %>% 
+  ggplot(aes(n_tips, dups+losses, color=targets)) + 
+  geom_point(size=.2) +
+  geom_smooth(method = "lm") +
+  facet_grid(~model) + 
+  scale_color_manual(values = palettes_method) +
+  theme_classic() + 
+  theme(panel.border = element_rect(fill=NA),
+        legend.position = "bottom")
 
+fig_s1 <- (plot_dist + sptree_plot + pp1_plot) +
+  plot_layout(widths = c(1.4, 1, 1.6)) + 
+  plot_annotation(tag_levels = 'A') & 
+  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
 ggsave("draft/figures/figS1.pdf", fig_s1, width = 14, height = 6)
+
+# fig_s1 <- (plot_ntips / ((plot_dist + sptree_plot + pp1_plot) +
+#                            plot_layout(widths = c(1.4, 1, 1.6)))) + 
+#   plot_layout(heights = c(.75,1)) 
+# ggsave("draft/figures/figS1.pdf", fig_s1, width = 14, height = 8)
 ```
+
+## Answers to Edu's comments
+
+```{r}
+df %>% 
+  filter(query %in% seeds) %>% 
+  count(query, singleton) %>% 
+  ggplot(aes(n, singleton, fill=singleton)) +
+  geom_boxplot()
+
+blast_filtered_eval <- filter(blast, 
+                         query %in% paste0("AF-", aln_seeds, "-F1"),
+                         evalue<as.numeric(params$eval_both)) %>% 
+  group_by(query) %>% 
+  slice_head(n = params$max_seqs) %>% 
+  ungroup()
+
+fs_filtered_eval <- filter(fs, 
+                      query %in% paste0("AF-", aln_seeds, "-F1"),
+                      evalue<as.numeric(params$eval_both)) %>% 
+  group_by(query) %>% 
+  slice_head(n = params$max_seqs) %>% 
+  ungroup()
+
+df_filtered_eval <- full_join(blast_filtered_eval, fs_filtered_eval,
+                by = c("query", "target"), suffix=c("_Blast", "_Foldseek")) %>%
+    filter(query!=target) %>% 
+  left_join(self, by="query") %>%
+  mutate(pident_Foldseek = 100*pident_Foldseek, 
+         qcov_Foldseek = 100*qcov_Foldseek,
+         singleton = case_when(is.na(evalue_Foldseek) ~ "only Blast",
+                               is.na(evalue_Blast) ~ "only Foldseek", 
+                               .default = "Common"))
+
+select(df_filtered_eval, query, target, singleton) %>% 
+  mutate(method="only-eval") %>% 
+  rbind(select(df_filtered, query, target, singleton) %>% 
+          mutate(method="qcov-tcov")) %>% 
+  group_by(method, singleton) %>% 
+  summarise(n = n()) %>%
+  mutate(freq = n / sum(n))
+
+res %>% 
+  group_by(singleton, model) %>% 
+  summarise(median(nn_dist/max_dist))
+```

### 2024-07-12 workflow/envs/sp_python.post-deploy.sh

@@ -6,4 +6,4 @@ cd build
 cmake -DUserInstallOption=ON ../
 make install
 
-cp -r build/lib/pdbx/ $CONDA_PREFIX/lib/python*/
+cp -r build/lib/pdbx/ $CONDA_PREFIX/lib/python3.10/

### 2024-07-12 workflow/envs/sp_tree.post-deploy.sh

@@ -6,7 +6,8 @@ cp bin/astral-pro $CONDA_PREFIX/bin/
 cd -
 
 # install Ranger-DTL
-# wget https://compbio.mit.edu/ranger-dtl/ranger-dtl-linux.tar.gz -O $CONDA_PREFIX/lib/RangerDTL.tar.gz
-# tar xvzf $CONDA_PREFIX/lib/RangerDTL.tar.gz -C $CONDA_PREFIX/bin/ #--strip-components=2 ./bin/foldmason
+wget https://compbio.engr.uconn.edu/wp-content/uploads/sites/2447/2019/08/Linux.zip -O $CONDA_PREFIX/lib/RangerDTL.zip
+unzip $CONDA_PREFIX/lib/RangerDTL.zip -d $CONDA_PREFIX/lib/RangerDTL
+chmod +x $CONDA_PREFIX/lib/RangerDTL/Linux/CorePrograms/*
+mv $CONDA_PREFIX/lib/RangerDTL/Linux/CorePrograms/* $CONDA_PREFIX/bin/
 
-# 404 error. let's wait and see if they fix it, for now use one already installed.

### 2024-07-12 workflow/scripts/analyze_apro.R

@@ -19,22 +19,25 @@ apro_trees <- read.tree(text = sapply(apro_files, readLines))
 names(apro_trees) <- sapply(str_split(basename(apro_files), "_"), function(x) paste0(x[2], "_", x[4]))
 
 trees_df <- fortify(apro_trees) 
-nodes_df <- trees_df %>%   group_by(.id) %>% 
+nodes_df <- trees_df %>%
+  group_by(.id) %>% 
   mutate(ordered = rank(y)) %>% 
   left_join(nodes_clades) %>% 
   separate(.id, c("targets", "model")) %>%  
   filter(!isTip, label!="") %>% 
   mutate(label = gsub("\\[|\\]|\\'", "", str_replace_all(label, "[a-z]{1,2}[0-9]=", ""))) %>% 
-  separate(label, c("pp1", "pp2", "pp3", "f1", "f2", "f3", "q1", "q2", "q3"), ";", convert = TRUE) %>% 
+  separate(label,
+           c("pp1", "pp2", "pp3", "f1", "f2", "f3", "q1", "q2", "q3"), ";", 
+           convert = TRUE) %>% 
   left_join(nodes_clades) %>% 
   mutate(freq=f1+f2+f3)
 
 
 pp1_plot <- nodes_df %>% 
-    ggplot(aes(q1, label, fill=model, color=model, group=model)) +
+    ggplot(aes(q1, label_int, fill=model, color=model, group=model)) +
     geom_linerange(aes(x=x, xmin=xmin, xmax=xmax, color=model),
                    data = . %>% 
-                     group_by(model, label) %>% 
+                     group_by(model, label_int) %>% 
                      summarise(x=mean(q1), xmin=min(q1), xmax=max(q1)),
                    position = position_dodge(width=.7), color="grey70"
                    ) +

### 2024-07-12 workflow/scripts/analyze_runtimes.R

@@ -7,7 +7,7 @@ df <- read_delim(snakemake@input[["time"]],
                  show_col_types = FALSE) %>% 
   mutate(basename=gsub(".txt", "", basename),
          step=basename(dirname),
-         mutate(max_rss_gb = max_rss/1024)) %>% 
+         max_rss_gb = max_rss/1024) %>% 
   separate(basename, c("seed", "gene", "method", "alphabet", "model"))
 
 

### 2024-07-12 workflow/scripts/foldtree/compute_scores.py

@@ -25,7 +25,7 @@ if __name__ == '__main__':
             targets = line[1]
             alphabet = line[2]
             model = line[3]
-            tree = toytree.tree(line[4], format = 0)
+            tree = toytree.tree(line[4], tree_format = 0)
             tree = treescore.label_leaves(tree, taxidmap, uniprot_df)
             overlap = treescore.getTaxOverlap(tree.treenode)
 

### 2024-07-12 workflow/scripts/plot_example.R

@@ -8,13 +8,13 @@ theme_set(theme_classic())
 
 seeds <- readLines(snakemake@input[["ids"]])
 
-seed_sp <- gsub("_examples.ids", "", basename(snakemake@input[["ids"]]))
+seed_sp <- snakemake@params[["seed"]]
 seed_dir <- gsub("trees", "seeds", dirname(snakemake@input[["trees"]]))
 
 taxmap <- read_delim(snakemake@input[["taxidmap"]], 
                      delim = "\t", col_names = c("label", "Tax_ID"),
                      show_col_types = FALSE) %>% 
-  left_join(read_delim(snakemake@input[["meta"]],
+  left_join(read_delim(snakemake@input[["table"]],
                        show_col_types = FALSE), by = "Tax_ID") %>% 
   dplyr::select(label, Proteome_ID, mnemo)
 
@@ -42,7 +42,7 @@ plot_domains = list()
 
 for (seed in seeds) {
   print(seed)
-  fls <- list.files(paste0("results/test/seeds/", seed_sp, "/", seed), full.names = TRUE)
+  fls <- list.files(paste0(seed_dir, "/", seed_sp, "/", seed), full.names = TRUE)
   ids <- fls[grepl("(blast|fs).*ids$", fls)]
 
   x <- sapply(ids, readLines, simplify = FALSE)
@@ -70,7 +70,7 @@ for (seed in seeds) {
 
   conf_df <- tibble(label=union(x$blast, x$fs)) %>% 
     left_join(taxmap, by="label") %>% 
-    mutate(confidence_file=paste0("results/data/structures/",
+    mutate(confidence_file=paste0(snakemake@params[["struct_dir"]],
                                   Proteome_ID,
                                   "/confidence/",
                                   label,
@@ -122,7 +122,8 @@ for (seed in seeds) {
     summarise(mn=mean(conf)) %>% 
     mutate(scaled_mn=length(names(seqs))*mn/100)
   
-  plot_3di_aln <- ggmsa(aln_3di_trimmed, seq_name = FALSE, font = NULL, show.legend = TRUE, border = NA) +
+  plot_3di_aln <- ggmsa(aln_3di_trimmed, seq_name = FALSE, 
+                        font = NULL, show.legend = TRUE, border = NA) +
     labs(title = paste(seed, "- 3Di union hits cleaned aln")) + 
     geom_line(mapping = aes(x=pos_trim, y=scaled_mn), data = scaled_conf_df, inherit.aes = F) +
     coord_cartesian() + 

### 2024-07-22 workflow/notebooks/draft.Rmd
added draft.rmd

@@
 
+blast_filt <- filter(blast, 
+                         # query %in% paste0("AF-", aln_seeds, "-F1"),
+                         length/qlen*100>params$coverage, 
+                         length/slen*100>params$coverage,
+                         evalue<as.numeric(params$eval_both)) %>% 
+  group_by(query) %>% 
+  slice_head(n = params$max_seqs) %>% 
+  ungroup()
+
+fs_filt <- filter(fs, 
+                      # query %in% paste0("AF-", aln_seeds, "-F1"),
+                      qcov*100>params$coverage, 
+                      tcov*100>params$coverage, 
+                      evalue<as.numeric(params$eval_both)) %>% 
+  group_by(query) %>% 
+  slice_head(n = params$max_seqs) %>% 
+  ungroup()
+
+df_filt <- full_join(blast_filt, fs_filt,
+                by = c("query", "target"), suffix=c("_Blast", "_Foldseek")) %>%
+    filter(query!=target) %>% 
+  left_join(self, by="query") %>%
+  mutate(pident_Foldseek = 100*pident_Foldseek, 
+         qcov_Foldseek = 100*qcov_Foldseek,
+         singleton = case_when(is.na(evalue_Foldseek) ~ "only Blast",
+                               is.na(evalue_Blast) ~ "only Foldseek", 
+                               .default = "Common"))
 # grouped_brhs <- brhs %>%
@@
     distinct()
+```
+
+## Fig1
+
+```{r fig1}
+
+fig1a <- df %>%
+  group_by(singleton) %>%
+  summarise(n = n()) %>%
+  ungroup() %>%
+  mutate(prop=n/sum(n), Filtering="Before") %>%
+  bind_rows(df_filt %>%
+            group_by(singleton) %>%
+            summarise(n = n()) %>%
+            ungroup() %>%
+            mutate(prop=n/sum(n), Filtering="After")) %>%
+  mutate(Filtering=factor(Filtering, labels = c("After", "Before"), ordered = T),
+         singleton=case_when(singleton=="only Blast"~"Only Bp", 
+                             singleton=="only Foldseek" ~ "Only Fs",
+                             .default = singleton)) %>%
+  ggplot(aes(singleton, n, fill=singleton)) +
+  geom_bar(aes(alpha=Filtering), stat = "identity", position = position_dodge2(reverse=TRUE)) +
+  geom_text(aes(group=Filtering, label=paste0(round(prop, 2)*100, "%")),
+            position = position_dodge2(reverse = TRUE, width = 0.9),
+            vjust=1.3, color="white", size=3) +
+  scale_fill_manual(values = palette_singleton_s, guide="none") +
+  scale_alpha_manual(values = c(1,.6)) +
+  scale_y_continuous(expand = c(0,0), labels = scales::label_number(scale = 1e-6)) +
+  labs(fill="Class", x="Class", y="# aligned protein pairs (Mln)") +
+  theme(legend.position = "bottom", legend.box="vertical")
+
+fig1b <- lddt_df %>% 
+  ggplot(aes(mean_lddt)) + 
+  geom_density(fill="#BF4375", size=1) +
+  scale_y_continuous(expand = expansion(mult = c(0, .1))) +
+  labs(x="Mean pLDDT", y="") +
+  geom_vline(xintercept = 40, linetype=2) +
+  theme(axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.line.y = element_blank())
+fig1c <- tibble(seeds) %>% 
+  mutate(`Has > 4 common` = seeds %in% paste0("AF-",aln_seeds,"-F1")) %>% 
+  ggplot(aes(fill=`Has > 4 common`, y=1)) + 
+  geom_bar() +
+  coord_cartesian(expand = 0) +
+  scale_fill_manual(values = c("grey80", "black")) +
+  labs(x="Has > 4 Common hits") +
+  theme(legend.position = "none", 
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.line.y = element_blank())
+
+sps <- read_delim("data/input/Hsapopi_set_sp.txt")
+
+sptree_plot <- fortify(sptree) %>% 
+  left_join(sps, by=c("label"="mnemo")) %>% 
+  mutate(species=gsub("[^A-Z].+? ", ". ", species)) %>% 
+  ggtree() +
+  geom_tiplab(aes(label=species), nudge_x=.2, fontface="italic") + 
+  geom_nodelab(color="firebrick", 
+               nudge_x = -.15, hjust = 1, 
+               nudge_y = .3, size=2.5)+
+  scale_x_continuous(expand = expansion(mult = c(0.2, .7)))
+
+ggsave("draft/figures/Fig1/tree.pdf", sptree_plot, width = 4, height = 4)
+
+fig1 <- (fig1b | (fig1c/fig1d + plot_layout(heights = c(10,1)))) +
+#   plot_layout(widths = c(1,2)) + 
+#   plot_annotation(tag_levels = 'A') & 
+#   theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
+ggsave("draft/figures/Fig1/fig1a.pdf", fig1a, width = 3, height = 4)
+ggsave("draft/figures/Fig1/fig1b.pdf", fig1b, width = 4, height = 2)
+ggsave("draft/figures/Fig1/fig1c.pdf", fig1c, width = 4, height = 1)
+
 ```
@@
   filter(model!="FTPY", targets=="common") %>% 
-  ggplot(aes(color=model, variance_r2t)) + 
-  geom_density(size=1, alpha=.5) +
+  ggplot(aes(fill=model, variance_r2t)) + 
+  # geom_density(size=1, alpha=.5) +
+  geom_histogram(position = "dodge") +
     # geom_histogram(position = "dodge") + 
   # coord_cartesian(expand = 0) +
-  scale_x_continuous(limits = c(0,.2)) + 
-  # facet_grid(~targets) + 
+  scale_x_continuous(limits = c(0,.2), expand = expansion(mult=0)) + 
+  scale_y_continuous(expand = expansion(mult=0),
+                     breaks = c(0,100,200)) +
+  # facet_grid(model~.) +
   labs(x="Variance Root-to-tip") +
-  scale_color_manual(values = palettes_model) +
+  scale_fill_manual(values = palettes_model) +
   theme(legend.position = 'bottom',

### 2024-07-22 workflow/scripts/functions.R

@@ -27,6 +27,9 @@ names(palette_singleton) <- c("common", "only_blast", "only_fs", "union")
 palette_singleton_p <- c("#00A08A", "#d63a3a", "#ffaa0e", "#6f63bb")
 names(palette_singleton_p) <- c("Common", "only Blast", "only Foldseek", "Union")
 
+palette_singleton_s <- c("#00A08A", "#d63a3a", "#ffaa0e", "#6f63bb")
+names(palette_singleton_s) <- c("Common", "Only Bp", "Only Fs", "Union")
+
 palette_brh <- c("#00A08A","#d63a3a", "#ffaa0e")
 names(palette_brh) <- c("common","blast_brh", "fs_brh")
 

### 2024-09-20 workflow/notebooks/draft.Rmd
switched to notung

--- 2024-07-12 workflow/notebooks/draft.Rmd
@@
 
-tax <- read_delim(params_dataset$taxons_file, show_col_types=F)
+tax <- read_delim(glue(outdir, "/meta/", params_dataset$dataset, "_uniprot_genomes.tsv"), 
+                  show_col_types=F)
 taxmap <- read_delim(glue(outdir, "/db/taxidmap"), 
@@
 
-sptree <- read.tree(params_dataset$species_tree_labels)
+sptree <- read.tree(params_dataset$species_tree)
 translate <- deframe(distinct(dplyr::select(taxmap, Proteome_ID, mnemo)))
@@
 # read input table
-lvls_tax <- rev(c(unique(tax$Clade), "archaea", "bacteria"))
+# lvls_tax <- rev(c(unique(tax$Clade), "archaea", "bacteria"))
 
@@
 
-table <- read_delim(params_dataset$input_table_file, show_col_types=F, 
-                    delim = "\t", col_names = table_columns) %>% 
-  left_join(tax, by = "Tax_ID") %>% 
-  mutate(Clade = ifelse(is.na(Clade), mnemo, Clade),
-         Clade = factor(Clade, levels = lvls_tax))
+# table <- read_delim(params_dataset$input_table_file, show_col_types=F, 
+#                     delim = "\t", col_names = table_columns) %>% 
+#   left_join(tax, by = "Tax_ID") %>% 
+#   mutate(Clade = ifelse(is.na(Clade), mnemo, Clade),
+#          Clade = factor(Clade, levels = lvls_tax))
 
-lddt_df <- read_delim(paste0(params$structure_dir, "/", tax$Proteome_ID, "/", tax$Proteome_ID, "_mean_plddt.tsv"),
+lddt_df <- read_delim(paste0(params$data_dir, "structures/", 
+                             tax$Proteome_ID, "/", tax$Proteome_ID, "_mean_plddt.tsv"),
            col_names = c("target", "mean_lddt"), show_col_types = FALSE)
@@
 ```{r cath}
-CATH_df <- read_delim(list.files(params$cath_dir, full.names = TRUE),
+CATH_df <- read_delim(list.files(glue(params$data_dir, "cath"), full.names = TRUE),
                       delim = "\t", id = "proteome", show_col_types = FALSE) %>%
@@
 
-nms <- read_delim("test/cath/cath-names.txt", 
+nms <- read_delim("resources/cath-names.txt", 
                   comment = "#", delim = "    ",
@@
 
-nms_pfam <- read_delim("test/cath/Pfam-A.clans.tsv", 
+nms_pfam <- read_delim("resources/Pfam-A.clans.tsv", 
            col_names = c("Pfam", "clan", "boh", "abbr", "name"), 
@@
 
+## Fig1
+
+```{r fig1}
+fig1a <- df %>%
+  group_by(singleton) %>%
+  summarise(n = n()) %>%
+  ungroup() %>%
+  mutate(prop=n/sum(n), Filtering="Before") %>%
+  bind_rows(df_filt %>%
+            group_by(singleton) %>%
+            summarise(n = n()) %>%
+            ungroup() %>%
+            mutate(prop=n/sum(n), Filtering="After")) %>%
+  mutate(Filtering=factor(Filtering, labels = c("After", "Before"), ordered = T),
+         singleton=case_when(singleton=="only Blast"~"Only Bp", 
+                             singleton=="only Foldseek" ~ "Only Fs",
+                             .default = singleton)) %>%
+  ggplot(aes(singleton, n, fill=singleton)) +
+  geom_bar(aes(alpha=Filtering), stat = "identity", position = position_dodge2(reverse=TRUE)) +
+  geom_text(aes(group=Filtering, label=paste0(round(prop, 2)*100, "%")),
+            position = position_dodge2(reverse = TRUE, width = 0.9),
+            vjust=1.3, color="white", size=3) +
+  scale_fill_manual(values = palette_singleton_s, guide="none") +
+  scale_alpha_manual(values = c(1,.6)) +
+  scale_y_continuous(expand = c(0,0), labels = scales::label_number(scale = 1e-6)) +
+  labs(fill="Class", x="Class", y="# aligned protein pairs (Mln)") +
+  theme(legend.position = "bottom", legend.box="vertical")
+
+fig1b <- lddt_df %>% 
+  ggplot(aes(mean_lddt)) + 
+  geom_density(fill="#BF4375", size=1) +
+  scale_y_continuous(expand = expansion(mult = c(0, .1))) +
+  labs(x="Mean pLDDT", y="") +
+  geom_vline(xintercept = 40, linetype=2) +
+  theme(axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.line.y = element_blank())
+
+ggsave("draft/figures/fig1_lddt.pdf", fig1b, width = 3, height = 4)
+ggsave("draft/figures/fig1_cnts.pdf", fig1a, width = 4, height = 2)
+```
+
 ## Fig2
@@
 ```{r fig2}
-evals <- c(Inf, 1e-2, 1e-3, 1e-5)
-min_covs <- c(0, 30, 80)
 eval_p <- ggplot(df, aes(-log10(evalue_Blast), -log10(evalue_Foldseek))) + 
@@
   scale_fill_gradientn(colors=wespal) +
-  geom_hline(yintercept = -log10(evals), linetype=2) + 
-  geom_vline(xintercept = -log10(evals), linetype=2) + 
+  ggpubr::stat_cor(label.x.npc = .025) +
+  # geom_hline(yintercept = -log10(evals), linetype=2) + 
+  # geom_vline(xintercept = -log10(evals), linetype=2) + 
   coord_fixed(expand = 0) +
-  labs(x="-log10(Evalue Blast)",y="-log10(Evalue Foldseek)") +
+  labs(x="-log10(Evalue BlastP)",y="-log10(Evalue Foldseek)") +
   scale_x_continuous(limits = c(0, 20)) +
@@
   scale_fill_gradientn(colors=wespal) +
+  ggpubr::stat_cor(label.x.npc = .025) +
   # scale_fill_viridis_c(option = "C") +
@@
   scale_y_continuous(limits = c(0, 100)) +
-  labs(x="Identity Blast",y="Identity Foldseek") +
+  labs(x="% Identity BlastP",y="% Identity Foldseek") +
   theme(legend.position = "none", panel.grid = element_blank())
@@
   stat_bin_2d(breaks=seq(0,100,2)) + 
-  geom_hline(yintercept = min_covs, linetype=2) + 
-  geom_vline(xintercept = min_covs, linetype=2) + 
+  ggpubr::stat_cor(label.x.npc = .025) +
+  # geom_hline(yintercept = min_covs, linetype=2) + 
+  # geom_vline(xintercept = min_covs, linetype=2) + 
   scale_x_continuous(limits = c(0, 100)) +
   scale_y_continuous(limits = c(0, 100)) +
-  labs(x="Query cov. Blast",y="Query cov. Foldseek") +
+  labs(x="Query cov. BlastP",y="Query cov. Foldseek") +
   scale_fill_gradientn(colors=wespal) +
@@
 
-df_multi <- df_filtered %>% 
-    # head(n = 100000) %>%
+df_multi <- df_filt %>% 
+  select(query, target, singleton) %>% 
+    head(n = 1000) %>%
   left_join(CATH_df, by=c("query"="target")) %>% 
@@
   facet_grid(name~.) +
-  labs(y="% in group") +
+  labs(y="% in Singleton group", x="Jaccard Index", 
+       fill="Target sets") +
   scale_y_continuous(expand = expansion(0.01,0)) + 
   scale_fill_manual(values=palette_singleton_p) +
-  theme(legend.position = "bottom",
+  theme(legend.position = "top",
         panel.background = element_rect(fill=NA, color="black"),
@@
 
+disco_fls <- list.files(glue(outdir, "/reco/disco"), pattern = "output.nwk", full.names = T)
+
+disco_rf <- NULL
+for (file in disco_fls){
+  a <- read.tree(file)
+  rf <- TreeDist::RobinsonFoulds(sptree, a, normalize = T)
+  ntips <- sapply(a, function(x) length(x$tip.label))
+  disco_rf <- bind_rows(disco_rf, 
+                        tibble(ntips=ntips, rf=rf, bn=rep(basename(file), length(rf))))
+}
+
 # Read all trees
@@
 rf_text <- group_by(rf_df, model.x, model.y) %>% 
-  filter(as.integer(model.x)<=as.integer(model.y)) %>% 
+  # filter(as.integer(model.x)<=as.integer(model.y)) %>% 
   summarise(value=median(RF, na.rm=T))
@@
 
+n_tips <- sapply(ts, function(x) length(x$tip.label)) %>% 
+  enframe() %>%
+  separate(name, c("gene", "target", "alphabet", "model")) %>%
+  rename("n_tips"="value")
+
+n_sps <- sapply(ts, function(x) 
+  length(unique(taxidmap[taxidmap$target %in% x$tip.label, ]$Tax_ID))) %>% 
+  enframe() %>%
+  separate(name, c("gene", "target", "alphabet", "model")) %>%
+  rename("n_sps"="value")
+
+var_df <- sapply(ts, function(x)
+  var(adephylo::distRoot(phytools::midpoint_root(x)))) %>% 
+  enframe() %>%
+  separate(name, c("gene", "target", "alphabet", "model")) %>%
+  rename("variance_r2t"="value")
+
 # Reconciliation ranger+TCS scores
-reco <- read_delim(glue(outdir,"/reco/", params_dataset$seed, "_DTL.tsv"), 
-                   show_col_types = FALSE)
-
-scores <- read_delim(glue(outdir,"/reco/", params_dataset$seed, "_scores.tsv"), 
-                     show_col_types = FALSE) %>% 
-  left_join(reco) %>% 
+reco <- read_delim(glue(outdir,"/reco/", params_dataset$seed, "_notung.tsv"), 
+                   show_col_types = FALSE, col_names = c("gene", "dups", "losses")) %>% 
+  separate(gene, c("gene", "target", "alphabet", "model")) %>% 
+  left_join(var_df) %>% 
+  left_join(n_tips) %>% 
+  left_join(n_sps) %>% 
   mutate(model=factor(model, levels=models),
-         scores, n_events = (dups+losses)/n_tips)
+         n_events = (dups+losses)/n_tips) %>% 
+  rename("targets"="target")
 ```
@@
 ```{r fig3}
-my_comparisons <- list( c("LG", "3Di"), c("3Di", "GTR"), c("LG", "GTR") )
-plot_bs <- df_bs %>% 
-  mutate(targets=target) %>% 
-  # group_by(gene, target, model) %>% 
-  # summarise(mn_bs=mean(support)) %>% 
-  ggplot(aes(model, support, fill=targets)) +
-  # geom_violin(position = position_dodge(width = 1)) +
-  geom_boxplot(outlier.size = .1) +
-  # ggdist::stat_pointinterval(position = position_dodge(width = 1), color="black") +
-  scale_fill_manual(values = palettes_method) +
-  theme_classic() +
-  labs(y="Boostrap", x="") 
-  # theme(legend.position = "bottom") 
-  # ggpubr::stat_compare_means(aes(group = model),
-  #                            comparisons = my_comparisons,
-  #                            label = "p.format")
 
-rf_plot <- rf_df %>% 
-  ggplot(aes(RF)) +
-  geom_tile(aes(x=0.5, y=0.5, fill=value), data = rf_text) +
-  geom_text(aes(x=0.5, y=.5, label=round(value, 2)), color="white", data = rf_text) +
-  geom_rug(linewidth=.1, data= . %>% filter(as.integer(model.x)>as.integer(model.y))) +
-  geom_density(aes(y = after_stat(scaled)), data= . %>% filter(as.integer(model.x)>as.integer(model.y))) + 
-  scale_fill_distiller(palette="YlGnBu", direction = -1, limits = c(0,1), 
-                       name="Median RF") + 
-  scale_x_continuous(breaks = c(0,.25,.5, .75,1), labels = function(x) sprintf("%.2g", x)) +
-  coord_cartesian(xlim = c(0,1), expand=0) +
-  # facet_grid(method~method2, switch = "both") +
-  facet_grid(model.x~model.y) +
-  # ggh4x::facet_nested(targets+model~targets2+model2, nest_line = element_line(linetype = 1)) +
-  labs(x="RF Distance") +
-  theme_classic() +
-  theme(legend.position = "bottom", 
-        panel.background = element_rect(fill='transparent'),
-        strip.placement = "outside",
-        strip.background = element_blank(),
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        # strip.text = element_text(size=6),
-        axis.text.x = element_text(size=6),
-        axis.text.y = element_blank(),
-        axis.line.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        axis.title.y = element_blank(),
-        legend.margin=margin(0,0,0,0),
-        legend.box.margin=margin(-20,0,0,0))
-
-comps <- NULL
-idx <- 1
-for (el in df_trees$target %>% unique()){
-  for (el2 in df_trees$target %>% unique()){
-    if (el>el2){
-    comps[[idx]] <- c(el, el2)
-    idx <- idx+1
-    }
-  }
-}
-
-ttest <- scores %>%
-  filter(model!="FTPY", id %in% id_tmp) %>%
-  group_by(model) %>% 
-  rstatix::t_test(n_events~targets) %>% 
-  rstatix::add_xy_position(x="model", dodge = 0.8)
-
-bxp <- scores %>% 
-  filter(model!="FTPY", id %in% id_tmp) %>%
-  ggpubr::ggboxplot(x = "model", y = "n_events", fill = "targets",
-                    palette = palettes_method)
-
-bxp +
-ggpubr::stat_pvalue_manual(ttest, label = "p")
-
-ggplot(scores,aes(n_tips, dups+losses, color=targets)) + 
-  geom_point(size=.1) +
-  geom_smooth(method = "lm", alpha=.2) +
-  facet_wrap(~model, nrow = 1,scales = "free") +
-  # coord_cartesian(xlim = c(0,50), ylim = c(0,250)) +
-  scale_color_manual(values = palettes_method)
-
-plot_DL <- scores %>% 
+plot_DL <- reco %>% 
   filter(model!="FTPY") %>% 
@@
   geom_boxplot(outlier.size = .1) +
-  labs(x="", y="(D+L)/# Tips") +
+  labs(y="(D+L)/# Tips", x="",fill="Target sets") +
   # ggthemes::scale_fill_solarized() +
-  scale_fill_manual(values = palettes_method)
-  # theme(legend.position = "bottom")
+  scale_fill_manual(values = palettes_method) +
+  theme(legend.position = "none")
 
-var_r2t <- scores %>% 
+plot_rf <- disco_rf %>% 
+  filter(ntips>10) %>% 
+  # filter(ntips>=length(sptree$tip.label)/2) %>%
+  separate(bn, c("seed", "targets", "alphabet", "model")) %>% 
+  mutate(model=factor(model, levels=models)) %>% 
+  ggplot(aes(model, rf, fill=targets)) + 
+  geom_boxplot() + 
+  scale_fill_manual(values = palettes_method) +
+  labs(x="", y="Norm. RF to species tree", fill="Target sets")
+
+nodes_clades <- fortify(sptree) %>% 
+  # filter(!isTip) %>% 
+  mutate(ordered = rank(y)) %>% 
+  mutate(label_int = fct_reorder(label, y)) %>%
+  select(label_int, ordered)
+
+apro_files <- list.files(glue(outdir, "/reco"), 
+                         pattern = "*support.nwk", full.names = TRUE)
+apro_trees <- read.tree(text = sapply(apro_files, readLines))
+names(apro_trees) <- sapply(str_split(basename(apro_files), "_"), 
+                            function(x) paste0(x[2], "_", x[4]))
+
+trees_df <- fortify(apro_trees) 
+nodes_df <- trees_df %>% 
+  group_by(.id) %>% 
+  mutate(ordered = rank(y)) %>% 
+  left_join(nodes_clades) %>% 
+  separate(.id, c("targets", "model")) %>% 
+  filter(!isTip, label!="") %>% 
+  mutate(label = gsub("\\[|\\]|\\'", "", str_replace_all(label, "[a-z]{1,2}[0-9]=", ""))) %>% 
+  separate(label, c("pp1", "pp2", "pp3", "f1", "f2", "f3", "q1", "q2", "q3"), 
+           ";", convert = TRUE) %>% 
+  mutate(freq=f1+f2+f3, 
+         model=factor(model, levels=models))
+
+plot_apro <- nodes_df %>%
+    filter(model!="FTPY") %>% 
+  ggplot(aes(model, q1, fill=targets)) + 
+  geom_boxplot() + 
+  scale_fill_manual(values = palettes_method) +
+  labs(x="", y="Quartet support", fill="Target sets")
+
+# firstrow <- (plot_rf + plot_DL + 
+#     plot_layout(guides = "collect"))
+# 
+# secondrow <- ( + 
+#     plot_layout(guides = "collect", widths = c(1,2)))
+
+fig3 <- plot_rf + plot_apro + plot_DL +
+  plot_layout(guides = "collect") +
+  plot_annotation(tag_levels = "A") & 
+  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"),
+        legend.position = "bottom")
+
+ggsave("draft/figures/fig3.pdf", fig3, width = 10, height = 4)
+# reco_rnkgs_plot <- reco %>% 
+#   group_by(gene, target) %>% 
+#   slice_min(n_events, with_ties = T) %>% 
+#   group_by(target, model) %>% 
+#   count() %>% 
+#   ggplot(aes(target, n, fill=model)) + 
+#   geom_bar(stat = "identity") +
+#   geom_text(aes(label=n), position = position_stack()) +
+#   scale_fill_manual(values = palettes_model)
+```
+
+```{r FigS1}
+plot_bs <- df_bs %>% 
+  mutate(targets=target,
+         model=factor(model, levels=models)) %>% 
+  ggplot(aes(model, support, fill=targets)) +
+  geom_boxplot(outlier.size = .1) +
+  scale_fill_manual(values = palettes_method) +
+  theme_classic() +
+  labs(y="Boostrap", x="", fill="") +
+  theme(legend.position = "bottom")
+
+var_r2t <- reco %>% 
   filter(model!="FTPY", targets=="common") %>% 
-  ggplot(aes(color=model, variance_r2t)) + 
-  geom_density(size=1, alpha=.5) +
-  # geom_histogram(position = "dodge") + 
-  # coord_cartesian(expand = 0) +
-  scale_x_continuous(limits = c(0,.2)) + 
-  # facet_grid(~targets) + 
-  labs(x="Variance Root-to-tip") +
-  scale_color_manual(values = palettes_model) +
-  theme(legend.position = 'bottom',
+  ggplot(aes(variance_r2t, model)) +
+  geom_violin(scale = "width", color="black", fill="grey40") +
+  scale_x_continuous(limits = c(0,.5), expand = expansion(mult=0, add = c(0,0.05))) +
+  labs(x="Variance Root-to-tip distance") +
+  # scale_fill_manual(values = palettes_model) +
+  theme(legend.position = 'none',
         legend.margin=margin(0,0,0,0),
@@
 
-secondcolumn <- (plot_DL + plot_bs + 
-    plot_layout(guides = "collect", widths = c(5,3)) &
-    theme(legend.position = 'bottom',
-          legend.margin=margin(0,0,0,0),
-          legend.box.margin=margin(-20,0,0,0))) / var_r2t
+rf_plot <- rf_text %>% 
+  ggplot(aes(model.x, model.y)) +
+  geom_tile(aes( fill=value)) +
+  geom_text(aes(label=round(value, 2)), color="white") +
+  scale_fill_distiller(palette="YlGnBu", direction = -1, limits = c(0,1), 
+                       name="Median RF") + 
+  labs(x="Median RF Distance", y="") +
+  theme_classic() +
+  theme(legend.position = "none", 
+        axis.line = element_blank(),
+        axis.ticks = element_blank())
 
-fig3 <- (rf_plot | secondcolumn) +
-  plot_annotation(tag_levels = 'A') & 
-  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
-ggsave("draft/figures/fig3.pdf", fig3, width = 11, height = 6)
-```
-
-```{r FigS1}
 df_red <- select(df_filtered, query, target, singleton)
@@
   ) + 
+  labs(x="Normalised internode distance", y="Method", fill="") +
   scale_x_continuous(expand = c(0,0.01), limits = c(0,1)) +
@@
 
-labeled_sptree <- read.tree("data/sptrees/homo_internal.spTree.nw")
-nodes_clades <- fortify(labeled_sptree) %>% 
-  # filter(!isTip) %>% 
-  mutate(ordered = rank(y)) %>% 
-  mutate(label_int = fct_reorder(label, y)) %>%
-  select(label_int, ordered)
+fig_s1 <- ((plot_bs / rf_plot) + 
+             plot_layout(heights = c(2,1)) | 
+             (plot_dist / var_r2t) + 
+             plot_layout(heights = c(2,1))) +
+  plot_layout(widths = c(1, 2)) +
+  plot_annotation(tag_levels = 'A') & 
+  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
+ggsave("draft/figures/figS1.pdf", fig_s1, width = 10, height = 6)
+```
 
-clades <- read_delim(params_dataset$taxons_file, show_col_types = FALSE)
-
-apro_files <- list.files(glue(outdir, "/reco"), 
-                         pattern = "*support.nwk", full.names = TRUE)
-apro_trees <- read.tree(text = sapply(apro_files, readLines))
-names(apro_trees) <- sapply(str_split(basename(apro_files), "_"), 
-                            function(x) paste0(x[2], "_", x[4]))
-
-trees_df <- fortify(apro_trees) 
-nodes_df <- trees_df %>% 
-  group_by(.id) %>% 
-  mutate(ordered = rank(y)) %>% 
-  left_join(nodes_clades) %>% 
-  separate(.id, c("targets", "model")) %>% 
-  filter(!isTip, label!="") %>% 
-  mutate(label = gsub("\\[|\\]|\\'", "", str_replace_all(label, "[a-z]{1,2}[0-9]=", ""))) %>% 
-  separate(label, c("pp1", "pp2", "pp3", "f1", "f2", "f3", "q1", "q2", "q3"), 
-           ";", convert = TRUE) %>% 
-  mutate(freq=f1+f2+f3)
-
+```{r FigS2}
 pp1_plot <- nodes_df %>%
@@
 
-nodes_df %>% 
-  ggplot(aes(label_int, f1+f2+f3, fill=model)) + 
-  geom_bar(stat="identity", position = "dodge") +
-  facet_grid(targets~.) +
-  scale_fill_manual(values=palettes_model)
+labeled_sptree <- sptree
+labeled_sptree$tip.label <- tax$Species[match(labeled_sptree$tip.label, tax$mnemo)]
 
 sptree_plot <- ggtree(labeled_sptree) + 
-  geom_tiplab() + 
+  geom_tiplab(aes(label=sub("[^A-Z].+? ", ". ", label)),fontface="italic",
+              align = TRUE) + 
   geom_nodelab(color="firebrick", 
                nudge_x = -.15, hjust = 1, 
-               nudge_y = .3, size=2)+
-  scale_x_continuous(expand = c(.2,.2))
+               nudge_y = .3, size=3.5)+
+  xlim(c(-4, 21))
 
-plot_ntips <- scores %>% 
-  filter(model!="FTPY") %>% 
-  ggplot(aes(n_tips, dups+losses, color=targets)) + 
-  geom_point(size=.2) +
-  geom_smooth(method = "lm") +
-  facet_grid(~model) + 
-  scale_color_manual(values = palettes_method) +
-  theme_classic() + 
-  theme(panel.border = element_rect(fill=NA),
-        legend.position = "bottom")
-
-fig_s1 <- (plot_dist + sptree_plot + pp1_plot) +
-  plot_layout(widths = c(1.4, 1, 1.6)) + 
+fig_s2 <- (sptree_plot + pp1_plot) +
+  plot_layout(widths = c(1, 1.5)) +
   plot_annotation(tag_levels = 'A') &
   theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
-ggsave("draft/figures/figS1.pdf", fig_s1, width = 14, height = 6)
-
-# fig_s1 <- (plot_ntips / ((plot_dist + sptree_plot + pp1_plot) +
-#                            plot_layout(widths = c(1.4, 1, 1.6)))) + 
-#   plot_layout(heights = c(.75,1)) 
-# ggsave("draft/figures/figS1.pdf", fig_s1, width = 14, height = 8)
+ggsave("draft/figures/figS2.pdf", fig_s2, width = 12, height = 6, device = cairo_pdf)
 ```
-
-## Answers to Edu's comments
-
-```{r}
-df %>% 
-  filter(query %in% seeds) %>% 
-  count(query, singleton) %>% 
-  ggplot(aes(n, singleton, fill=singleton)) +
-  geom_boxplot()
-
-blast_filtered_eval <- filter(blast, 
-                         query %in% paste0("AF-", aln_seeds, "-F1"),
-                         evalue<as.numeric(params$eval_both)) %>% 
-  group_by(query) %>% 
-  slice_head(n = params$max_seqs) %>% 
-  ungroup()
-
-fs_filtered_eval <- filter(fs, 
-                      query %in% paste0("AF-", aln_seeds, "-F1"),
-                      evalue<as.numeric(params$eval_both)) %>% 
-  group_by(query) %>% 
-  slice_head(n = params$max_seqs) %>% 
-  ungroup()
-
-df_filtered_eval <- full_join(blast_filtered_eval, fs_filtered_eval,
-                by = c("query", "target"), suffix=c("_Blast", "_Foldseek")) %>%
-    filter(query!=target) %>% 
-  left_join(self, by="query") %>%
-  mutate(pident_Foldseek = 100*pident_Foldseek, 
-         qcov_Foldseek = 100*qcov_Foldseek,
-         singleton = case_when(is.na(evalue_Foldseek) ~ "only Blast",
-                               is.na(evalue_Blast) ~ "only Foldseek", 
-                               .default = "Common"))
-
-select(df_filtered_eval, query, target, singleton) %>% 
-  mutate(method="only-eval") %>% 
-  rbind(select(df_filtered, query, target, singleton) %>% 
-          mutate(method="qcov-tcov")) %>% 
-  group_by(method, singleton) %>% 
-  summarise(n = n()) %>%
-  mutate(freq = n / sum(n))
-
-res %>% 
-  group_by(singleton, model) %>% 
-  summarise(median(nn_dist/max_dist))
-```

### 2024-09-20 workflow/scripts/compare_trees.R

@@ -23,7 +23,18 @@ df_ml <- read_delim(snakemake@input[["mltrees"]], delim = "\t",
 ts <- read.tree(text = df_trees$tree)
 names(ts) <- paste(df_trees$gene, df_trees$target, df_trees$alphabet, df_trees$model, sep = "_")
 
-# # get BS distribution
+var_df <- sapply(ts, function(x) 
+  var(adephylo::distRoot(phytools::midpoint_root(x)))) %>% 
+  enframe() %>%
+  separate(name, c("gene", "target", "alphabet", "model")) %>%
+  rename("variance_r2t"="value")
+
+n_tips <- sapply(ts, function(x) length(x$tip.label)) %>% 
+  enframe() %>%
+  separate(name, c("gene", "target", "alphabet", "model")) %>%
+  rename("n_tips"="value")
+
+# get BS distribution
 df_bs <- fortify(ts) %>%
   filter(!isTip) %>%
   mutate(support=as.numeric(label)) %>%
@@ -90,7 +101,9 @@ rf_df <- filter(df_trees, target=="common") %>%
   inner_join(x=., y = ., by = c("gene", "target")) %>%
   filter(!is.na(tree.x), !is.na(tree.y)) %>%
   rowwise() %>%
-  mutate(RF = phangorn::RF.dist(normalize = TRUE, read.tree(text = tree.x), read.tree(text=tree.y)),
+  mutate(RF = phangorn::RF.dist(normalize = TRUE, check.labels = FALSE,
+                                read.tree(text = tree.x),
+                                read.tree(text=tree.y)),
          n_tips = length(read.tree(text = tree.x)$tip.label)) %>%
   select(-tree.x, -tree.y)
 
@@ -98,7 +111,7 @@ rf_text <- group_by(rf_df, model.x, model.y) %>%
   filter(as.integer(model.x)<=as.integer(model.y)) %>% 
   summarise(value=median(RF, na.rm=T))
 
-# plto the heatmap of dist distribution and median value
+# plot the heatmap of dist distribution and median value
 rf_plot <- rf_df %>% 
   ggplot(aes(RF)) +
   geom_tile(aes(x=0.5, y=0.5, fill=value), data = rf_text) +
@@ -127,15 +140,20 @@ rf_plot <- rf_df %>%
         axis.title.y = element_blank())
 
 
-# Reconciliation ranger+TCS scores
-reco <- read_delim(snakemake@input[["reco"]], show_col_types = FALSE) 
[…]

### 2024-09-20 workflow/scripts/functions.R

@@ -10,9 +10,9 @@ fs_columns <- c("query","target","pident","length","mismatch","gapopen",
                 "qstart","qend","sstart","send","evalue","bitscore",
                 "lddt","alntmscore","rmsd","prob","qcov","tcov")
 
-models <- c("LG", "3Di", "GTR", "QT", "FT", "FTPY")
+models <- c("LG", "QT", "FT", "FTPY", "3Di", "GTR")
 
-palettes_model <- c("#BF360C", "#75C323", "#40A79D", "#FFA000", "#D4E157","#646B00")
+palettes_model <- c("#BF360C", "#FFA000", "#D4E157","#646B00", "#75C323", "#40A79D")
 names(palettes_model) <- models
 palettes_model_d <- colorspace::darken(palettes_model, amount = 0.6)
 names(palettes_model_d) <- models

### 2024-09-20 workflow/scripts/plot_example.R

@@ -34,8 +34,9 @@ gff <- read_delim(c(snakemake@input[["gff"]]),
                   show_col_types = FALSE) %>% 
   filter(type %in% c(interesting_types, "Chain"))
 
-reco <- read_delim(snakemake@input[["reco"]], show_col_types = FALSE) %>% 
-  mutate(model=factor(model, levels=models))
+reco <- read_delim(snakemake@input[["reco"]], 
+                   show_col_types = FALSE, col_names = c("gene", "dups", "losses")) %>% 
+  separate(gene, c("id", "targets", "alphabet", "model")) 
 
 plot_list = list()
 plot_domains = list()
@@ -144,7 +145,8 @@ for (seed in seeds) {
     theme_void() + 
     theme(axis.text.x = element_text())
 
-  trees_files <- fls[grepl("nwk$", fls)]
+  trees_files <- fls[grepl(".nwk$", fls)]
+  trees_files <- trees_files[!grepl("_rnm.nwk$", trees_files)]
 
   trees <- read.tree(text = sapply(trees_files, readLines))
   names(trees) <- gsub(paste0(seed,"_|(3Di|aa)_|.nwk$"), "", basename(trees_files))
@@ -168,9 +170,9 @@ for (seed in seeds) {
 
   a <- plot_dom(trees[["union_LG"]], seed, gff, prot, singleton_df, "LG")
   b <- plot_dom(trees[["union_FT"]], seed, gff, prot, singleton_df, "FT")
-  plot_tree_domain <- (a[[2]] | a[[1]]) | (b[[2]] | b[[1]])
+  plot_tree_domain <- (a[[2]] + a[[1]] + b[[2]] + b[[1]]) + plot_layout(nrow = 1)
 
-  plot_ranger <- reco %>% 
+  plot_notung <- reco %>% 
     filter(id==seed) %>% 
     left_join(singleton_df_red) %>% 
     ggplot(aes(targets, (dups+losses)/n, color=model)) + 
@@ -205,7 +207,7 @@ for (seed in seeds) {
   title <- ifelse(is.null(title), seed, paste(seed, "-", title))
 
   first_row <- (plot_seed | plot_venn) + plot_layout(widths = c(4.5,1))
-  bottom_row <- ((plot_conf / plot_ranger) | tree_plot) + plot_layout(widths = c(1, 5))
+  bottom_row <- ((plot_conf / plot_notung) | tree_plot) + plot_layout(widths = c(1, 5))
[…]

### 2024-11-06 workflow/notebooks/draft.Rmd
BIG rework: now it can accept a list of ids or an homolog set

@@
 source("workflow/scripts/functions.R")
-theme_set(theme_classic())
+theme_set(theme_classic(base_family = "Aptos"))
 
@@
 params <- yaml::read_yaml("config/params.yaml")
-params_dataset <- yaml::read_yaml("config/Hsapopi_test.yaml")
+params_dataset <- yaml::read_yaml("config/Hsap_draft.yaml")
 
-outdir <- glue(params$outdir, params_dataset$dataset)
+homodir <- glue(params$outdir, "homology/", params_dataset$homology_dataset)
+phylodir <- glue(params$outdir, "phylogeny/", params_dataset$phylo_dataset)
 ```
@@
 ```{r read_stuff, cache=FALSE}
-seeds <- readLines(glue(outdir, "/ids/UP000005640_common.ids"))
+seeds <- readLines(params_dataset$test_seeds)
 
-tax <- read_delim(glue(outdir, "/meta/", params_dataset$dataset, "_uniprot_genomes.tsv"), 
+tax <- read_delim(glue("results/data/meta/",
+                       params_dataset$homology_dataset, "_uniprot_genomes.tsv"), 
                   show_col_types=F)
-taxmap <- read_delim(glue(outdir, "/db/taxidmap"), 
+taxmap <- read_delim(glue(homodir, "/db/taxidmap"), 
                      delim = "\t", col_names = c("label", "Tax_ID"),
@@
 
-taxidmap <- read_delim(glue(outdir,"/db/taxidmap"), 
+taxidmap <- read_delim(glue(homodir,"/db/taxidmap"), 
                        col_names = c("target", "Tax_ID"), show_col_types=F)
@@
 
-# read input table
-# lvls_tax <- rev(c(unique(tax$Clade), "archaea", "bacteria"))
-
 table_columns <- c('Proteome_ID', 'Tax_ID', 'count1', 'count2', 'count3', 
                    'genome', 'source', 'species', 'mnemo')
-
-# table <- read_delim(params_dataset$input_table_file, show_col_types=F, 
-#                     delim = "\t", col_names = table_columns) %>% 
-#   left_join(tax, by = "Tax_ID") %>% 
-#   mutate(Clade = ifelse(is.na(Clade), mnemo, Clade),
-#          Clade = factor(Clade, levels = lvls_tax))
 
@@
 
-aln_seeds <- readLines(glue(outdir,"/ids/", params_dataset$seed, "_aln.ids"))
+aln_seeds <- readLines(glue(phylodir,"/ids/", params_dataset$seed, "_aln.ids"))
 ```
@@
 ```{r read_self_blast}
-blast_self <- read_delim(glue(outdir, "/homology/allvall/",
+blast_self <- read_delim(glue(homodir, "/allvall/",
                               params_dataset$seed, "_", params_dataset$seed, "_blast.tsv"),
@@
 
-fs_self <- read_delim(glue(outdir, "/homology/allvall/",
-                              params_dataset$seed, "_", params_dataset$seed, "_fs.tsv"), 
+fs_self <- read_delim(glue(homodir, "/allvall/", params_dataset$seed, 
+                           "_", params_dataset$seed, "_fs.tsv"), 
                       show_col_types = FALSE,
@@
 
-```{r read_fs}
-fs <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_fs.tsv"), 
+```{r read_homo}
+fs <- read_delim(glue(homodir, "/", params_dataset$seed, "_fs.tsv"), 
                  col_names = fs_columns, show_col_types = FALSE) %>% 
@@
 
-fs_brhs <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_fs_brh.tsv"),
-                     col_names = c("query", "target"), show_col_types = FALSE) %>% 
-  # filter(query %in% seeds) %>%
-  mutate(method="fs_brh") 
-```
-
-```{r read_blast}
-blast <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_blast.tsv"), 
+blast <- read_delim(glue(homodir, "/", params_dataset$seed, "_blast.tsv"), 
                     col_names = blast_columns, show_col_types = FALSE) %>% 
@@
   mutate(evalue = ifelse(evalue==0, 1e-180, evalue))
-
-blast_brhs <- read_delim(glue(outdir, "/homology/", params_dataset$seed, "_blast_brh.tsv"),
-                     col_names = c("query", "target"), show_col_types = FALSE) %>% 
-  # filter(query %in% seeds) %>%
-  mutate(method="blast_brh")
 ```
@@
 
-brhs <- rbind(blast_brhs, fs_brhs) %>% 
-    group_by(query, target, method) %>%
-    mutate(n=1) %>%
-    pivot_wider(names_from = method, values_fill = 0, values_from = n) %>%
-    mutate(common = ifelse(fs_brh+blast_brh==2, 1, 0))
-
 blast_filt <- filter(blast, 
-                         # query %in% paste0("AF-", aln_seeds, "-F1"),
                          length/qlen*100>params$coverage, 
@@
 fs_filt <- filter(fs, 
-                      # query %in% paste0("AF-", aln_seeds, "-F1"),
                       qcov*100>params$coverage, 
@@
                                .default = "Common"))
-# grouped_brhs <- brhs %>%
+
+# blast_filtered <- filter(blast, 
+#                          query %in% paste0("AF-", aln_seeds, "-F1"),
+#                          length/qlen*100>params$coverage, 
+#                          length/slen*100>params$coverage,
+#                          evalue<as.numeric(params$eval_both)) %>% 
 #   group_by(query) %>% 
-#   summarise(fs_brh = sum(fs_brh), blast_brh=sum(blast_brh), common_brh=sum(common))
-blast_filtered <- filter(blast, 
-                         query %in% paste0("AF-", aln_seeds, "-F1"),
-                         length/qlen*100>params$coverage, 
-                         length/slen*100>params$coverage,
-                         evalue<as.numeric(params$eval_both)) %>% 
-  group_by(query) %>% 
-  slice_head(n = params$max_seqs) %>% 
-  ungroup()
-
-fs_filtered <- filter(fs, 
-                      query %in% paste0("AF-", aln_seeds, "-F1"),
-                      qcov*100>params$coverage, 
-                      tcov*100>params$coverage, 
-                      evalue<as.numeric(params$eval_both)) %>% 
-  group_by(query) %>% 
-  slice_head(n = params$max_seqs) %>% 
-  ungroup()
-
-df_filtered <- full_join(blast_filtered, fs_filtered,
-                by = c("query", "target"), suffix=c("_Blast", "_Foldseek")) %>%
-    filter(query!=target) %>% 
-  left_join(self, by="query") %>%
-  mutate(pident_Foldseek = 100*pident_Foldseek, 
-         qcov_Foldseek = 100*qcov_Foldseek,
-         singleton = case_when(is.na(evalue_Foldseek) ~ "only Blast",
-                               is.na(evalue_Blast) ~ "only Foldseek", 
-                               .default = "Common"))
+#   slice_head(n = params$max_seqs) %>% 
+#   ungroup()
+# 
+# fs_filtered <- filter(fs, 
+#                       query %in% paste0("AF-", aln_seeds, "-F1"),
+#                       qcov*100>params$coverage, 
+#                       tcov*100>params$coverage, 
+#                       evalue<as.numeric(params$eval_both)) %>% 
+#   group_by(query) %>% 
+#   slice_head(n = params$max_seqs) %>% 
+#   ungroup()
+# 
+# df_filtered <- full_join(blast_filtered, fs_filtered,
+#                 by = c("query", "target"), suffix=c("_Blast", "_Foldseek")) %>%
+#     filter(query!=target) %>% 
+#   left_join(self, by="query") %>%
+#   mutate(pident_Foldseek = 100*pident_Foldseek, 
+#          qcov_Foldseek = 100*qcov_Foldseek,
+#          singleton = case_when(is.na(evalue_Foldseek) ~ "only Blast",
+#                                is.na(evalue_Blast) ~ "only Foldseek", 
+#                                .default = "Common"))
 ```
@@
 ```{r cath}
-CATH_df <- read_delim(list.files(glue(params$data_dir, "cath"), full.names = TRUE),
+CATH_df <- read_delim(paste0(params$data_dir, "cath/", tax$Proteome_ID, "_cath.tsv"),
                       delim = "\t", id = "proteome", show_col_types = FALSE) %>%
-  # filter(!is.na(Gene3D)) %>% 
-  mutate(proteome = gsub("_cath.tsv", "", basename(proteome)),
-         n_3d_domains = str_count(Gene3D, ";"),
-         n_Pfam_domains = str_count(Pfam, ";")) %>%
-  # filter(n_domains==1) %>% 
+  mutate(proteome = gsub("_cath.tsv", "", basename(proteome))) %>%
+         # n_3d_domains = str_count(Gene3D, ";"),
+         # n_Pfam_domains = str_count(Pfam, ";")
   mutate(target=paste0("AF-", Entry, "-F1")) %>% 
-    select(-proteome, Entry)
+  rename("CATH"="Gene3D") %>% 
+  select(target, Pfam, CATH, -proteome, -Entry) %>% 
+  mutate(CAT = gsub("(\\d+\\.\\d+\\.\\d+)\\.\\d+", "\\1", CATH),
+         CA = gsub("(\\d+\\.\\d+)\\.\\d+", "\\1", CAT),
+         C = gsub("\\.\\d+", "\\1", CA))
 
-nms <- read_delim("resources/cath-names.txt", 
-                  comment = "#", delim = "    ",
-                  col_names = c("Gene3D", "abbr", "name"), 
-                  show_col_types = FALSE) %>% 
-  select(-abbr) %>% 
-  mutate(name=gsub("^:", "", name))
+ortho_dbs <- read_delim(paste0(params$data_dir, "ids/", tax$Proteome_ID, "_ids.tsv"),
+                      delim = "\t", id = "proteome", show_col_types = FALSE) %>% 
+  mutate(proteome = gsub("_ids.tsv", "", basename(proteome))) %>% 
+  mutate(target=paste0("AF-", Entry, "-F1")) %>% 
+  select(-proteome, -Entry)
 
-nms_pfam <- read_delim("resources/Pfam-A.clans.tsv", 
-           col_names = c("Pfam", "clan", "boh", "abbr", "name"), 
-           show_col_types = FALSE) %>% 
-  select(-name, -clan, -boh)
+ortho_dbs <- full_join(CATH_df, ortho_dbs)
 
-long_CATH_df <- CATH_df %>% 
-  # filter(target %in% unique(blast$query)) %>% 
-  mutate(Gene3D=strsplit(Gene3D, ";")) %>% 
-  unnest(Gene3D) %>% 
-  separate(col = Gene3D, into = c("C", "A", "T", "H"), sep = "\\.", remove = FALSE) %>% 
-  mutate(CA = paste(C,A,sep = "."), CAT=paste(C,A,T,sep = ".")) %>% 
-  left_join(nms, by=c("C"="Gene3D")) %>%  
-  left_join(nms, by=c("CA"="Gene3D"), suffix = c("_C", "_CA")) %>% 
-  left_join(nms, by=c("CAT"="Gene3D"), suffix = c("_CA", "_CAT")) %>% 
-  left_join(nms, by=c("Gene3D"), suffix = c("_CAT", "_CATH"))
-
-long_CATH_nms <- long_CATH_df %>%
-    select(-c(Entry, Pfam, n_Pfam_domains, target, n_3d_domains)) %>% 
-    distinct()
+# nms <- read_delim("resources/cath-names.txt", 
+#                   comment = "#", delim = "    ",
+#                   col_names = c("Gene3D", "abbr", "name"), 
+#                   show_col_types = FALSE) %>% 
+#   select(-abbr) %>% 
+#   mutate(name=gsub("^:", "", name))
+# 
+# nms_pfam <- read_delim("resources/Pfam-A.clans.tsv", 
+#            col_names = c("Pfam", "clan", "boh", "abbr", "name"), 
+#            show_col_types = FALSE) %>% 
+#   select(-name, -clan, -boh)
+# 
+# long_CATH_df <- CATH_df %>% 
+#   # filter(target %in% unique(blast$query)) %>% 
+#   mutate(Gene3D=strsplit(Gene3D, ";")) %>% 
+#   unnest(Gene3D) %>% 
+#   separate(col = Gene3D, into = c("C", "A", "T", "H"), sep = "\\.", remove = FALSE) %>% 
+#   mutate(CA = paste(C,A,sep = "."), CAT=paste(C,A,T,sep = ".")) %>% 
+#   left_join(nms, by=c("C"="Gene3D")) %>%  
+#   left_join(nms, by=c("CA"="Gene3D"), suffix = c("_C", "_CA")) %>% 
+#   left_join(nms, by=c("CAT"="Gene3D"), suffix = c("_CA", "_CAT")) %>% 
+#   left_join(nms, by=c("Gene3D"), suffix = c("_CAT", "_CATH"))
+# 
+# long_CATH_nms <- long_CATH_df %>%
+#     select(-c(Entry, Pfam, n_Pfam_domains, target, n_3d_domains)) %>% 
+#     distinct()
 ```
@@
   geom_bar(aes(alpha=Filtering), stat = "identity", position = position_dodge2(reverse=TRUE)) +
-  geom_text(aes(group=Filtering, label=paste0(round(prop, 2)*100, "%")),
-            position = position_dodge2(reverse = TRUE, width = 0.9),
-            vjust=1.3, color="white", size=3) +
+  # geom_text(aes(group=Filtering, label=paste0(round(prop, 2)*100, "%")),
+  #           position = position_dodge2(reverse = TRUE, width = 0.9),
+  #           vjust=1.3, color="white", size=3) +
   scale_fill_manual(values = palette_singleton_s, guide="none") +
@@
   scale_y_continuous(expand = c(0,0), labels = scales::label_number(scale = 1e-6)) +
-  labs(fill="Class", x="Class", y="# aligned protein pairs (Mln)") +
+  labs(fill="Class", x="Class", y="Milions of aligned protein pairs") +
   theme(legend.position = "bottom", legend.box="vertical")
@@
 
-ggsave("draft/figures/fig1_lddt.pdf", fig1b, width = 3, height = 4)
-ggsave("draft/figures/fig1_cnts.pdf", fig1a, width = 4, height = 2)
+# ggsave("draft/figures/fig1_lddt.pdf", fig1b, width = 3, height = 4)
+# ggsave("draft/figures/fig1_cnts.pdf", fig1a, width = 4, height = 2)
 ```
@@
 
-df_multi <- df_filt %>% 
-  select(query, target, singleton) %>% 
-    head(n = 1000) %>%
-  left_join(CATH_df, by=c("query"="target")) %>% 
-  filter(!is.na(Pfam), !is.na(Gene3D)) %>% 
-  left_join(CATH_df, by="target") %>% 
+dbs <- colnames(ortho_dbs)[2:dim(ortho_dbs)[2]]
+# Loop through each column name and apply a function
+jaccard_df <- map_dfr(dbs, ~ {
+  col_name <- .x
+select(df_filt, query, target, singleton) %>% 
+    # filter(query=="AF-A0A024RBG1-F1") %>%
+    left_join(select(ortho_dbs, target, col_name), by=c("query"="target")) %>% 
+    left_join(select(ortho_dbs, target, col_name), by=c("target")) %>% 
+    rename("query_set"=paste0(col_name, ".x"), "target_set"=paste0(col_name, ".y")) %>% 
+    # Exclude queries without any domain
+    filter(!is.na(query_set), !is.na(target_set)) %>% 
+    group_by(query, singleton) %>%
+    distinct(query, singleton, query_set, target_set) %>% 
+    summarise(set_q = strsplit(paste(query_set, collapse = ""), ";"),
+              set_t = strsplit(paste(target_set, collapse = ""), ";")) %>%
     rowwise() %>%
-  mutate(Gene3D.x=strsplit(Gene3D.x, ";"),
-         Pfam.x=strsplit(Pfam.x, ";"),
-         Gene3D.y=strsplit(Gene3D.y, ";"),
-         Pfam.y=strsplit(Pfam.y, ";")) %>% 
-  mutate(n_3d_common=length(intersect(Gene3D.x,Gene3D.y)),
-         n_Pfam_common=length(intersect(Pfam.x,Pfam.y))) %>% 
-  # filter(n_Pfam_common>0, n_3d_common>0) %>% 
-  select(-c(Pfam.x,Pfam.y,Gene3D.x,Gene3D.y)) %>% 
-  mutate(Jaccard_3d=n_3d_common/pmax(n_3d_domains.x,n_3d_domains.y),
-         Jaccard_Pfam=n_Pfam_common/pmax(n_Pfam_domains.x,n_Pfam_domains.y)) %>% 
-  select(query, target, singleton, contains("Jaccard"), contains("common"))
+    mutate(jaccard = length(intersect(set_q, set_t))/length(union(set_q, set_t))) %>% 
+    select(-set_q, -set_t) %>% 
+    mutate(db=col_name)
+
+})
+
+ggplot(jaccard_df, aes(jaccard, fill=singleton)) +
+  geom_histogram(position = "dodge") +
+  facet_wrap(~db, scales = "free_y") +
+  scale_y_continuous(expand = expansion(0.01,0)) + 
+  scale_fill_manual(values=palette_singleton_p) +
+  theme_bw() +
+  theme(legend.position = "top",
+        # panel.background = element_rect(fill=NA, color="black"),
+        # axis.line = element_blank()
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(), 
+        strip.background = element_rect(fill=NA)
+        # panel.grid.major.x = element_line()
+        )
+
+ggsave("test/bench.pdf", width = 8, height = 5, device = cairo_pdf)
+# plot_dbs_list[[db]] <- p_db
+
+# df_multi <- df_filt %>% 
+#   select(query, target, singleton) %>% 
+#     head(n = 100) %>%
+#   left_join(CATH_df, by=c("query"="target")) %>% 
+#   filter(!is.na(Pfam), !is.na(Gene3D)) %>% 
+#   left_join(CATH_df, by="target") %>% 
+#   rowwise() %>% 
+#   mutate(Gene3D.x=strsplit(Gene3D.x, ";"),
+#          Pfam.x=strsplit(Pfam.x, ";"),
+#          Gene3D.y=strsplit(Gene3D.y, ";"),
+#          Pfam.y=strsplit(Pfam.y, ";")) %>% 
+#   mutate(n_3d_common=length(intersect(Gene3D.x,Gene3D.y)),
+#          n_Pfam_common=length(intersect(Pfam.x,Pfam.y))) %>% 
+#   # filter(n_Pfam_common>0, n_3d_common>0) %>% 
+#   select(-c(Pfam.x,Pfam.y,Gene3D.x,Gene3D.y)) %>% 
+#   mutate(Jaccard_3d=n_3d_common/pmax(n_3d_domains.x,n_3d_domains.y),
+#          Jaccard_Pfam=n_Pfam_common/pmax(n_Pfam_domains.x,n_Pfam_domains.y)) %>% 
+#   select(query, target, singleton, contains("Jaccard"), contains("common"))
 

### 2024-11-06 workflow/envs/reco.post-deploy.sh

@@ -4,3 +4,18 @@ unzip $CONDA_PREFIX/lib/Notung-2.9.1.5.zip  -d $CONDA_PREFIX/lib/
 
 echo -e '#!/usr/bin/env bash\njava -jar $CONDA_PREFIX/lib/Notung-2.9.1.5.jar $@' > $CONDA_PREFIX/bin/notung
 chmod +x $CONDA_PREFIX/bin/notung
+
+
+git clone https://github.com/JSdoubleL/DISCO $CONDA_PREFIX/lib/DISCO
+sed -i '1s/^/\#!\/usr\/bin\/env python3\n/' $CONDA_PREFIX/lib/DISCO/disco.py
+
+chmod +x $CONDA_PREFIX/lib/DISCO/disco.py
+cp $CONDA_PREFIX/lib/DISCO/disco.py $CONDA_PREFIX/bin
+
+
+# install astral-pro
+git clone https://github.com/chaoszhang/ASTER $CONDA_PREFIX/lib/ASTER
+cd $CONDA_PREFIX/lib/ASTER
+make astral-pro
+cp bin/astral-pro $CONDA_PREFIX/bin/
+cd -

### 2024-11-06 workflow/envs/sp_tree.post-deploy.sh

@@ -1,13 +1,0 @@
-# install astral-pro
-git clone https://github.com/chaoszhang/ASTER $CONDA_PREFIX/lib/ASTER
-cd $CONDA_PREFIX/lib/ASTER
-make astral-pro
-cp bin/astral-pro $CONDA_PREFIX/bin/
-cd -
-
-# install Ranger-DTL
-wget https://compbio.engr.uconn.edu/wp-content/uploads/sites/2447/2019/08/Linux.zip -O $CONDA_PREFIX/lib/RangerDTL.zip
-unzip $CONDA_PREFIX/lib/RangerDTL.zip -d $CONDA_PREFIX/lib/RangerDTL
-chmod +x $CONDA_PREFIX/lib/RangerDTL/Linux/CorePrograms/*
-mv $CONDA_PREFIX/lib/RangerDTL/Linux/CorePrograms/* $CONDA_PREFIX/bin/
-

### 2024-11-06 workflow/envs/sp_utils.post-deploy.sh

@@ -1,6 +1,0 @@
-# install foldmason
-# find other binaries here: https://mmseqs.com/foldmason/
-mkdir -p $CONDA_PREFIX/lib/
-wget https://mmseqs.com/foldmason/foldmason-linux-avx2.tar.gz -O $CONDA_PREFIX/lib/foldmason.tar.gz
-
-tar xvzf $CONDA_PREFIX/lib/foldmason.tar.gz -C $CONDA_PREFIX/bin/ --strip-components=2 ./bin/foldmason

### 2024-11-06 workflow/scripts/analyze_bl.R

@@ -5,7 +5,7 @@ source("workflow/scripts/functions.R")
 theme_set(theme_classic())
 
 # all trees
-df_trees <- read_delim(snakemake@input[["trees"]], show_col_types = FALSE,
+df_trees <- read_delim(snakemake@input[["trees"]], delim = "\t", show_col_types = FALSE,
                        col_names = c("gene", "targets", "alphabet", "model", "tree")) %>% 
   mutate(model=factor(model, levels=models)) %>% 
   filter(targets=="common")

### 2024-11-06 workflow/scripts/compare_sptree.R

@@ -1,0 +1,124 @@
+suppressPackageStartupMessages(library(tidyverse))
+suppressPackageStartupMessages(library(ggtree))
+suppressPackageStartupMessages(library(patchwork))
+source("workflow/scripts/functions.R")
+theme_set(theme_classic())
+
+# read species tree
+sptree <- read.tree(snakemake@input[["sptree"]])
+
+taxidmap <- read_delim(snakemake@input[["taxidmap"]], 
+                       col_names = c("target", "Tax_ID"), show_col_types = FALSE)
+
+# trees Dataframe
+df_trees <- read_delim(snakemake@input[["trees"]],
+                       show_col_types = FALSE,
+                       col_names = c("gene", "target", "alphabet", "model", "tree")) %>% 
+  mutate(model=factor(model, levels=models))
+
+# ML trees with selected model
+df_ml <- read_delim(snakemake@input[["mltrees"]], delim = "\t", 
+                    show_col_types=FALSE) %>% 
+  # mutate(fn=gsub("_mltrees.txt", "", basename(fn))) %>% 
+  # separate(fn, c("gene", "target", "alphabet", "model"), sep = "_") %>% 
+  mutate(model=factor(gsub("3DI", "3Di", gsub("GTR20", "GTR", gsub("\\+.*", "", Model))), levels=models), 
+         rate=sub("\\+", "", gsub("GTR20|3DI|LG|AF|LLM", "", Model)))
+
+# Disco similarity to sptree
+disco_fls <- c(snakemake@input[["disco"]])
+
+# Astral-pro results
+clades <- read_delim(snakemake@input[["table"]], show_col_types = FALSE)
+
+apro_files <- snakemake@input[["apro_trees"]]
+apro_trees <- read.tree(text = sapply(apro_files, readLines))
+names(apro_trees) <- sapply(str_split(basename(apro_files), "_"), 
+                            function(x) paste0(x[2], "_", x[4]))
+
+
+
+# 1. Compute astral-pro quartet
+nodes_clades <- fortify(sptree) %>% 
+  # filter(!isTip) %>% 
+  mutate(ordered = rank(y)) %>% 
+  mutate(label_int = fct_reorder(label, y)) %>%
+  select(label_int, ordered)
+
+trees_df <- fortify(apro_trees) 
+nodes_df <- trees_df %>% 
+  group_by(.id) %>% 
[…]

### 2024-11-06 workflow/scripts/fasta2phylip.py

@@ -1,0 +1,14 @@
+#!/usr/bin/env python
+# adapted from https://github.com/jvollme/fasta2phylip/blob/master/fasta2phylip.py
+import os
+from Bio import AlignIO
+
+def main():
+	infile = open(snakemake.input[0], "r")
+	outfile = open(snakemake.output[0], "w")
+	alignments = AlignIO.parse(infile, "fasta")
+	AlignIO.write(alignments, outfile, "phylip-relaxed")
+	infile.close()
+	outfile.close()
+	
+main()

### 2024-11-06 workflow/scripts/functions.R

@@ -10,7 +10,7 @@ fs_columns <- c("query","target","pident","length","mismatch","gapopen",
                 "qstart","qend","sstart","send","evalue","bitscore",
                 "lddt","alntmscore","rmsd","prob","qcov","tcov")
 
-models <- c("LG", "QT", "FT", "FTPY", "3Di", "GTR", "AF", "LLM")
+models <- c("LG", "FM", "FT", "FTPY", "3Di", "GTR", "AF", "LLM")
 
 palettes_model <- c("#BF360C", "#FFA000", "#D4E157","#646B00", "#75C323", "#40A79D", "#33ceff", "#3396ff")
 names(palettes_model) <- models

### 2024-11-06 workflow/scripts/get_QFO_sptree.R

@@ -1,0 +1,25 @@
+library(ape)
+
+t <- read.nhx("resources/species_tree_QFO2020.nx")
+t <- TreeTools::Preorder(t@phylo)
+t$tip.label <- gsub("_.*", "", t$tip.label)
+t <- TreeTools::Subtree(t, getMRCA(t, tip = c("HUMAN", "GIAIC")))
+t$tip.label <- gsub("CANFA", "CANLF", t$tip.label)
+t <- TreeTools::AddTip(t, getMRCA(t, tip = c("DANRE", "ORYLA")), label = "LEPOC")
+t <- TreeTools::AddTip(t, getMRCA(t, tip = c("DROME", "BOMMO")), label = "TRICA")
+t <- TreeTools::AddTip(t, where = "TRIVA", label = "TRIV3")
+t <- TreeTools::AddTip(t, where = "SCHMA", label = "HELRO")
+t <- TreeTools::AddTip(t, where = "SORBI", label = "MAIZE")
+# Some species are not found in Euka dataset and this causes a bug in A-Pro
+# mammals <- c("PONAB", "NOMLE", "MACMU",
+#              "CALJA", "OTOGA", "CAVPO",
+#              "SPETR", "RABIT", "PIG", 
+#              "HORSE", "CANLF", "AILME", "MYOLU", "LOXAF", "SARHA", "ORNAN", "CHICK")
+# t <- TreeTools::DropTip(t, mammals)
+write.tree(t, "data/sptrees/QFO.nwk")
+
+# p <- ggtree(t) + 
+#   geom_tiplab() + 
+#   xlim(c(0,30))
+# 
+# ggsave("test/qfo.pdf", width = 8, height = 12)

### 2024-11-06 workflow/scripts/get_sets.R

@@ -1,0 +1,59 @@
+suppressPackageStartupMessages(library(tidyverse))
+source("workflow/scripts/functions.R")
+
+ids <- read_delim(snakemake@input[["seed_file"]], show_col_types = FALSE,
+                  delim = "\t", col_names = c("group", "target"))
+ids_todo <- unique(ids$group)
+oprefix <- dirname(dirname(snakemake@output[[1]][1]))
+
+if (ncol(ids)>1){
+  print("The orthogroups are already set: only union set will be computed")
+  for (id in ids_todo){
+    hits <- filter(ids, group==id) %>% 
+      pull(target)
+    writeLines(hits, paste0(oprefix, "/", id, "/", id, "_common.ids"))
+    } 
+  } else {
+  print("proceding with phylomeDB approach")
+  blast <- read_delim(snakemake@input[["blast"]],
+                      show_col_types=FALSE, col_names = blast_columns) %>% 
+    mutate(type="blast")
+  
+  fs <- read_delim(snakemake@input[["fs"]],
+                      show_col_types=FALSE, col_names = fs_columns) %>% 
+    mutate(type="fs")
+  
+  df <- full_join(select(blast, query, target, type), select(fs, query, target, type),
+            by = c("query", "target"), suffix=c("_blast", "_fs")) %>% 
+    mutate(singleton = case_when(is.na(type_fs) ~ "only_blast",
+                          is.na(type_blast) ~ "only_fs", .default = "common"))
+
+  for (id in ids_todo){
+    blast_hits <- filter(df, query==id) %>% 
+      filter(type_blast=="blast", target!=id) %>% 
+      slice_head(n = 150) %>% 
+      pull(target)
+    blast_hits <- c(id, blast_hits)
+    fs_hits <- filter(df, query==id) %>% 
+      filter(type_fs=="fs", target!=id) %>% 
+      slice_head(n = snakemake@params[["max_seqs"]]-1) %>% 
+      pull(target)
+    fs_hits <- c(id, fs_hits)
+
+    if ("blast" %in% snakemake@params[["modes"]]) {
+      writeLines(blast_hits, paste0(oprefix,  "/", id, "/", id, "_blast.ids"))
+    }
+    if ("fs" %in% snakemake@params[["modes"]]) {
+      writeLines(fs_hits, paste0(oprefix,  "/", id, "/", id, "_fs.ids"))
+    }
[…]

### 2024-11-06 workflow/scripts/mask_structures.py

@@ -17,7 +17,7 @@ if __name__ == '__main__':
         low_confidence_sites = list(confidence[confidence['confidenceScore']<min_lddt]['residueNumber'])
 
         for site in low_confidence_sites:
-            record.seq = record.seq[:site-1] + 'X' + record.seq[site:]
+            record.seq = record.seq[:site-1] + '-' + record.seq[site:]
 
         out_seqs.append(record)
 

### 2024-11-06 workflow/scripts/plot_example.R

@@ -18,7 +18,6 @@ taxmap <- read_delim(snakemake@input[["taxidmap"]],
                        show_col_types = FALSE), by = "Tax_ID") %>% 
   dplyr::select(label, Proteome_ID, mnemo)
 
-
 sptree <- read.tree(snakemake@input[["sptree"]])
 translate <- deframe(distinct(dplyr::select(taxmap, Proteome_ID, mnemo)))
 sptree$tip.label <- names(translate)[match(translate, sptree$tip.label)]

### 2024-11-07 workflow/notebooks/draft.Rmd
implemented eggnog and oma benchmark

@@
 
+# ortho_dbs %>% 
+#   left_join(taxmap, by=c("target"="label")) %>% 
+#   group_by(mnemo) %>% 
+#   summarise_all(~ sum(is.na(.))/nrow(ortho_dbs)) %>% 
+#   filter(!is.na(mnemo)) %>% 
+#   select(-target) %>% 
+#   pivot_longer(!mnemo) %>% 
+#   ggplot(aes(name, value)) + 
+#   geom_boxplot() + 
+#   labs(x="DB", y="% NA")
+
 # nms <- read_delim("resources/cath-names.txt", 
@@
 dbs <- colnames(ortho_dbs)[2:dim(ortho_dbs)[2]]
+dbs <- setdiff(dbs, c("InParanoid"))
+dbs <- dbs[1:5]
+
+binwidth <- 0.1
+
 # Loop through each column name and apply a function
+jaccard_df_mono <- map_dfr(dbs, ~ {
+  col_name <- .x
+  select(df_filt, query, target, singleton) %>% 
+    left_join(select(ortho_dbs, target, col_name), by=c("query"="target")) %>% 
+    left_join(select(ortho_dbs, target, col_name), by=c("target")) %>% 
+    rename("query_set"=paste0(col_name, ".x"), "target_set"=paste0(col_name, ".y")) %>% 
+    mutate(n_domain_q = str_count(pattern = "\\;", query_set),
+           n_domain_t = str_count(pattern = "\\;", target_set)) %>% 
+    filter(n_domain_q==1) %>% 
+    # Exclude queries without any domain
+    filter(!is.na(query_set), !is.na(target_set)) %>% 
+    mutate(is_same = query_set==target_set) %>% 
+    group_by(query, singleton) %>%
+    summarise(jaccard=sum(is_same)/n()) %>% 
+    # select(-set_q, -set_t) %>% 
+    mutate(db=col_name)
+
+})
+
+p_dbs_mono <- ggplot(jaccard_df_mono, aes(jaccard, fill=singleton)) +
+  geom_histogram(aes(group=singleton, y=after_stat(binwidth*density)), 
+                 position = "dodge", binwidth = binwidth) +
+  facet_wrap(~db, ncol = 5) +
+  scale_y_continuous(expand = expansion(0.01,0)) + 
+  scale_fill_manual(values=palette_singleton_p) +
+  theme_bw(base_family = "Aptos") +
+  labs(title="Only monodomain", x="Jaccard Index", y="Scaled density", fill="Target sets") +
+  theme(legend.position = "none",
+        strip.background = element_rect(fill=NA))
+
+# Loop through each column name and apply a function
+jaccard_df_NA <- map_dfr(dbs, ~ {
+  col_name <- .x
+  select(df_filt, query, target, singleton) %>% 
+    # filter(query=="AF-A0A024RBG1-F1") %>%
+    left_join(select(ortho_dbs, target, col_name), by=c("query"="target")) %>% 
+    left_join(select(ortho_dbs, target, col_name), by=c("target")) %>% 
+    rename("query_set"=paste0(col_name, ".x"), "target_set"=paste0(col_name, ".y")) %>% 
+    # Exclude queries without any domain
+    filter(!is.na(query_set)) %>% 
+    group_by(query, singleton) %>%
+    distinct(query, singleton, query_set, target_set) %>% 
+    summarise(set_q = strsplit(paste(query_set, collapse = ""), ";"),
+              set_t = strsplit(paste(target_set, collapse = ""), ";")) %>%
+    rowwise() %>%
+    mutate(jaccard = length(intersect(set_q, set_t))/length(union(set_q, set_t))) %>% 
+    select(-set_q, -set_t) %>% 
+    mutate(db=col_name)
+
+})
+
+p_dbs_NA <- ggplot(jaccard_df_NA, aes(jaccard, fill=singleton)) +
+  geom_histogram(aes(group=singleton, y=after_stat(binwidth*density)), 
+                 position = "dodge", binwidth = binwidth) +
+  facet_wrap(~db, ncol = 5) +
+  scale_y_continuous(expand = expansion(0.01,0)) + 
+  scale_fill_manual(values=palette_singleton_p) +
+  theme_bw(base_family = "Aptos") +
+  labs(title="NA=mismatch", x="Jaccard Index", y="Scaled density", fill="Target sets") +
+  theme(legend.position = "none",
+        strip.background = element_rect(fill=NA))
+
 jaccard_df <- map_dfr(dbs, ~ {
@@
 
-ggplot(jaccard_df, aes(jaccard, fill=singleton)) +
-  geom_histogram(position = "dodge") +
-  facet_wrap(~db, scales = "free_y") +
+p_dbs <- ggplot(jaccard_df, aes(jaccard, fill=singleton)) +
+  geom_histogram(aes(group=singleton, y=after_stat(binwidth*density)), 
+                 position = "dodge", binwidth = binwidth) +
+  facet_wrap(~db, ncol = 5) +
   scale_y_continuous(expand = expansion(0.01,0)) + 
   scale_fill_manual(values=palette_singleton_p) +
-  theme_bw() +
+  theme_bw(base_family = "Aptos") +
+  labs(title="Default", x="Jaccard Index", y="Scaled density", fill="Target sets") +
   theme(legend.position = "top",
-        # panel.background = element_rect(fill=NA, color="black"),
-        # axis.line = element_blank()
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(), 
-        strip.background = element_rect(fill=NA)
-        # panel.grid.major.x = element_line()
-        )
+        strip.background = element_rect(fill=NA))
 
-ggsave("test/bench.pdf", width = 8, height = 5, device = cairo_pdf)
-# plot_dbs_list[[db]] <- p_db
+p_dbs / p_dbs_NA / p_dbs_mono
+
+# ggsave("test/bench.pdf", p_dbs, width = 8, height = 5, device = cairo_pdf)
 

### 2024-11-07 workflow/scripts/compare_sptree.R

@@ -121,4 +121,4 @@ reco_plot <- (plot_rf | plot_apro | plot_DL) +
   plot_layout(guides = "collect") &
   theme(legend.position = "bottom") 
 
-ggsave(snakemake@output[["reco"]], reco_plot, width = 7, height = 4)
+ggsave(snakemake@output[["reco"]], reco_plot, width = 8, height = 4)

### 2024-11-07 workflow/scripts/compare_trees.R

@@ -26,15 +26,18 @@ df_ml <- read_delim(snakemake@input[["mltrees"]], delim = "\t",
 
 # Disco similarity to sptree
 disco_fls <- c(snakemake@input[["disco"]])
+# disco_fls <- list.files("results/phylogeny/eggnog_benchmark/reco/disco/", full.names = T, pattern = "out*")
 
 # Astral-pro results
 clades <- read_delim(snakemake@input[["table"]], show_col_types = FALSE)
 
 apro_files <- snakemake@input[["apro_trees"]]
+# apro_files <- list.files("results/phylogeny/eggnog_benchmark/reco/apro/", full.names = T, pattern = "supp*")
+apro_files <- apro_files[sapply(apro_files, function(x) file.size(x))>0]
+
 apro_trees <- read.tree(text = sapply(apro_files, readLines))
 names(apro_trees) <- sapply(str_split(basename(apro_files), "_"), 
                             function(x) paste0(x[2], "_", x[4]))
-
 
 # 1. Compute astral-pro quartet
 nodes_clades <- fortify(sptree) %>% 
@@ -210,7 +213,7 @@ rf_plot <- rf_df %>%
 
 
 # Reconciliation notung
-reco <- read_delim(snakemake@input[["reco"]], 
+reco <- read_delim("results/phylogeny/eggnog_benchmark/reco/UP000005640_notung.tsv", 
                      show_col_types = FALSE, col_names = c("gene", "dups", "losses")) %>% 
   separate(gene, c("gene", "target", "alphabet", "model")) %>% 
   left_join(var_df) %>% 

### 2024-11-07 workflow/scripts/ecod_exp.R

@@ -1,0 +1,5 @@
+ecod <- read_delim("~/Downloads/ecod.develop292.domains.txt", 
+                   comment = "#", delim = " ", col_names = FALSE)
+df_filt$target
+indb <- gsub("AF-", "", gsub("-F1", "_F1", unique(names(db_fasta))))
+length(intersect(indb, ecod$X1))/length(indb)

### 2024-11-07 workflow/scripts/functions.R

@@ -10,9 +10,11 @@ fs_columns <- c("query","target","pident","length","mismatch","gapopen",
                 "qstart","qend","sstart","send","evalue","bitscore",
                 "lddt","alntmscore","rmsd","prob","qcov","tcov")
 
-models <- c("LG", "FM", "FT", "FTPY", "3Di", "GTR", "AF", "LLM")
+models <- c("LG", "FM", "FT", "FTPY", 
+            "3Di", "GTR", "AF", "LLM", "part")
 
-palettes_model <- c("#BF360C", "#FFA000", "#D4E157","#646B00", "#75C323", "#40A79D", "#33ceff", "#3396ff")
+palettes_model <- c("#BF360C", "#FFA000", "#D4E157","#646B00",
+                    "#75C323", "#40A79D", "#33ceff", "#3396ff", "purple")
 names(palettes_model) <- models
 palettes_model_d <- colorspace::darken(palettes_model, amount = 0.6)
 names(palettes_model_d) <- models

### 2024-11-07 workflow/scripts/get_QFO_sptree.R

@@ -1,25 +1,0 @@
-library(ape)
-
-t <- read.nhx("resources/species_tree_QFO2020.nx")
-t <- TreeTools::Preorder(t@phylo)
-t$tip.label <- gsub("_.*", "", t$tip.label)
-t <- TreeTools::Subtree(t, getMRCA(t, tip = c("HUMAN", "GIAIC")))
-t$tip.label <- gsub("CANFA", "CANLF", t$tip.label)
-t <- TreeTools::AddTip(t, getMRCA(t, tip = c("DANRE", "ORYLA")), label = "LEPOC")
-t <- TreeTools::AddTip(t, getMRCA(t, tip = c("DROME", "BOMMO")), label = "TRICA")
-t <- TreeTools::AddTip(t, where = "TRIVA", label = "TRIV3")
-t <- TreeTools::AddTip(t, where = "SCHMA", label = "HELRO")
-t <- TreeTools::AddTip(t, where = "SORBI", label = "MAIZE")
-# Some species are not found in Euka dataset and this causes a bug in A-Pro
-# mammals <- c("PONAB", "NOMLE", "MACMU",
-#              "CALJA", "OTOGA", "CAVPO",
-#              "SPETR", "RABIT", "PIG", 
-#              "HORSE", "CANLF", "AILME", "MYOLU", "LOXAF", "SARHA", "ORNAN", "CHICK")
-# t <- TreeTools::DropTip(t, mammals)
-write.tree(t, "data/sptrees/QFO.nwk")
-
-# p <- ggtree(t) + 
-#   geom_tiplab() + 
-#   xlim(c(0,30))
-# 
-# ggsave("test/qfo.pdf", width = 8, height = 12)

### 2024-11-07 workflow/scripts/get_sample_benchmark.R

@@ -1,0 +1,58 @@
+library(tidyverse)
+library(glue)
+
+n_sample <- 500
+min_seqs <- 4
+
+params <- yaml::read_yaml("config/params.yaml")
+params_dataset <- yaml::read_yaml("config/Hsap_draft.yaml")
+tax <- read_delim(glue("results/data/meta/",
+                       params_dataset$homology_dataset, "_uniprot_genomes.tsv"), 
+                  show_col_types=F)
+
+# some sequences do not have structures so to simplify our lives 
+# we only get those ortthogroups with at least 4 existying files
+db_fasta <- Biostrings::readAAStringSet("results/homology/hsap_euka/db/all_seqs_fsdb_ss.fa")
+
+dbs <- read_delim(paste0(params$data_dir, "ids/", tax$Proteome_ID, "_ids.tsv"),
+           delim = "\t", id = "proteome", show_col_types = FALSE) %>% 
+  mutate(proteome = gsub("_ids.tsv", "", basename(proteome)), 
+         target=paste0("AF-", Entry, "-F1")) %>% 
+  select(target, OMA, eggNOG) %>% 
+  filter(target %in% names(db_fasta)) %>% 
+  # as one id can have multiple eggnog we need to unnest
+  mutate(eggNOG=strsplit(eggNOG, ";"),
+         OMA=gsub(";", "", OMA)) %>%
+  unnest(eggNOG)
+
+eggNOGs <- dbs %>% 
+  select(target, eggNOG) %>% 
+  filter(!is.na(eggNOG)) %>% 
+  group_by(eggNOG) %>% 
+  summarise(dist=n_distinct(target)) %>% 
+  filter(dist>min_seqs) %>% 
+  pull(eggNOG) %>% 
+  sample(n_sample, replace = FALSE)
+
+dbs %>% 
+  filter(eggNOG %in% eggNOGs) %>% 
+  select(eggNOG, target) %>% 
+  arrange(eggNOG) %>% 
+  write_delim("data/seeds/eggnog_benchmark.txt", col_names = FALSE, delim = "\t")
+
+OMAs <- dbs %>% 
+  select(target, OMA) %>% 
+  filter(!is.na(OMA)) %>% 
+  group_by(OMA) %>% 
+  summarise(dist=n_distinct(target)) %>% 
+  filter(dist>min_seqs) %>% 
+  pull(OMA) %>% 
[…]

### 2024-11-07 workflow/scripts/get_sptrees.R

@@ -1,0 +1,51 @@
+library(ape)
+
+# For our sampling some species are not in eggnog or OMA, we can prune them 
+# from the sptree.
+params <- yaml::read_yaml("config/params.yaml")
+params_dataset <- yaml::read_yaml("config/Hsap_draft.yaml")
+tax <- read_delim(glue("results/data/meta/",
+                       params_dataset$homology_dataset, "_uniprot_genomes.tsv"), 
+                  show_col_types=F)
+taxmap <- read_delim(glue(homodir, "/db/taxidmap"), 
+                     delim = "\t", col_names = c("label", "Tax_ID"),
+                     show_col_types = FALSE) %>% 
+  left_join(tax, by = "Tax_ID") %>% 
+  dplyr::select(label, Proteome_ID, mnemo)
+
+euka_tree <- read.tree(params_dataset$species_tree)
+
+dbs <- read_delim(paste0(params$data_dir, "ids/", tax$Proteome_ID, "_ids.tsv"),
+                  delim = "\t", id = "proteome", show_col_types = FALSE) %>% 
+  mutate(proteome = gsub("_ids.tsv", "", basename(proteome))) %>% 
+  mutate(target=paste0("AF-", Entry, "-F1")) %>% 
+  select(-proteome, -Entry)
+
+missing_df <- dbs %>%
+  right_join(taxmap, by=c("target"="label")) %>%
+  group_by(mnemo) %>%
+  select(mnemo, eggNOG, OMA) %>% 
+  summarise_all(~ sum(is.na(.))/n())
+
+oma_missing <- missing_df[missing_df$OMA==1,]$mnemo
+eggnog_missing <- missing_df[missing_df$eggNOG==1,]$mnemo
+
+oma_tree <- TreeTools::DropTip(euka_tree, oma_missing)
+write.tree(oma_tree, "data/sptrees/homo_oma.nwk")
+eggnog_tree <- TreeTools::DropTip(euka_tree, eggnog_missing)
+write.tree(eggnog_tree, "data/sptrees/homo_eggnog.nwk")
+
+# get QFO species tree and do some small modifications for the 
+# Foldtree benchmark
+t <- treeio::read.nhx("resources/species_tree_QFO2020.nx")
+t <- TreeTools::Preorder(t@phylo)
+t$tip.label <- gsub("_.*", "", t$tip.label)
[…]

### 2024-11-29 workflow/notebooks/draft.Rmd
added orthobenchmark and finally resplit workflow by set

--- 2024-11-06 workflow/notebooks/draft.Rmd
@@
 source("workflow/scripts/functions.R")
-theme_set(theme_classic(base_family = "Aptos"))
+theme_set(theme_classic(base_family = "Helvetica"))
 
@@
 
+df %>% 
+  group_by(singleton) %>% 
+  summarise(n(), n()/nrow(df))
+
+df_filt %>% 
+  group_by(singleton) %>% 
+  summarise(n(), n()/nrow(df_filt))
+
 # blast_filtered <- filter(blast, 
@@
 #                                .default = "Common"))
+
+# seed_fa <- Biostrings::readAAStringSet("results/data/fastas/UP000005640.fa")
+# length_df <- Biostrings::width(seed_fa)
+# names(length_df) <- names(seed_fa)
+
 ```
@@
 
+threed_df <- read_delim(paste0(params$data_dir, "ids/", tax$Proteome_ID, "_3ds.tsv"),
+           delim = "\t", id = "proteome", show_col_types = FALSE) %>% 
+  mutate(proteome = gsub("_ids.tsv", "", basename(proteome))) %>% 
+  mutate(target=paste0("AF-", Entry, "-F1")) %>% 
+  select(-proteome, -Entry) %>% 
+  left_join(taxmap, by=c("target"="label")) %>% 
+  mutate(sp = gsub("..*_", "", `Entry Name`))
+
 ortho_dbs <- read_delim(paste0(params$data_dir, "ids/", tax$Proteome_ID, "_ids.tsv"),
@@
 ortho_dbs <- full_join(CATH_df, ortho_dbs)
+
+# ortho_dbs %>% 
+#   left_join(taxmap, by=c("target"="label")) %>% 
+#   group_by(mnemo) %>% 
+#   summarise_all(~ sum(is.na(.))/nrow(ortho_dbs)) %>% 
+#   filter(!is.na(mnemo)) %>% 
+#   select(-target) %>% 
+#   pivot_longer(!mnemo) %>% 
+#   ggplot(aes(name, value)) + 
+#   geom_boxplot() + 
+#   labs(x="DB", y="% NA")
 
@@
   ggplot(aes(mean_lddt)) + 
-  geom_density(fill="#BF4375", size=1) +
+  geom_density(fill=NA, size=1) +
   scale_y_continuous(expand = expansion(mult = c(0, .1))) +
@@
 
-# ggsave("draft/figures/fig1_lddt.pdf", fig1b, width = 3, height = 4)
-# ggsave("draft/figures/fig1_cnts.pdf", fig1a, width = 4, height = 2)
+fig1 <- (fig1a | fig1b) + plot_layout(widths = c(1,2))
+
+ggsave("draft/figures/Fig1_tmp.pdf", fig1, width = 8, height = 4)
 ```
@@
   # geom_vline(xintercept = -log10(evals), linetype=2) + 
-  coord_fixed(expand = 0) +
+  coord_cartesian(expand = 0) +
   labs(x="-log10(Evalue BlastP)",y="-log10(Evalue Foldseek)") +
@@
   # scale_fill_viridis_c(option = "C") +
-  coord_fixed(expand = 0) +
+  coord_cartesian(expand = 0) +
   scale_x_continuous(limits = c(0, 100)) +
@@
   scale_fill_gradientn(colors=wespal) +
-  coord_fixed(expand = 0) +
+  coord_cartesian(expand = 0) +
   theme(legend.position = "none", panel.grid = element_blank()) 
@@
   geom_density() + 
-  # scale_y_continuous(expand = c(0.01,0)) +
-  # scale_x_continuous(limits = c(params$low_confidence, 100),
-  #                    expand = c(0,0)) +
-  scale_color_manual(values = palette_singleton_p) + 
-  labs(x="LDDT") + 
+  scale_y_continuous(expand = expansion(c(0.02,0.02))) +
+  scale_x_continuous(limits = c(0, 1), 
+                     expand = expansion(c(0.02,0.02))) +
+  scale_color_manual(values = palette_singleton_p, guide=FALSE) + 
+  labs(x="LDDT", color = "Target sets") + 
   theme(legend.position = "none", 
@@
   # coord_cartesian(expand = 0) +
-  labs(x="TM-Score") + 
-  scale_color_manual(values = palette_singleton_p)+ 
+  scale_y_continuous(expand = expansion(c(0.02,0.02))) +
+  scale_x_continuous(limits = c(0, 1), 
+                     expand = expansion(c(0.02,0.02))) +
+  labs(x="TM-Score", color = "Target sets") + 
+  scale_color_manual(values = palette_singleton_p, guide=FALSE)+ 
   theme(legend.position = "none", 
@@
   geom_density(bw=1) + 
-  # scale_y_continuous(expand = c(0.01,0)) +
-  scale_x_continuous(limits = c(params$low_confidence, 100)) +
-  scale_color_manual(values = palette_singleton_p) +   
-  labs(x="Mean target pLDDT") + 
+  scale_y_continuous(expand = expansion(c(0.02,0.02))) +
+  scale_x_continuous(limits = c(params$low_confidence, 100), 
+                     expand = expansion(c(0.02,0.02))) +
+  scale_color_manual(values = palette_singleton_p, guide=FALSE) +   
+  labs(x="Mean target pLDDT", color = "Target sets") + 
   theme(legend.position = "none", 
@@
 dbs <- colnames(ortho_dbs)[2:dim(ortho_dbs)[2]]
+dbs <- setdiff(dbs, c("InParanoid"))
+dbs <- dbs[2:5]
+
+binwidth <- 0.1
+
 # Loop through each column name and apply a function
+# jaccard_df_mono <- map_dfr(dbs, ~ {
+#   col_name <- .x
+#   select(df_filt, query, target, singleton) %>% 
+#     left_join(select(ortho_dbs, target, col_name), by=c("query"="target")) %>% 
+#     left_join(select(ortho_dbs, target, col_name), by=c("target")) %>% 
+#     rename("query_set"=paste0(col_name, ".x"), "target_set"=paste0(col_name, ".y")) %>% 
+#     mutate(n_domain_q = str_count(pattern = "\\;", query_set)) %>% 
+#     # Get only monodomain queries
+#     filter(n_domain_q==1) %>% 
+#     filter(!is.na(query_set), !is.na(target_set)) %>% 
+#     mutate(
+#       jaccard = map2_dbl(strsplit(query_set, ";"), strsplit(target_set, ";"),
+#                          ~ length(intersect(.x, .y)) / length(union(.x, .y)))
+#     ) %>% 
+#     group_by(query, singleton) %>% 
+#     summarise(avg_Jaccard = mean(jaccard))  %>% 
+#     mutate(db=col_name)
+# })
+# 
+# jaccard_df_NA <- map_dfr(dbs, ~ {
+#   col_name <- .x
+#   select(df_filt, query, target, singleton) %>%
+#     left_join(select(ortho_dbs, target, col_name), by=c("query"="target")) %>%
+#     left_join(select(ortho_dbs, target, col_name), by=c("target")) %>%
+#     rename("query_set"=paste0(col_name, ".x"), "target_set"=paste0(col_name, ".y")) %>%
+#     # If target has na no we count mismatch
+#     filter(!is.na(query_set)) %>%
+#     mutate(
+#       jaccard = map2_dbl(strsplit(query_set, ";"), strsplit(target_set, ";"),
+#                          ~ length(intersect(.x, .y)) / length(union(.x, .y)))
+#     )  %>%
+#     group_by(query, singleton) %>%
+#     summarise(avg_Jaccard = mean(jaccard)) %>%
+#     mutate(db=col_name)
+# 
+# })
+
 jaccard_df <- map_dfr(dbs, ~ {
@@
   select(df_filt, query, target, singleton) %>% 
-    # filter(query=="AF-A0A024RBG1-F1") %>%
     left_join(select(ortho_dbs, target, col_name), by=c("query"="target")) %>% 
@@
     rename("query_set"=paste0(col_name, ".x"), "target_set"=paste0(col_name, ".y")) %>% 
-    # Exclude queries without any domain
+    # If any of query or target has na no we dont count
     filter(!is.na(query_set), !is.na(target_set)) %>% 
+    mutate(
+      jaccard = map2_dbl(strsplit(query_set, ";"), strsplit(target_set, ";"),
+                         ~ length(intersect(.x, .y)) / length(union(.x, .y)))
+    )  %>% 
     group_by(query, singleton) %>% 
-    distinct(query, singleton, query_set, target_set) %>% 
-    summarise(set_q = strsplit(paste(query_set, collapse = ""), ";"),
-              set_t = strsplit(paste(target_set, collapse = ""), ";")) %>%
-    rowwise() %>%
-    mutate(jaccard = length(intersect(set_q, set_t))/length(union(set_q, set_t))) %>% 
-    select(-set_q, -set_t) %>% 
+    summarise(avg_Jaccard = mean(jaccard)) %>% 
     mutate(db=col_name)
@@
 
-ggplot(jaccard_df, aes(jaccard, fill=singleton)) +
-  geom_histogram(position = "dodge") +
-  facet_wrap(~db, scales = "free_y") +
-  scale_y_continuous(expand = expansion(0.01,0)) + 
-  scale_fill_manual(values=palette_singleton_p) +
-  theme_bw() +
-  theme(legend.position = "top",
-        # panel.background = element_rect(fill=NA, color="black"),
-        # axis.line = element_blank()
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(), 
-        strip.background = element_rect(fill=NA)
-        # panel.grid.major.x = element_line()
-        )
+p_dbs <- ggplot(jaccard_df, aes(avg_Jaccard, 1, color=singleton, alpha=db)) +
+  stat_ecdf(pad=FALSE) +
+  scale_color_manual(values=palette_singleton_p) +
+  # scale_linetype_manual(values = c(2,2,2,2,1)) +
+  scale_alpha_manual(values = seq(0.3, 1, length.out=4)) +
+  coord_fixed() +
+  scale_y_continuous(expand = expansion(0.0,0)) +  
+  scale_x_continuous(expand = expansion(add = c(0,0.01))) +  
+  guides(colour = guide_legend(order = 1),  
+         alpha = guide_legend(order = 2)) +
+  theme(panel.border = element_rect(fill = NA), 
+        legend.position = "bottom",
+        panel.grid.major = element_line(linetype = 2)) + 
+  # theme_bw(base_family = "Aptos") +
+  labs(x="Mean Jaccard Index", y="ECDF", 
+       alpha="DB", linetype="DB", color="Target sets")
 
-ggsave("test/bench.pdf", width = 8, height = 5, device = cairo_pdf)
-# plot_dbs_list[[db]] <- p_db
+# p_dbs_mono <-  ggplot(jaccard_df_mono, aes(avg_Jaccard, 1, color=singleton, 
+#                                 linetype=db, alpha=db)) +
+#   stat_ecdf(pad=FALSE) +
+#   scale_color_manual(values=palette_singleton_p) +
+#   scale_linetype_manual(values = c(2,2,2,2,1)) +
+#   scale_alpha_manual(values = c(0.4,0.6,0.8,1,1)) +
+#   coord_fixed() +
+#   scale_y_continuous(expand = expansion(0.02,0)) +  
+#   scale_x_continuous(expand = expansion(0.02,0)) +  
+#   theme_bw(base_family = "Aptos") +
+#   labs(title="Only Monodomain queries", x="Mean Jaccard Index", y="ECDF", 
+#        alpha="DB", linetype="DB", color="Target sets")
+# 
+# p_dbs_NA <-  ggplot(jaccard_df_NA, aes(avg_Jaccard, 1, color=singleton,
+#                                 linetype=db, alpha=db)) +
+#   stat_ecdf(pad=FALSE) +
+#   scale_color_manual(values=palette_singleton_p) +
+#   scale_linetype_manual(values = c(2,2,2,2,1)) +
+#   scale_alpha_manual(values = c(0.4,0.6,0.8,1,1)) +
+#   coord_fixed() +
+#   scale_y_continuous(expand = expansion(0.02,0)) +
+#   scale_x_continuous(expand = expansion(0.02,0)) +
+#   theme_bw(base_family = "Aptos") +
+#     labs(title="NA=mismatch", x="Mean Jaccard Index", y="ECDF", 
+#        alpha="DB", linetype="DB", color="Target sets")
 
-# df_multi <- df_filt %>% 
-#   select(query, target, singleton) %>% 
-#     head(n = 100) %>%
-#   left_join(CATH_df, by=c("query"="target")) %>% 
-#   filter(!is.na(Pfam), !is.na(Gene3D)) %>% 
-#   left_join(CATH_df, by="target") %>% 
-#   rowwise() %>% 
-#   mutate(Gene3D.x=strsplit(Gene3D.x, ";"),
-#          Pfam.x=strsplit(Pfam.x, ";"),
-#          Gene3D.y=strsplit(Gene3D.y, ";"),
-#          Pfam.y=strsplit(Pfam.y, ";")) %>% 
-#   mutate(n_3d_common=length(intersect(Gene3D.x,Gene3D.y)),
-#          n_Pfam_common=length(intersect(Pfam.x,Pfam.y))) %>% 
-#   # filter(n_Pfam_common>0, n_3d_common>0) %>% 
-#   select(-c(Pfam.x,Pfam.y,Gene3D.x,Gene3D.y)) %>% 
-#   mutate(Jaccard_3d=n_3d_common/pmax(n_3d_domains.x,n_3d_domains.y),
-#          Jaccard_Pfam=n_Pfam_common/pmax(n_Pfam_domains.x,n_Pfam_domains.y)) %>% 
-#   select(query, target, singleton, contains("Jaccard"), contains("common"))
-
-CATH_plot <- df_multi %>% 
-    pivot_longer(contains("Jaccard"), values_to = "Jaccard") %>% 
-    mutate(name=gsub("Jaccard_", "", name),
-           name=gsub("3d", "CATH", name),
-           Jaccard=cut_width(Jaccard, .1, boundary = 0)) %>% 
-  filter(!is.na(Jaccard)) %>% 
-  group_by(name, singleton, Jaccard) %>% 
-  summarise(n=n()) %>% 
-  mutate(freq=n/sum(n)) %>%
-  ggplot(aes(Jaccard, freq, fill=singleton)) + 
-  geom_bar(stat = "identity",position = "dodge") + 
-  facet_grid(name~.) +
-  labs(y="% in Singleton group", x="Jaccard Index", 
-       fill="Target sets") +
-  scale_y_continuous(expand = expansion(0.01,0)) + 
-  scale_fill_manual(values=palette_singleton_p) +
-  theme(legend.position = "top",
-        panel.background = element_rect(fill=NA, color="black"),
-        axis.line = element_blank(),
-        axis.text.x = element_text(size=8))
-
-# prop_p <- group_by(df_filtered) %>% 
-#   count(query, singleton) %>% 
-#   group_by(query) %>% 
-#   mutate(nn=sum(n), prop=n/nn) %>% 
-#   ggplot(aes(prop, singleton, color=singleton)) + 
-#   geom_boxplot() + 
-#   labs(x="Proportion in tree", y="") +
-#   scale_color_manual(values = palette_singleton_p) + 
-#   theme(legend.position = "none")
+# jaccard_plot <- (p_dbs_NA | p_dbs_mono) + 
+#   plot_layout(guides = "collect", 
+#               axis_titles = "collect") & theme(legend.position = 'bottom')
 
 fig2_first_row <- (pident_p | eval_p | qcov_p)
-fig2_second_row <- ((lddt_p / alntmscore_p / plddt_p) | CATH_plot) + plot_layout(widths = c(1,2))
+fig2_second_row <- ((lddt_p / alntmscore_p / plddt_p) | p_dbs) + 
+  plot_layout(guides = "collect") & 
+  theme(legend.position = 'bottom')
 fig2 <- (fig2_first_row / fig2_second_row) +
+  plot_layout(heights = c(1,1.3)) +
   plot_annotation(tag_levels = 'A') &
-  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
+  theme(plot.tag = element_text(size = 10, face = "bold", family = "Aptos"))
 
-ggsave("draft/figures/fig2.pdf", fig2, width = 10.5, height = 8)
+ggsave("draft/figures/Fig2.pdf", fig2, width = 8, height = 7, device = cairo_pdf)
+# ggsave("draft/figures/Fig_Supp_CATH.pdf", jaccard_plot, width = 8, height = 4, device = cairo_pdf)
 ```
@@
 ```{r read_trees}
-df_trees <- read_delim(glue(outdir,"/trees/", params_dataset$seed, "_unrooted_trees.txt"),
+glue("results/phylogeny/", params_dataset$phylo_dataset,"/trees/", params_dataset$seed, "_unrooted_trees.txt")
+
+df_trees <- read_delim("results/old/test/trees/UP000005640_unrooted_trees.txt",
                        show_col_types = FALSE,
                        col_names = c("gene", "target", "alphabet", "model", "tree")) %>% 
-  mutate(model=factor(model, levels=models)) %>% 
-  filter(model!="FTPY")
+  mutate(model=factor(model, levels=models))  
+  # filter(model!="FTPY")
 
 # ML trees with selected model and LL
-df_ml <- read_delim(glue(outdir,"/trees/", params_dataset$seed, "_mltrees.txt"), 
+glue("results/phylogeny/", params_dataset$phylo_dataset,"/trees/", params_dataset$seed, "_mltrees.txt")
+df_ml <- read_delim("results/old/test/trees/UP000005640_mltrees.txt", 
                     delim = "\t", 
@@
 
-disco_fls <- list.files(glue(outdir, "/reco/disco"), pattern = "output.nwk", full.names = T)
+glue("results/phylogeny/", params_dataset$phylo_dataset, "/reco/disco")
+
+disco_fls <- list.files("results/old/test/reco/disco/", pattern = "output.nwk", full.names = T)
 
@@
 # Reconciliation ranger+TCS scores
-reco <- read_delim(glue(outdir,"/reco/", params_dataset$seed, "_notung.tsv"), 
+glue("results/phylogeny/", params_dataset$phylo_dataset, "/reco/", params_dataset$seed, "_notung.tsv")
+reco <- read_delim("results/old/test/reco/UP000005640_notung.tsv", 
                    show_col_types = FALSE, col_names = c("gene", "dups", "losses")) %>% 
@@
   filter(ntips>10) %>% 
-  # filter(ntips>=length(sptree$tip.label)/2) %>%
   separate(bn, c("seed", "targets", "alphabet", "model")) %>% 
@@
 
-apro_files <- list.files(glue(outdir, "/reco"), 
+apro_files <- list.files("results/old/test/reco/", 
                          pattern = "*support.nwk", full.names = TRUE)
+
 apro_trees <- read.tree(text = sapply(apro_files, readLines))
@@
 
-df_red <- select(df_filtered, query, target, singleton)
+df_red <- select(df_filt, query, target, singleton)
 
@@
 ```{r FigS2}
-pp1_plot <- nodes_df %>%
-    filter(model!="FTPY") %>% 
-    ggplot(aes(q1, label_int, fill=model, color=model, group=model)) +
-    geom_linerange(aes(x=x, xmin=xmin, xmax=xmax, color=model),
-                   data = . %>% 
-                     group_by(model, label_int) %>% 
-                     summarise(x=mean(q1), xmin=min(q1), xmax=max(q1)),
-                   position = position_dodge(width=.7), color="grey70") +
-    geom_point(aes(shape=targets),
-               position = position_dodge(width=.7)) +
-  scale_fill_manual(values=palettes_model) +
-  scale_color_manual(values = palettes_model_d) +
-  scale_shape_manual(values = c(21,22,23,24)) +
-  labs(y="Node") +
-  guides(color = guide_legend(override.aes = list(pch = 21, size=3)),
-         shape = guide_legend(override.aes = list(size=3))) +
-  theme(legend.position = "bottom", legend.box="vertical") + 
-  theme_bw()
+# pp1_plot <- nodes_df %>%
+#     filter(model!="FTPY") %>% 
+#     ggplot(aes(q1, label_int, fill=model, color=model, group=model)) +
+#     geom_linerange(aes(x=x, xmin=xmin, xmax=xmax, color=model),
+#                    data = . %>% 
+#                      group_by(model, label_int) %>% 
+#                      summarise(x=mean(q1), xmin=min(q1), xmax=max(q1)),
+#                    position = position_dodge(width=.7), color="grey70") +
+#     geom_point(aes(shape=targets),
+#                position = position_dodge(width=.7)) +
+#   scale_fill_manual(values=palettes_model) +
+#   scale_color_manual(values = palettes_model_d) +
+#   scale_shape_manual(values = c(21,22,23,24)) +
+#   labs(y="Node") +
+#   guides(color = guide_legend(override.aes = list(pch = 21, size=3)),
+#          shape = guide_legend(override.aes = list(size=3))) +
+#   theme(legend.position = "bottom", legend.box="vertical") + 
+#   theme_bw()
+# 
+# labeled_sptree <- sptree
+# labeled_sptree$tip.label <- tax$Species[match(labeled_sptree$tip.label, tax$mnemo)]
+# 
+# sptree_plot <- ggtree(labeled_sptree) + 
+#   geom_tiplab(aes(label=sub("[^A-Z].+? ", ". ", label)),fontface="italic",
+#               align = TRUE) + 
+#   geom_nodelab(color="firebrick", 
+#                nudge_x = -.15, hjust = 1, 
+#                nudge_y = .3, size=3.5)+
+#   xlim(c(-4, 21))
+# 
+# fig_s2 <- (sptree_plot + pp1_plot) +
+#   plot_layout(widths = c(1, 1.5)) +
+#   plot_annotation(tag_levels = 'A') &
+#   theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
+# ggsave("draft/figures/figS2.pdf", fig_s2, width = 12, height = 6, device = cairo_pdf)
+```
 
-labeled_sptree <- sptree
-labeled_sptree$tip.label <- tax$Species[match(labeled_sptree$tip.label, tax$mnemo)]
+# Review plots
 
-sptree_plot <- ggtree(labeled_sptree) + 
-  geom_tiplab(aes(label=sub("[^A-Z].+? ", ". ", label)),fontface="italic",
-              align = TRUE) + 
-  geom_nodelab(color="firebrick", 
-               nudge_x = -.15, hjust = 1, 
-               nudge_y = .3, size=3.5)+
-  xlim(c(-4, 21))
+### Experimental vs predicted
 
-fig_s2 <- (sptree_plot + pp1_plot) +
-  plot_layout(widths = c(1, 1.5)) +
-  plot_annotation(tag_levels = 'A') &
-  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
-ggsave("draft/figures/figS2.pdf", fig_s2, width = 12, height = 6, device = cairo_pdf)
+```{r why_no_exp}
+
+plot_PDB <- threed_df %>%
+  group_by(mnemo) %>% 
+  filter(!is.na(mnemo)) %>% 
+  summarise_all(~ (1-sum(is.na(.))/n())*100) %>% 
+  left_join(select(tax, mnemo, Species)) %>% 
+  ggplot(aes(PDB, fct_reorder(Species, PDB))) + 
+  geom_text(aes(label=signif(PDB,2)), family="Aptos", 
+            size=2.5, nudge_x = 1, hjust = 0) + 
+  geom_bar(stat="identity") + 
+  scale_x_continuous(limits = c(0,100), expand = expansion(add = c(0,2))) +
+  labs(y="Species", x="% Proteins with PDB entry") + 
+  theme(axis.text.y = element_text(face = "italic"))
+
+ggsave("draft/figures/review_nopdb.png", plot_PDB, width = 4, height = 3, dpi = 600)
 ```
+
+### Paralogy issue in phylome
+
+```{r}
+# find results/phylogeny/hsap_1kseeds/seeds/UP000005640/ -name "*.ids" -print0 | while IFS= read -r -d '' file; do  awk -v fname="$(basename "$file" | sed 's/.ids//' | sed 's/_/\t/')"  '{print $0 "\t" fname}' "$file"; done > test/paranoia/ids_by_set.txt
+df_ids <- read_delim("draft/figures/ids_by_set.txt", 
+                 col_names = c("target", "query", "set")) %>% 
+  filter(query %in% aln_seeds)
+
+overlap_df <- filter(df_ids, query!=target) %>% 
+  group_by(query, set) %>% 
+  summarise(targets=list(target)) %>% 
+  # head(100) %>%
+  cross_join(., .) %>% 
+  filter(set.x==set.y, query.x!=query.y) %>% 
+  rowwise() %>% 
+  mutate(n =  length(union(targets.x, targets.y)), 
+         overlap = length(intersect(targets.x, targets.y))/n) %>% 
+  select(query.x, query.y, overlap, n, set.x) %>% 
+  filter(query.x>query.y)
+
+overlap_df %>% 
+  mutate(is_not0 = overlap>0) %>% 
+  count(set.x, is_not0) %>% 
+  pivot_wider(names_from = is_not0, values_from = n) %>% 
+  mutate(proportion = `TRUE`/(`FALSE`+`TRUE`)*100)
+
+```

### 2024-11-29 workflow/scripts/analyze_orthobench.R

@@ -1,0 +1,120 @@
+suppressPackageStartupMessages(library(tidyverse))
+library(patchwork)
+source("workflow/scripts/functions.R")
+theme_set(theme_classic())
+
+ids_df <- read_delim(snakemake@input[["ids"]], 
+                     show_col_types = FALSE, delim = "\t")
+# ids_df <- read_delim(paste0("results/data/ids/", tax$Proteome_ID, "_ids.tsv"))
+
+eval <- snakemake@params[["evalue"]]
+# eval <- 1e-3
+min_size <- snakemake@params[["min_seqs"]]
+# min_size <- 5
+dataset <- snakemake@wildcards[["db"]]
+# dataset <- "oma"
+
+seed_fa <- Biostrings::readAAStringSet(snakemake@input[["fa"]])
+# seed_fa <- Biostrings::readAAStringSet("results/ortho_bench/eggnog/db/db.fa")
+
+length_df <- Biostrings::width(seed_fa)
+names(length_df) <- names(seed_fa)
+
+colnames(ids_df) <- str_to_lower(colnames(ids_df))
+
+true_set <- ids_df %>% 
+  mutate(entry = paste0("AF-", entry, "-F1")) %>% 
+  filter(entry %in% names(length_df)) %>%
+  select(entry, match(dataset, colnames(ids_df))) %>% 
+  rename(dataset = 2) %>% 
+  rowwise() %>%
+  mutate(dataset = strsplit(dataset, ";")) %>%
+  unnest(dataset)
+
+# fs_bench <- read_delim("results/ortho_bench/eggnog//hits_fs.tsv", col_names = fs_columns) %>%
+#   filter(evalue < eval, query != target) %>%
+#   mutate(method = "fs", qcov = qcov * 100)
+fs_bench <- read_delim(snakemake@input[["fs"]], col_names = fs_columns, 
+                       show_col_types = FALSE, delim = "\t") %>% 
+  filter(evalue < eval, query != target) %>% 
+  mutate(method = "fs", qcov = qcov * 100)
+
+# blast_bench <- read_delim("results/ortho_bench/eggnog/hits_blast.tsv", col_names = blast_columns)  %>%
+#   filter(evalue < eval, query != target) %>%
+#   mutate(method = "blast")
+blast_bench <- read_delim(snakemake@input[["blast"]], col_names = blast_columns, 
+                          show_col_types = FALSE, delim = "\t")  %>% 
[…]

### 2024-11-29 workflow/scripts/analyze_runtimes.R

@@ -3,13 +3,59 @@ library(patchwork)
 source("workflow/scripts/functions.R")
 theme_set(theme_bw())
 
-df <- read_delim(snakemake@input[["time"]], 
-                 show_col_types = FALSE) %>% 
-  mutate(basename=gsub(".txt", "", basename),
-         step=basename(dirname),
-         max_rss_gb = max_rss/1024) %>% 
-  separate(basename, c("seed", "gene", "method", "alphabet", "model"))
-
+# all_fls <- list.files("results/phylogeny/hsap_1kseeds/benchmarks/", full.names = T, 
+#            include.dirs = F, pattern = "*txt", recursive = T) 
+# 
+# stats <- read_delim("tmp.stats") %>% 
+#   separate(file, c("gene", "targets", "alphabet"), sep = "_") %>%
+#   mutate(alphabet = gsub(".alg", "", alphabet)) %>% 
+#   separate(alphabet, c("alphabet", "clean")) %>% 
+#   filter(!is.na(clean))
+# 
+# tmp <- read_delim(all_fls[1])
+# df_tmp <- sapply(all_fls, function(x) readLines(x)[-1]) %>% 
+#   enframe() %>% 
+#   separate(value, colnames(tmp), convert = TRUE) 
+# 
+# df <- df_tmp %>% 
+#   mutate(method = basename(dirname(name)), name = basename(name),
+#          max_rss_gb = max_rss/1024) %>% 
+#   separate(name, c("seed", "gene", "targets", "alphabet", "model"), sep = "_") %>% 
+#   mutate(targets = gsub(".txt", "", targets),
+#          alphabet = gsub(".txt", "", alphabet),
+#          model = gsub(".txt", "", 
+#                       case_when(method=="foldtree" ~ NA,
+#                                 method=="fastme" ~ NA, 
+#                                 .default = model))) %>% 
+#   left_join(stats)
+# 
+# 
+# # Alignment
+# # df %>% 
+# #   filter(method %in% c("mafft", "foldmason")) %>% 
+# #   select(targets, alphabet, `h:m:s`, num_seqs, avg_len, method, max_rss_gb) %>% 
+# #   ggplot(aes(avg_len, `h:m:s`, color=targets)) + 
+# #   geom_point() + 
+# #   facet_grid(~method)
+# 
+# 
+# 
+# df <- read_delim(snakemake@input[["time"]], 
[…]

### 2024-11-29 workflow/scripts/compare_sptree.R

@@ -117,8 +117,31 @@ plot_rf <- disco_rf %>%
   labs(x="", y="Norm. RF to species tree (only if > 10 tips)", fill="Target sets")  +
   theme(legend.position = "none")
 
-reco_plot <- (plot_rf | plot_apro | plot_DL) + 
+rank_plot <- reco %>% 
+  group_by(gene, target) %>% 
+  mutate(rank = factor(dense_rank(n_events), levels = seq(0,9,0.5))) %>% 
+  ggplot(aes(rank, fill=model, alpha=target)) +
+  # geom_bar(position = "dodge") + 
+  geom_histogram(stat="count", position = "dodge") +
+  # facet_grid(target~.) +
+  scale_y_continuous(expand = expansion(c(0,0))) +
+  scale_fill_manual(values = palettes_model) + 
+  scale_alpha_manual(values = seq(1, 0.3, length.out=4))
+
+reco_plot <- ((plot_rf | plot_apro | plot_DL) / (rank_plot)) + 
   plot_layout(guides = "collect") &
   theme(legend.position = "bottom") 
 
-ggsave(snakemake@output[["reco"]], reco_plot, width = 8, height = 4)
+ggsave(snakemake@output[["reco"]], reco_plot, width = 8, height = 6)
+
+
+# nodes_df %>% 
+#   group_by(node, target) %>% 
+#   mutate(rank = factor(dense_rank(-q1), levels = seq(0,9,0.5))) %>% 
+#   ggplot(aes(rank, fill=model, alpha=target)) +
+#   geom_histogram(stat="count", position = "dodge") +
+#   # facet_grid(target~.) +
+#   scale_y_continuous(expand = expansion(c(0,0))) +
+#   scale_fill_manual(values = palettes_model) + 
+#   scale_alpha_manual(values = seq(1, 0.3, length.out=4))
+

### 2024-11-29 workflow/scripts/functions.R

@@ -10,10 +10,10 @@ fs_columns <- c("query","target","pident","length","mismatch","gapopen",
                 "qstart","qend","sstart","send","evalue","bitscore",
                 "lddt","alntmscore","rmsd","prob","qcov","tcov")
 
-models <- c("LG", "FM", "FT", "FTPY", 
+models <- c("LG", "FM", "QT",  "FT", "FTPY", 
             "3Di", "GTR", "AF", "LLM", "part")
 
-palettes_model <- c("#BF360C", "#FFA000", "#D4E157","#646B00",
+palettes_model <- c("#BF360C", "#FFA000", "#FFA000", "#D4E157","#646B00",
                     "#75C323", "#40A79D", "#33ceff", "#3396ff", "purple")
 names(palettes_model) <- models
 palettes_model_d <- colorspace::darken(palettes_model, amount = 0.6)
@@ -22,7 +22,8 @@ names(palettes_model_d) <- models
 
 palettes_method <- c("#00A08A", "#d63a3a", "#ffaa0e", "#6f63bb")
 names(palettes_method) <- c("common", "blast", "fs", "union")
-
+palettes_method_d <- colorspace::darken(palettes_method, amount = 0.3)
+names(palettes_method_d) <- c("common", "blast", "fs", "union")
 palette_singleton <- c("#00A08A", "#d63a3a", "#ffaa0e", "#6f63bb")
 names(palette_singleton) <- c("common", "only_blast", "only_fs", "union")
 

### 2024-11-29 workflow/scripts/get_ids.R

@@ -1,0 +1,38 @@
+suppressPackageStartupMessages(library(tidyverse))
+
+min_size <- snakemake@params[["min_seqs"]]
+dataset <- snakemake@wildcards[["db"]]
+
+tax <- read_delim(snakemake@input[["meta"]], 
+                  show_col_types=F)
+
+in_db <- read_delim(snakemake@input[["idmap"]], 
+                    show_col_types = FALSE,col_names = c("entry", "taxid"))
+
+# remove all those that are not in your dbs anyway!
+ids_df <- read_delim(snakemake@input[["ids"]], show_col_types = FALSE) 
+colnames(ids_df) <- str_to_lower(colnames(ids_df))
+
+ids_df <- ids_df %>% 
+  mutate(entry = paste0("AF-", entry, "-F1")) %>% 
+  left_join(in_db) %>% 
+  filter(!is.na(taxid)) %>% 
+  select(entry, match(dataset, colnames(ids_df))) %>% 
+  rename(dataset = 2) %>% 
+  filter(!is.na(dataset)) %>% 
+  rowwise() %>%
+  mutate(dataset = strsplit(dataset, ";")) %>%
+  unnest(dataset)
+
+
+# get groups with at least X members
+groups_df <- ids_df %>% 
+  count(dataset) %>% 
+  filter(n > min_size)
+
+ids_df %>%
+  right_join(groups_df) %>%  
+  pull(entry) %>% 
+  unique() %>% 
+  writeLines(snakemake@output[[1]])
+

### 2024-12-04 workflow/notebooks/draft.Rmd
added foldtree benchmark

@@
 library(glue)
+library(ggh4x)
 source("workflow/scripts/functions.R")
@@
   summarise(n(), n()/nrow(df_filt))
-
-# blast_filtered <- filter(blast, 
-#                          query %in% paste0("AF-", aln_seeds, "-F1"),
-#                          length/qlen*100>params$coverage, 
-#                          length/slen*100>params$coverage,
-#                          evalue<as.numeric(params$eval_both)) %>% 
-#   group_by(query) %>% 
-#   slice_head(n = params$max_seqs) %>% 
-#   ungroup()
-# 
-# fs_filtered <- filter(fs, 
-#                       query %in% paste0("AF-", aln_seeds, "-F1"),
-#                       qcov*100>params$coverage, 
-#                       tcov*100>params$coverage, 
-#                       evalue<as.numeric(params$eval_both)) %>% 
-#   group_by(query) %>% 
-#   slice_head(n = params$max_seqs) %>% 
-#   ungroup()
-# 
-# df_filtered <- full_join(blast_filtered, fs_filtered,
-#                 by = c("query", "target"), suffix=c("_Blast", "_Foldseek")) %>%
-#     filter(query!=target) %>% 
-#   left_join(self, by="query") %>%
-#   mutate(pident_Foldseek = 100*pident_Foldseek, 
-#          qcov_Foldseek = 100*qcov_Foldseek,
-#          singleton = case_when(is.na(evalue_Foldseek) ~ "only Blast",
-#                                is.na(evalue_Blast) ~ "only Foldseek", 
-#                                .default = "Common"))
-
-# seed_fa <- Biostrings::readAAStringSet("results/data/fastas/UP000005640.fa")
-# length_df <- Biostrings::width(seed_fa)
-# names(length_df) <- names(seed_fa)
-
 ```
@@
 ortho_dbs <- full_join(CATH_df, ortho_dbs)
-
-# ortho_dbs %>% 
-#   left_join(taxmap, by=c("target"="label")) %>% 
-#   group_by(mnemo) %>% 
-#   summarise_all(~ sum(is.na(.))/nrow(ortho_dbs)) %>% 
-#   filter(!is.na(mnemo)) %>% 
-#   select(-target) %>% 
-#   pivot_longer(!mnemo) %>% 
-#   ggplot(aes(name, value)) + 
-#   geom_boxplot() + 
-#   labs(x="DB", y="% NA")
-
-# nms <- read_delim("resources/cath-names.txt", 
-#                   comment = "#", delim = "    ",
-#                   col_names = c("Gene3D", "abbr", "name"), 
-#                   show_col_types = FALSE) %>% 
-#   select(-abbr) %>% 
-#   mutate(name=gsub("^:", "", name))
-# 
-# nms_pfam <- read_delim("resources/Pfam-A.clans.tsv", 
-#            col_names = c("Pfam", "clan", "boh", "abbr", "name"), 
-#            show_col_types = FALSE) %>% 
-#   select(-name, -clan, -boh)
-# 
-# long_CATH_df <- CATH_df %>% 
-#   # filter(target %in% unique(blast$query)) %>% 
-#   mutate(Gene3D=strsplit(Gene3D, ";")) %>% 
-#   unnest(Gene3D) %>% 
-#   separate(col = Gene3D, into = c("C", "A", "T", "H"), sep = "\\.", remove = FALSE) %>% 
-#   mutate(CA = paste(C,A,sep = "."), CAT=paste(C,A,T,sep = ".")) %>% 
-#   left_join(nms, by=c("C"="Gene3D")) %>%  
-#   left_join(nms, by=c("CA"="Gene3D"), suffix = c("_C", "_CA")) %>% 
-#   left_join(nms, by=c("CAT"="Gene3D"), suffix = c("_CA", "_CAT")) %>% 
-#   left_join(nms, by=c("Gene3D"), suffix = c("_CAT", "_CATH"))
-# 
-# long_CATH_nms <- long_CATH_df %>%
-#     select(-c(Entry, Pfam, n_Pfam_domains, target, n_3d_domains)) %>% 
-#     distinct()
 ```
@@
   geom_bar(aes(alpha=Filtering), stat = "identity", position = position_dodge2(reverse=TRUE)) +
-  # geom_text(aes(group=Filtering, label=paste0(round(prop, 2)*100, "%")),
-  #           position = position_dodge2(reverse = TRUE, width = 0.9),
-  #           vjust=1.3, color="white", size=3) +
   scale_fill_manual(values = palette_singleton_s, guide="none") +
@@
   ggpubr::stat_cor(label.x.npc = .025) +
-  # geom_hline(yintercept = -log10(evals), linetype=2) + 
-  # geom_vline(xintercept = -log10(evals), linetype=2) + 
   coord_cartesian(expand = 0) +
@@
   ggpubr::stat_cor(label.x.npc = .025) +
-  # scale_fill_viridis_c(option = "C") +
   coord_cartesian(expand = 0) +
@@
   ggpubr::stat_cor(label.x.npc = .025) +
-  # geom_hline(yintercept = min_covs, linetype=2) + 
-  # geom_vline(xintercept = min_covs, linetype=2) + 
   scale_x_continuous(limits = c(0, 100)) +
@@
   geom_density() + 
-  # coord_cartesian(expand = 0) +
   scale_y_continuous(expand = expansion(c(0.02,0.02))) +
@@
   plot_layout(guides = "collect") & 
-  theme(legend.position = 'bottom')
+  theme(legend.position = 'top')
 fig2 <- (fig2_first_row / fig2_second_row) +
@@
 ```{r read_trees}
-glue("results/phylogeny/", params_dataset$phylo_dataset,"/trees/", params_dataset$seed, "_unrooted_trees.txt")
-
-df_trees <- read_delim("results/old/test/trees/UP000005640_unrooted_trees.txt",
+models_draft <- setdiff(models, c("QT", "FTPY", "AF", "LLM"))
+df_trees <- read_delim("test/flashforward/out/unrooted_trees.txt",
                        show_col_types = FALSE,
                        col_names = c("gene", "target", "alphabet", "model", "tree")) %>% 
-  mutate(model=factor(model, levels=models))  
-  # filter(model!="FTPY")
+  mutate(model=factor(model, levels=models)) %>% 
+  filter(!is.na(tree), model %in% models_draft) %>% 
+  # some FM trees have negative branch lengths
+  mutate(tree = gsub(":-", ":", tree)) 
+# 722*8*4=23104 trees
 
 # ML trees with selected model and LL
-glue("results/phylogeny/", params_dataset$phylo_dataset,"/trees/", params_dataset$seed, "_mltrees.txt")
-df_ml <- read_delim("results/old/test/trees/UP000005640_mltrees.txt", 
+df_ml <- read_delim("draft/mltrees.txt", 
                     delim = "\t", 
@@
 
-glue("results/phylogeny/", params_dataset$phylo_dataset, "/reco/disco")
-
-disco_fls <- list.files("results/old/test/reco/disco/", pattern = "output.nwk", full.names = T)
-
-disco_rf <- NULL
-for (file in disco_fls){
-  a <- read.tree(file)
-  rf <- TreeDist::RobinsonFoulds(sptree, a, normalize = T)
-  ntips <- sapply(a, function(x) length(x$tip.label))
-  disco_rf <- bind_rows(disco_rf, 
-                        tibble(ntips=ntips, rf=rf, bn=rep(basename(file), length(rf))))
-}
+disco_fls <- list.files("test/flashforward/reco/", pattern = "^disco", full.names = T)
 
@@
 
-# Compare trees with RF distance
-rf_df <- filter(df_trees, target=="common") %>% 
-  select(-alphabet) %>% 
-  inner_join(x=., y = ., by = c("gene", "target")) %>%
-  filter(!is.na(tree.x), !is.na(tree.y)) %>%
-  rowwise() %>%
-  mutate(RF = phangorn::RF.dist(normalize = TRUE, 
-                                read.tree(text = tree.x), read.tree(text=tree.y)),
-         n_tips = length(read.tree(text = tree.x)$tip.label)) %>%
-  select(-tree.x, -tree.y)
-
-rf_text <- group_by(rf_df, model.x, model.y) %>% 
-  # filter(as.integer(model.x)<=as.integer(model.y)) %>% 
-  summarise(value=median(RF, na.rm=T))
-
-# get BS distribution
-df_bs <- fortify(ts) %>%
-  filter(!isTip) %>%
-  mutate(support=as.numeric(label)) %>%
-  separate(.id, c("gene", "target", "alphabet", "model")) %>%
-  filter(model!="QT") %>% 
-  filter(!is.na(support)) %>% 
-  select(gene, target, model, support)
+disco_rf <- get_disco_rf(disco_fls, sptree)
+df_bs <- get_bs_df(ts)
 
@@
   enframe() %>%
-  separate(name, c("gene", "target", "alphabet", "model")) %>%
+  separate(name, c("gene", "target", "alphabet", "model"), sep = "_") %>%
   rename("n_tips"="value")
 
-n_sps <- sapply(ts, function(x) 
-  length(unique(taxidmap[taxidmap$target %in% x$tip.label, ]$Tax_ID))) %>% 
+var_df <- sapply(ts, function(x) { 
+                 a <- phytools::midpoint_root(x) 
+                 var(ape::dist.nodes(a)[a$edge[1,][1],1:length(a$tip.label)])
+                 }) %>% 
   enframe() %>%
-  separate(name, c("gene", "target", "alphabet", "model")) %>%
-  rename("n_sps"="value")
-
-var_df <- sapply(ts, function(x)
-  var(adephylo::distRoot(phytools::midpoint_root(x)))) %>% 
-  enframe() %>%
-  separate(name, c("gene", "target", "alphabet", "model")) %>%
+  separate(name, c("gene", "target", "alphabet", "model"), sep="_") %>%
   rename("variance_r2t"="value")
@@
 # Reconciliation ranger+TCS scores
-glue("results/phylogeny/", params_dataset$phylo_dataset, "/reco/", params_dataset$seed, "_notung.tsv")
-reco <- read_delim("results/old/test/reco/UP000005640_notung.tsv", 
-                   show_col_types = FALSE, col_names = c("gene", "dups", "losses")) %>% 
-  separate(gene, c("gene", "target", "alphabet", "model")) %>% 
+# glue("results/phylogeny/", params_dataset$phylo_dataset, "/reco/", params_dataset$seed, "_notung.tsv")
+reco <- read_delim("test/flashforward/out/notung.txt", 
+                   show_col_types = FALSE, 
+                   col_names = c("gene", "target", "alphabet", "model", "dups", "losses")) %>% 
+  left_join(n_tips) %>% 
   left_join(var_df) %>% 
-  left_join(n_tips) %>% 
-  left_join(n_sps) %>% 
   mutate(model=factor(model, levels=models),
          n_events = (dups+losses)/n_tips) %>% 
+  filter(model %in% models_draft) %>% 
   rename("targets"="target")
@@
 ```{r fig3}
+xax <- ifelse(df_model$algorithm=="Dist.", "italic", "plain")
 
 plot_DL <- reco %>% 
-  filter(model!="FTPY") %>% 
-  ggplot(aes(model, n_events, fill=targets)) +
+  left_join(df_model) %>% 
+  ggplot(aes(weave_factors(model, data), n_events, 
+             color=algorithm, fill=targets)) + 
+  labs(y="(D+L)/# Tips", x="",fill="Target sets") +
   geom_boxplot(outlier.size = .1) +
-  labs(y="(D+L)/# Tips", x="",fill="Target sets") +
-  # ggthemes::scale_fill_solarized() +
+  scale_color_manual(values = c("black", "black"),
+                     labels = c("ML", expression(italic("Dist.")))) +
+  guides(x = "axis_nested", 
+         colour=guide_legend(title = "Method", override.aes=list(linetype = 0, size=0))) +
   scale_fill_manual(values = palettes_method) +
-  theme(legend.position = "none")
+    theme(
+    axis.text.x = element_text(face = xax),
+    ggh4x.axis.nestline.x = element_line(color="grey"),
+    ggh4x.axis.nesttext.x = element_text(colour = "grey30", 
+                                         face = "bold", size=10)
+  )
 
 plot_rf <- disco_rf %>% 
-  filter(ntips>10) %>% 
-  separate(bn, c("seed", "targets", "alphabet", "model")) %>% 
   mutate(model=factor(model, levels=models)) %>% 
-  ggplot(aes(model, rf, fill=targets)) + 
-  geom_boxplot() +
+  filter(nsps>length(sptree$tip.label)/2) %>%
+  # separate(bn, c("seed", "targets", "alphabet", "model")) %>% 
+  left_join(df_model) %>% 
+  ggplot(aes(weave_factors(model, data), rf, 
+             color=algorithm, fill=target)) + 
+  labs(x="", y="Norm. RF to species tree", fill="Target sets") +
+  geom_boxplot(outlier.size = .1) +
+  scale_color_manual(values = c("black", "black"),
+                     labels = c("ML", expression(italic("Dist.")))) +
+  guides(x = "axis_nested", 
+         colour=guide_legend(title = "Method", override.aes=list(linetype = 0, size=0))) +
   scale_fill_manual(values = palettes_method) +
-  labs(x="", y="Norm. RF to species tree", fill="Target sets")
+    theme(
+    axis.text.x = element_text(face = xax),
+    ggh4x.axis.nestline.x = element_line(color="grey"),
+    ggh4x.axis.nesttext.x = element_text(colour = "grey30", 
+                                         face = "bold", size=10)
+  )
 
-nodes_clades <- fortify(sptree) %>% 
-  # filter(!isTip) %>% 
-  mutate(ordered = rank(y)) %>% 
-  mutate(label_int = fct_reorder(label, y)) %>%
-  select(label_int, ordered)
-
-apro_files <- list.files("results/old/test/reco/", 
-                         pattern = "*support.nwk", full.names = TRUE)
+apro_files <- list.files("test/flashforward/reco/", 
+                         pattern = "^apro*", full.names = TRUE)
 
@@
 
-trees_df <- fortify(apro_trees) 
-nodes_df <- trees_df %>% 
-  group_by(.id) %>% 
-  mutate(ordered = rank(y)) %>% 
-  left_join(nodes_clades) %>% 
-  separate(.id, c("targets", "model")) %>% 
-  filter(!isTip, label!="") %>% 
-  mutate(label = gsub("\\[|\\]|\\'", "", str_replace_all(label, "[a-z]{1,2}[0-9]=", ""))) %>% 
-  separate(label, c("pp1", "pp2", "pp3", "f1", "f2", "f3", "q1", "q2", "q3"), 
-           ";", convert = TRUE) %>% 
-  mutate(freq=f1+f2+f3, 
-         model=factor(model, levels=models))
+trees_df <- get_apro_stats(apro_trees)
 
-plot_apro <- nodes_df %>%
-    filter(model!="FTPY") %>% 
-  ggplot(aes(model, q1, fill=targets)) + 
-  geom_boxplot() + 
+plot_apro <- trees_df %>% 
+  left_join(df_model) %>% 
+  ggplot(aes(weave_factors(model, data), q1, 
+             color=algorithm,fill=targets)) + 
+  labs(x="", y="Quartet support", fill="Target sets") + 
+  geom_boxplot(outlier.size = .1) +
+  scale_color_manual(values = c("black", "black"),
+                     labels = c("ML", expression(italic("Dist.")))) +
+  guides(x = "axis_nested", 
+         colour=guide_legend(title = "Method", override.aes=list(linetype = 0, size=0))) +
   scale_fill_manual(values = palettes_method) +
-  labs(x="", y="Quartet support", fill="Target sets")
-
-# firstrow <- (plot_rf + plot_DL + 
-#     plot_layout(guides = "collect"))
-# 
-# secondrow <- ( + 
-#     plot_layout(guides = "collect", widths = c(1,2)))
+    theme(
+    axis.text.x = element_text(face = xax),
+    ggh4x.axis.nestline.x = element_line(color="grey"),
+    ggh4x.axis.nesttext.x = element_text(colour = "grey30", 
+                                         face = "bold", size=10)
+  )
 
@@
   theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"),
-        legend.position = "bottom")
+        legend.position="bottom",
+        legend.margin=margin(0,0,0,0),
+        legend.box.margin=margin(-25,0,-10,0))
 
-ggsave("draft/figures/fig3.pdf", fig3, width = 10, height = 4)
-# reco_rnkgs_plot <- reco %>% 
-#   group_by(gene, target) %>% 
-#   slice_min(n_events, with_ties = T) %>% 
-#   group_by(target, model) %>% 
-#   count() %>% 
-#   ggplot(aes(target, n, fill=model)) + 
-#   geom_bar(stat = "identity") +
-#   geom_text(aes(label=n), position = position_stack()) +
-#   scale_fill_manual(values = palettes_model)
+ggsave("draft/figures/Fig3.pdf", fig3, width = 9, height = 2.8, device = cairo_pdf)
 ```
@@
 ```{r FigS1}
+# Compare trees with RF distance
+rf_df <- get_rf_df(df_trees)
+rf_text <- group_by(rf_df, model.x, model.y) %>% 
+  # filter(as.integer(model.x)<=as.integer(model.y)) %>% 
+  summarise(value=mean(RF, na.rm=T))
+
 plot_bs <- df_bs %>% 
+  filter(model!="FM") %>% 
   mutate(targets=target,
@@
 var_r2t <- reco %>% 
-  filter(model!="FTPY", targets=="common") %>% 
-  ggplot(aes(variance_r2t, model)) +
+  filter(targets=="common", model %in% models_draft) %>% 
+  ggplot(aes(variance_r2t, fct_reorder(model, variance_r2t))) +
   geom_violin(scale = "width", color="black", fill="grey40") +
   scale_x_continuous(limits = c(0,.5), expand = expansion(mult=0, add = c(0,0.05))) +
-  labs(x="Variance Root-to-tip distance") +
+  labs(x="Variance Root-to-tip distance", y="") +
   # scale_fill_manual(values = palettes_model) +
@@
   scale_fill_distiller(palette="YlGnBu", direction = -1, limits = c(0,1), 
-                       name="Median RF") + 
-  labs(x="Median RF Distance", y="") +
+                       name="Mean RF") + 
+  labs(x="Mean RF Distance", y="") +
   theme_classic() +
@@
 
-df_red <- select(df_filt, query, target, singleton)
+union_trees <- df_trees %>% 
+  filter(target=="union", model %in% models_draft)
+df_dist <- distance_to_seed(union_trees)
 
-res <- NULL
-for (idx in 1:nrow(df_trees)){
-  t <- ts[[idx]]
-  seed <-  t$tip.label[grepl(df_trees$gene[idx], t$tip.label)]
-  nn <- as.matrix(adephylo::distTips(t, method = "nNodes"))[seed, ]
-  max_dist <- max(nn)
-
-  row <- enframe(nn, name = "target", value="nn_dist") %>% 
-    mutate(query=seed, 
-           max_dist=max_dist,
-           targets=df_trees$target[idx],
-           alphabet=df_trees$alphabet[idx],
-           model=df_trees$model[idx]) %>% 
-    left_join(df_red, by=c("query", "target"))
-  res <- bind_rows(res, row)
-}
-
-# for what this should be grouped?
-plot_dist <- res %>% 
-  filter(query!=target) %>% 
+plot_dist <- df_dist %>% 
+  left_join(select(df_filt, query, target, singleton), 
+            by = c("query", "target")) %>%
+  filter(query!=target, !is.na(singleton)) %>% 
+  group_by(query) %>% 
+  mutate(max_dist = max(nn_dist)) %>% 
   group_by(query, singleton, model) %>%
@@
   ggplot(aes(median_dist, model,  fill=singleton)) + 
-  ggdist::stat_halfeye(alpha=.8, point_colour = NA, .width = 0, height=.6, 
-                       position = position_nudge(y=.15)) +
+  # ggdist::stat_halfeye(alpha=.8, point_colour = NA, .width = 0, height=.6, 
+  #                      position = position_nudge(y=.15)) +
   geom_boxplot(
-    width=.2,
-    position = position_dodge(),
+    # width=.2,
+    # position = position_dodge(),
     # removing outliers
@@
   theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
-ggsave("draft/figures/figS1.pdf", fig_s1, width = 10, height = 6)
+
+ggsave("draft/figures/Fig_S1.pdf", fig_s1, width = 10, height = 6)
 ```
@@
 ```
+
+### Heteroscedascity of DL normalization
+
+```{r}
+lm <- lm(dups+losses ~ n_tips, data = reco)
+reco$res <- lm$residuals
+
+tmp_plot <- ggplot(reco, aes(y = res, x = n_tips, fill=target)) + 
+  geom_point(pch=21) +
+  facet_wrap(~target) + 
+  geom_abline(slope = 0) + 
+  scale_fill_manual(values = palettes_method) + 
+  labs(x="Num. of Tips", 
+       y="Residuals of: D+L ~ Num. of Tips") + 
+  theme(legend.position = "none")
+
+ggsave("draft/figures/review_hetero_TMP.png", tmp_plot,
+       width = 4, height = 3)
+```

### 2024-12-04 workflow/scripts/analyze_orthobench.R

@@ -12,7 +12,7 @@ eval <- snakemake@params[["evalue"]]
 min_size <- snakemake@params[["min_seqs"]]
 # min_size <- 5
 dataset <- snakemake@wildcards[["db"]]
-# dataset <- "oma"
+# dataset <- "eggnog"
 
 seed_fa <- Biostrings::readAAStringSet(snakemake@input[["fa"]])
 # seed_fa <- Biostrings::readAAStringSet("results/ortho_bench/eggnog/db/db.fa")
@@ -48,6 +48,7 @@ blast_bench <- read_delim(snakemake@input[["blast"]], col_names = blast_columns,
   mutate(method = "blast")
 
 comb_df <- true_set %>% 
+  # head(1000) %>% 
   group_by(dataset) %>%
   filter(n() > min_size) %>% 
   summarize(pairs = list(combn(entry, 2, simplify = FALSE)), .groups = "drop") %>%
@@ -69,22 +70,28 @@ accuracy_df <- comb_df %>%
   summarise(recall_blast = sum(!is.na(qcov_blast))/n(),
             recall_fs = sum(!is.na(qcov_fs))/n(), 
             n = n_distinct(entry1),
-            avg_len = mean(length))
+            avg_len = mean(length),
+            avg_qcov_blast = mean(qcov_blast, na.rm = T),
+            avg_qcov_fs = mean(qcov_fs, na.rm = T))
 
 
 qcov_box <- comb_df %>% 
   pivot_longer(cols = c(qcov_blast, qcov_fs), values_to = "qcov") %>% 
-  mutate(class = case_when(is.na(method_blast) & is.na(method_fs) ~ "None",
+  mutate(class = case_when(is.na(method_blast) & is.na(method_fs) ~ NA,
                            is.na(method_fs) ~ "only Blast",
                            is.na(method_blast) ~ "only Foldseek",
                            .default = "Common")) %>% 
+  filter(!is.na(class)) %>% 
   ggplot(aes(qcov, class, fill=class)) + 
-  geom_boxplot() + 
-  geom_text(data = . %>% count(class), aes(x = 15, label = paste0("n: ", n/2)), nudge_y = +0.2) +
+  geom_violin() +
+  # facet_wrap(~name) +
+  geom_text(data = . %>% count(class), 
+            aes(x = 15, label = paste0("n: ", n/2)),
[…]

### 2024-12-04 workflow/scripts/functions.R

@@ -10,12 +10,22 @@ fs_columns <- c("query","target","pident","length","mismatch","gapopen",
                 "qstart","qend","sstart","send","evalue","bitscore",
                 "lddt","alntmscore","rmsd","prob","qcov","tcov")
 
-models <- c("LG", "FM", "QT",  "FT", "FTPY", 
-            "3Di", "GTR", "AF", "LLM", "part")
+models <- c("LG", "FM", "QT", "part", "FT",
+            "FTPY", "3Di", "GTR", "AF", "LLM")
 
-palettes_model <- c("#BF360C", "#FFA000", "#FFA000", "#D4E157","#646B00",
-                    "#75C323", "#40A79D", "#33ceff", "#3396ff", "purple")
+palettes_model <- c("#BF360C", "#FFA000", "#FFA000", "purple", "#D4E157","#646B00",
+                    "#75C323", "#40A79D", "#33ceff", "#3396ff")
 names(palettes_model) <- models
+
+
+df_model <- tibble(model = factor(models), 
+       data = c(rep("AA", 3), "Mixed", rep("3Di", 6)),
+       algorithm = c("ML", rep("Dist.", 2), "ML", 
+                     rep("Dist.", 2), rep("ML", 4))) %>% 
+  filter(!model %in% c("QT", "FTPY") )
+df_model$data <- factor(df_model$data, levels = c("AA", "Mixed", "3Di"))
+df_model$algorithm <- factor(df_model$algorithm, levels = c("ML", "Dist."))
+
 palettes_model_d <- colorspace::darken(palettes_model, amount = 0.6)
 names(palettes_model_d) <- models
 # "#FF0000" "#00A08A" "#F2AD00" "#F98400" "#5BBCD6"
@@ -137,4 +147,93 @@ plot_dom <- function(tree, seed, gff, prot, singleton_df, title="") {
   
   return (list(a, b))
 }
+
+
+get_disco_rf <- function(disco_fls, sptree) {
+  disco_rf <- NULL
+  for (file in disco_fls){
+    a <- read.tree(file)
+    rf <- TreeDist::RobinsonFoulds(sptree, a, normalize = T)
+    nsps <- sapply(a, function(x) length(unique(x$tip.label)))
+    disco_rf <- bind_rows(disco_rf, 
+                          tibble(nsps=nsps, rf=rf, bn=rep(basename(file), length(rf))))
+  }
+  disco_rf <- disco_rf %>% 
[…]

### 2024-12-04 workflow/scripts/tmp.R

@@ -1,0 +1,63 @@
+a <- df %>% 
+  # head(10000) %>% 
+  mutate(qcov = log10(qcov_Blast/qcov_Foldseek),
+         # evalue_ratio = evalue_Blast/evalue_Foldseek,
+         pident = log10(pident_Blast/pident_Foldseek)) %>% 
+  # filter(pident_ratio < 1)
+  select(c("qcov", "pident")) %>% 
+  pivot_longer(cols = everything())
+
+ggplot(a, aes(value, color=name)) +
+  geom_density() + 
+  scale_x_continuous(limits = c(-1, 1)) +
+  geom_vline(xintercept = 0) + 
+  scale_color_manual(values = c("#F2CB05", "#0378A6")) + 
+  labs(x="Log10(Blast/Foldseek)") +
+  theme(legend.position = "bottom", legend.title = element_blank())
+
+
+df_ml %>% 
+  filter(model!="LG") %>% 
+  # filter(gene=="AF-A0A075B6U4-F1") %>% 
+  group_by(gene, targets) %>% 
+  mutate(rank = factor(dense_rank(BIC))) %>% 
+  select(BIC, model, rank) %>% 
+  ungroup() %>% 
+  count(model, rank) %>% 
+  ggplot(aes(rank, model, fill=n)) + 
+  # geom_point(aes(size=n), pch=21) +
+  geom_tile() +
+  geom_text(aes(label=n), color="white") +
+  scale_fill_viridis_c() + 
+  coord_fixed(expand = F) +
+  labs(x="Rank", y="Model") + 
+  theme(legend.position = "none")
+
+
+dl <- read_delim("results/old/test/reco/UP000005640_DTL.tsv") %>% 
+  rename("gene"="id", "target"="targets")
+
+df_trees <- read_delim("results/old/test/trees/UP000005640_unrooted_trees.txt",
+                       show_col_types = FALSE,
+                       col_names = c("gene", "target", "alphabet", "model", "tree")) %>% 
+  rowwise() %>% 
+  mutate(model=factor(model, levels=models),
+         n_tips = length(read.tree(text = tree)$tip.label)) 
+
+a <- left_join(df_trees, dl) %>% 
+  filter(!is.na(reco_score))
+
+lm <- lm(dups+losses ~ n_tips, data = a)
+a$res <- lm$residuals
+
+tmp_plot <- ggplot(a, aes(y = res, x = n_tips, fill=target)) + 
[…]

### 2024-12-06 workflow/notebooks/draft.Rmd
update README

@@
   guides(colour = guide_legend(order = 1),  
-         alpha = guide_legend(order = 2)) +
+         alpha = guide_legend(title = "CATH Level", order = 2)) +
   theme(panel.border = element_rect(fill = NA), 
@@
 ```{r fig3}
-xax <- ifelse(df_model$algorithm=="Dist.", "italic", "plain")
+xax <- ifelse(df_model$algorithm=="Distance", "italic", "plain")
 
@@
   scale_color_manual(values = c("black", "black"),
-                     labels = c("ML", expression(italic("Dist.")))) +
+                     labels = c("ML", expression(italic("Distance")))) +
   guides(x = "axis_nested", 
@@
   scale_color_manual(values = c("black", "black"),
-                     labels = c("ML", expression(italic("Dist.")))) +
+                     labels = c("ML", expression(italic("Distance")))) +
   guides(x = "axis_nested", 
-         colour=guide_legend(title = "Method", override.aes=list(linetype = 0, size=0))) +
+         colour=guide_legend(title = "Method", 
+                             override.aes=list(linetype = 0, size=0))) +
   scale_fill_manual(values = palettes_method) +
@@
 
-```{r FigS1}
+```{r FigS2}
 # Compare trees with RF distance
@@
   ggplot(aes(model.x, model.y)) +
-  geom_tile(aes( fill=value)) +
+  geom_tile(aes(fill=value)) +
   geom_text(aes(label=round(value, 2)), color="white") +
@@
 
-fig_s1 <- ((plot_bs / rf_plot) + 
+fig_s2 <- ((plot_bs / rf_plot) + 
              plot_layout(heights = c(2,1)) | 
@@
 
-ggsave("draft/figures/Fig_S1.pdf", fig_s1, width = 10, height = 6)
+ggsave("draft/figures/Fig_S2.pdf", fig_s2, width = 10, height = 6)
 ```
 
-```{r FigS2}
+# ```{r FigS2}
 # pp1_plot <- nodes_df %>%
@@
 # ggsave("draft/figures/figS2.pdf", fig_s2, width = 12, height = 6, device = cairo_pdf)
-```
+<!-- ``` -->
 

### 2024-12-06 workflow/scripts/functions.R

@@ -20,11 +20,11 @@ names(palettes_model) <- models
 
 df_model <- tibble(model = factor(models), 
        data = c(rep("AA", 3), "Mixed", rep("3Di", 6)),
-       algorithm = c("ML", rep("Dist.", 2), "ML", 
-                     rep("Dist.", 2), rep("ML", 4))) %>% 
+       algorithm = c("ML", rep("Distance", 2), "ML", 
+                     rep("Distance", 2), rep("ML", 4))) %>% 
   filter(!model %in% c("QT", "FTPY") )
 df_model$data <- factor(df_model$data, levels = c("AA", "Mixed", "3Di"))
-df_model$algorithm <- factor(df_model$algorithm, levels = c("ML", "Dist."))
+df_model$algorithm <- factor(df_model$algorithm, levels = c("ML", "Distance"))
 
 palettes_model_d <- colorspace::darken(palettes_model, amount = 0.6)
 names(palettes_model_d) <- models
